### 2016-04-07 article.Rmd
submitted version

---
title: 'Disclosing funding sources for open access publication fees: the Open APC initiative'
author:
- Najko Jahn, Bielefeld University Library, Bielefeld University
- <a href="http://orcid.org/0000-0002-5111-2788">Marco Tullney</a>, Technische Informationsbibliothek (TIB) - German National Library of Science and Technology, Hannover, Germany
date: '7 April 2016'
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: peerj.sty
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: no
    theme: readable
bibliography: literatur.bib
csl: frontiers.csl
abstract: Publication fees in open access publishing hold a prominent place on the agenda of researchers, policy-makers, and academic publishers. This paper contributes to the evolving empirical basis on open access funding. It describes the Open APC initiative, in which German universities and research organizations share their expenditures for publication fees. As method, the initiative uses existing open data tools to aggregate and disseminate institutional spending on open access publication fees. In total, 29 German research organizations self-reported funding of 6,279 open access journal articles, which amounted to 8,039,339 €. The average payment for each article was 1,280 €, and the median payment 1,209 €. Our data-set comprises only 53 articles in hybrid journals. With an indexing coverage of 99 %, the findings reveal that the DOI agency CrossRef provides both comprehensive bibliographic coverage of the funded open access journal literature and disambiguated names of journal titles and publishing houses. We show that authority control of these bibliographic information is particularly relevant for the comparative study of the economical effects of open access publishing.
---

```{r, echo =FALSE}

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 12,
  fig.height = 8
)
options(scipen = 999, digits = 2)
knitr::knit_hooks$set(inline = function(x) {
      if(is.numeric(x)){
          return(prettyNum(x, big.mark=","))
      }else{
          return(x)
       }
   })
```

# Introduction{.unnumbered}

Publication fees, often paid by funders or universities, are a widely discussed open access business model. Yet, how and to what extent these activities are effective in terms of the number of supported research articles and associated costs remains under debate. This paper describes the Open APC initiative[^2], in which German universities and research organization share spendings on open access publication fees, and how it is currently implemented. More specifically, it addresses three problem areas when studying the economical effects of open access publishing: fragmentation of open access funding, variable pricing schemes and the comparison across research institutions. Such an approach extends methods and improves data collection activities for researchers and practitioners, as well as contribute to a better understanding of factors affecting the analysis of publication fees in open access publishing.

The rise of open access journals  matches the increasing relevance of publication fees in academic publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015]. To cover these fees, authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012]. However, collecting information about what was funded is in most cases difficult. One reason why payments made for open access journal publications are often hard to track is that, on the one hand, they are fragmented across the budgets of funding agencies, research institutions, and libraries, and, on the other hand, taken from personal budgets. Furthermore, open access funding mostly exists in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011]. Personal budgets stand in contrast with those support structures and are likely used to cover low price publication fees [@Solomon_2011; @Bj_rk_2015]. Along with the fragmentation of payments, funding for open access publications lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015]. It also remains unclear which factors contribute to price formation.[^4b] While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015]. Other factors leading to a complex landscape of variable pricing schemes [@Pinfield_2015] include submission or page charges [@BJ_RK_2012]. Hybrid journals substantially add to this complexity, because comprehensive offset systems to avoid paying for the same article twice, through subscription and publication fee, are rare, which, in turn, leads to the phenomena of "double dipping" in scholarly publishing [@Pinfield_2015].

This complex situation of fee-based open access publishing creates difficulties for researchers and practitioners alike. Because of fragmented payments, the extent of funding remains unclear. To increase transparency, some research funders have begun collecting and disclosing expenditures for open access journal articles as open data. As per definition, open data is data that "can be freely used, modified, and shared by anyone for any purpose" [@OpenDataHandbook]. Therefore, opening up information about the funding of open access journal articles promises to enhance the discussion about current and future business models in academic publishing. To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc], who both released data on publication fees they had funded. The British not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015]. Disclosed as publicly available spreadsheets, these data-sets self-report expenditures along with bibliographic information, including title, journal and publisher, persistent identifier to the publisher's version, and a link to a deposit in a subject repository. Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates. In the case of the Wellcome Trust, crowd-sourcing data cleaning activities through a Google spreadsheet in combination with checks against bibliographic sources massively improved the spending data (see comments in @wellcome_apc).

The open access landscape in Germany, which is the focus of this paper, shares the general problems of in-transparency regarding funding schemes and costs as discussed above. The Deutsche Forschungsgemeinschaft (DFG), the largest research funder in Germany, has been encouraging open access publishing since years. It launched its "Open-Access Publishing" program in 2009 that has strongly influenced the support of open access publication fees through funds managed by university libraries.[^3] With this program, the DFG aims to help universities to establish support structures for publishing in open access journals where authors are requested to pay a publication fee. To reduce administrative burdens, grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions. Examples include central invoicing schemes and related agreements between university libraries and publishers. Grantees are also required to report the institutional publication output and their fees paid for open access journal articles to the DFG on a regular basis, and to present the university-wide strategy to sustain the funds when DFG’s initial support runs out lately in 2019. The DFG enforces a set of criteria grantees have to comply with, leading to similar implementations for supporting open access publishing across German universities: these criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding VAT) [@Fournier_2013]. Research institutes organized in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this funding program, contributing to the diversity of schemes in Germany. In response, some organizations have adopted similar processes to support authors. The Max-Planck-Gesellschaft operates their long-lasting open access activities, including handling spending and publisher agreements centrally, through the Max Planck Digital Library [@Schimmer_2013; @Sikora_2015], while the Leibniz-Gemeinschaft set up a dedicated open access fund in 2016.

The growing share of articles published in fee-based open access journals in recent years has led to calls for an unified approach towards funding of publication fees. The Allianz der Wissenschaftsorganisationen[^10c], representing all major research organizations in Germany, thus marks transparency as a major means to sustain an "adequate open access publication system" [@allianz]. However, there are various ways to achieve this goal. The existing approaches in Austria and the United Kingdom have one institution in charge to collect and analyze the data. The history of the Open APC initiative is rather bottom-up: In May 2014, Bielefeld University Library began to share its expenditures for publication fees. The library put its approach to the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11] as a basis for discussion, and invited others to participate. Reflecting the increasing demand for publicly available data, contributions from Universität Regensburg and Universität Hannover followed soon after. As of writing, 29 universities and research institutes voluntary reported their data to the Open APC initiative to be included into a unified data-set of all expenditures.

In this paper, we present the technical workflow of the Open APC initiative. We describe how the data-set is curated and which tools are used in order to produce, disseminate, and preserve the Open APC spending data. Presenting our results, we will particularly discuss how and to what extent the use of the CrossRef index accommodates the demand of disambiguated bibliographic information about journals and publishers when reporting about funding of open access journal articles. CrossRef, a Digital Object Identifiers (DOI) registration agency for scholarly literature, associate these persistent links with metadata CrossRef members such as publishers and research societies contribute.<!--  Our analysis of the data contributed so far illustrates possible ways of using this data-set.  We are able to show the importance of reusable open data and encourage other possible data contributors to also publish their payment information through this initiative. The goal is to strengthen the insight on open access funding, enabling institutions to better plan for a transition to open access.
-->

# Methods and Materials{.unnumbered}

The major goal of the Open APC initiative is to gain insights into expenses for publication fees by collecting data directly from the institutions paying on behalf of the authors they support. These institutions can report best on the most important part: How much they have spent for each article.  For this aim, the Open APC initiative applies open data methods, designed around the idea of sharing the collected data as permissive as possible. At its core is a data pipeline to collect and to process the contributed data. In this section, we discuss the general idea of this pipeline and describe its stages: data submission (by spending institutions), merging contributions, re-using data, preserving data, and engaging with our communities to increase participation.

## General idea{.unnumbered}

Our approach drew on general open data guidelines that state to "keep things simple" and to "engage early and engage often" with data providers and potential users [@OpenDataHandbook]. We therefore chose a simple data scheme, built our pipeline around the popular social coding platform GitHub, and re-used information from bibliographic indexes.

We followed good practice of other open data projects to guarantee that the data is as open as possible. From the beginning, we emphasized that this also includes information on submissions -- date of submission, contributors, etc. -- and open file formats.
Because institutions should be allowed to self-report data and updates at different times, tracking initial submissions and data-set updates is particularly important.

All data and documentation reside, therefore, on GitHub. Git, a distributed version control system, powers GitHub and allows people to collaborate on software projects. Git keeps a log of changes made in the source code and manages to synchronize local copies of the very same software repository. Because of its distributed and social characteristics, Git in general and GitHub in particular are suitable for researchers to mutually curate and share research artifacts including data-sets, analyses, or visualizations [@Ram_2013].

In addition to GitHub, bibliographic indexes are an essential part of our data pipeline. We use CrossRef to normalize bibliographic metadata. This ensures automatic authority control for journal and publisher titles, which are the most appropriate levels of aggregations when analyzing expenditures for open access articles [@Pinfield_2015]. We also check automatically the indexing status of each article in Europe PubMed Central, a large bibliographic index for life-science literature, the multidisciplinary database Web of Science, and the open access source Directory of Open Access Journals (DOAJ).

## Data Submission{.unnumbered}

Participants use a template to self-report data (see Table 1). Data structured in this way enables the fetch of additional data from external sources, and ensures that the merge into a single data-set succeeds. By collecting as little data as possible from the institutions, this approach allows for changes and the addition of new fields if required.

|Variable               |Description                              |Required  |
|:----------------------|:----------------------------------------|:-------- |
|`institution`          |Top-level organisation e.g. MPG          | mandatory|
|`period` 		          |Year of APC payment (YYYY)               | mandatory|
|`euro` 		            |The amount paid in EURO (incl. VAT)      | mandatory|
|`doi`                  |Digital Object Identifier                | mandatory|
|`is_hybrid`            |Published in a access journal?           | mandatory|
|`publisher`          	|Name of publication house                | optional |
|`journal_full_title`   |Full name of periodical                  | optional |
|`issn` 		            |International Standard Serial Number.    | optional |

Table: Mandatory and optional data elements for disclosing funding of open access publication fees

The data scheme reflects how the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc [@Lawson_data_2015] organized data about expenses for open access publication fees. The project requests items that are already present at the research institutions for internal reporting purposes [@Pinfield_2015]. Guidance[^23] available to the participants furthermore refers to the principles of *tidy data* [@tidy-data]. This responds to issues encountered in the United Kingdom, when the inexperienced use of spreadsheet software like Excel lead to misaligned tables [@woodward_2014]. The principle of tidy data intends to reduce data cleaning efforts before statistical analyses. The three principles of tidy data state that, firstly, each variable must form a column, secondly, each observation must form a row, and, finally, each type of observational unit must form a table.  To prevent messy data, it is also crucial that unknown values are not left empty. In our case, we use the R convention `NA` for handling those values.

Data contributors collect the data in different systems. After preparing the data, it must be exported in the csv standard, a plain text file format for tabular data supported by most spreadsheet software that can be easily logged with version control systems like Git.

To illustrate a sample data contribution displaying mandatory information in the csv format:

```
"Hannover U",2013,1241.02,"10.1371/journal.pone.0063501",FALSE
```

Contributors store the csv file into a folder that also contains a README file. Written in plain text or Markdown, the README contains information about the data-set and the contributing institution. The submission itself is facilitated through GitHub's pull request mechanism: contributors fork the initiative's data, add their contributions, upload their modified copy to their repository, and request importing their data into the initiative's repository by alerting the maintainers through sending a "pull request".

## Data curation{.unnumbered}

<!--Bielefeld University Library acts as curator and merges the contributed spreadsheets into one general data-set.--> By using open software, we, as curators, retrieve additional data elements based on the article's DOI. This avoids manual cleaning efforts the other initiatives were faced with. Because the DOI is at the center of the data curation, data normalization starts with a check of the DOI columns for possible duplicates and white space. CrossRef is then queried for article metadata matching the particular DOI.

CrossRef provides several APIs to search for bibliographic information it indexes during DOI registration. In our case, we query the CrossRef REST API[^18] by using content negotiation. As resource type, we request the format `application/vnd.crossref.unixsd+xml`, which main function is to support text mining activities.[^5] The advantage of this XML format is that it distinguishes full and abbreviated journal titles as well as the media types of ISSNs, the International Standard Serial Number used to identify journals. It also contains license information and disambiguated publisher information, thus avoiding confusion about licensing and naming of publisher houses [@woodward_2014].

As a client, we use the R package `rcrossref` [@rcrossref] developed and maintained by the rOpenSci initiative.[^6] With the function `cr_cn`, the client supports all linking types. Table 2 summarizes the data elements we retrieve:

|Source     	|Data element  |Description                     |
|:--------------|:-------------------|:------------------------------------|
|CrossRef   	|`publisher` |Title of Publisher             |
|CrossRef   	|`journal_full_title` |Full Title of Journal  |
|CrossRef   	|`issn` |International Standard Serial Numbers (collapsed) |
|CrossRef   	|`issn_print` |ISSN print |
|CrossRef   	|`issn_electronic`  |ISSN electronic        |
|CrossRef   	|`license_ref`  |License of the article     |
|CrossRef   	|`indexed_in_crossref`  |Indexed in CrossRef? (logical)    |
|EuropePMC    	|`pmid`  |PubMed ID                 |
|EuropePMC    	|`pmcid` |PubMed Central ID         |
|Web of Science |`ut` |Web of Science record ID             |
|DOAJ           |`doaj` |Is the journal indexed in the DOAJ? (logical)    |

Table: Sources used to automatically enrich the Open APC data-set

In addition to CrossRef, the indexing status for each article in Europe PubMed Central and the Web of Science is checked. Europe PubMed Central, one of the largest database for life science literature, offers a public RESTful web services to access more than 24 million records and 870,000 deposited open access articles [@epmc_2014]. Information from the Web of Science, which provides indexing of open access journal literature [@Walters_2011], is retrieved through the Thomson Reuters Article Match Retrieval Service[^8]. This web interface is available to Bielefeld University Library as part of its Web of Science subscription. Finally, an automatic match with the DOAJ, a comprehensive and openly available registry of open access journals, is performed.

After disambiguating and enriching the data, we append the rows into a single csv file that records all expenses. The data additions are logged with Git and pushed to the source code repository on GitHub.

## Re-use{.unnumbered}

We use the main README file of our data repository as general guide to the Open APC data collection and to present sample analyses. The README itself is written in R Markdown, an authoring format that allows for the combination of the R code, results, and text within one document.[^10] Compiled with knitr [@knitr], the README reflects new data submissions after every build. Sample results include tables summarizing the number of supported articles and costs per institution as well as figures that illustrate the distribution of publication fees.

A blog hosted on GitHub presents all new data contributions.[^12] Since it is technically based on Jekyll,[^13] a static site generator, we can use the same development cycle we have for producing the README files. We write blog posts, which include information about the providing institution and the contributed data-set, in R Markdown, build the posts with `knitr` and, then, log and deploy the files to GitHub with Jekyll, making it possible to reproduce the analytical steps.

## Preservation{.unnumbered}

The data-set itself is licensed under the Open Database License (ODbL) v1.0,[^13a] permitting re-use under conditions of attribution and share-alike. To preserve the collected data, a GitLab installation hosted by Bielefeld University and administrated by Bielefeld University Library, mirrors a copy of the GitHub repository and is accessible via a DOI.[^22]

To provide reference points to particular data contributions, unique version names refer to data submissions or other changes in the source code, which will be then distributed as releases through GitHub. This approach allows others to not only to refer, but also to re-use or self-archive particular snapshots of the data.

## Engagement{.unnumbered}

Besides technical measures, social aspects are crucial to make an open data initiative successful. For the Open APC initiative, we focus on engaging librarians and collaborating within existing networks. In Germany, the working group "Elektronisches Publizieren" of the Deutsche Initiative für Netzwerkinformation (DINI) coordinates these efforts.[^11] The Open APC initiative aims at increasing participation also through the use of social media, workshops, and regular community calls.

A GitHub wiki[^15] shares guidance and interim reports. If participants want to report bugs or propose new functionality, GitHub's "issues" mechanism involves users while keeping track of the discussion. To ensure a constructive environment for all participants, we refer to the code of conduct from Contributor Covenant.[^17]

<!--From our experience, the process described in this chapter has played an important role in getting many institutions to participate in voluntarily sharing information on their expenses for APCs. The data scheme has been revised at the end of 2015, and transformation to a new scheme has happened without any complications. Further adaption, e.g. adding subject classification if a suitable source can be identified, is possible. We encourage institutions and researchers to further work on the data pipeline while preserving its openness and the low entry barrier for new data contributors.-->

# Results{.unnumbered}

```{r, echo=FALSE, cache = FALSE}
# load data
my_apc <- read.csv("data/apc_de.csv", header = TRUE, sep =",", stringsAsFactors = FALSE)

# data cleaning
library(dplyr)
# remove Bielefelds Zero value
my_apc <- filter(my_apc, euro > 0)
# remove intech book chapters
my_apc <-  filter(my_apc, !is.na(journal_full_title))
# clean publisher titles, because metadata could not be fetched
my_apc$publisher[grep("BioMed Central", my_apc$publisher, ignore.case = T)] <- "Springer Science + Business Media"
my_apc$publisher[grep("Impact Journals, LLC", my_apc$publisher, ignore.case = T)] <- "Impact Journals LLC"
my_apc$publisher[grep("e-Century Publishing Corporation", my_apc$publisher, ignore.case = T)] <- "e-Century Publ. Corp."
my_apc$publisher[grep("Academic Journals", my_apc$publisher, ignore.case = T)] <- "Academic Journals"
# fix plos name change in journal titles
my_apc$journal_full_title <- gsub("PLoS", "PLOS", my_apc$journal_full_title)
my_apc$journal_full_title <- gsub("PloS","PLOS", my_apc$journal_full_title)
my_apc$journal_full_title <- gsub("PlOS","PLOS", my_apc$journal_full_title)
my_apc$journal_full_title <- gsub("Sustainibility","Sustainability", my_apc$journal_full_title)
my_apc$journal_full_title <- gsub("Surgeal","Surgical", my_apc$journal_full_title)
my_apc$journal_full_title <- gsub("Oncotarget","OncoTarget", my_apc$journal_full_title)
my_apc$journal_full_title <- gsub("BMC cancer","BMC Cancer", my_apc$journal_full_title)

```

```{r results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(scales)

```

## Cost Data{.unnumbered}

```{r}
apc_time <- my_apc %>%
  group_by(period) %>%
  summarize(articles = n())
```

On February 19th 2016,[^16a] the Open APC initiative covered `r nrow(my_apc)` articles, whose publication fees were centrally paid by `r length(unique(my_apc$institution))` German universities and research institutions. The number of supported open access journal articles, which were reported to this initiative, grew over the years (see Figure 1). While one institution disclosed `r filter(my_apc, period == min(my_apc$period)) %>% nrow()` payments made in `r as.character(min(my_apc$period))`, the majority shared their expenditures from 2013 onwards. With `r max(apc_time$articles)` articles, the year `r apc_time[apc_time$articles == max(apc_time$articles),"period"]` was best represented in our data-set. Because of a time lag between payments made and reporting them to the Open APC initiative, only `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institution were able to partly contribute their cost data for 2015 at the time of this analysis.

```{r articles_year, fig.width=6, fig.height=2.5,fig.cap="Growth of Open APC Initiative"}
apc_time <- my_apc %>%
  group_by(period) %>%
  summarize(articles = n())
ggplot(apc_time, aes(factor(period), cumsum(articles), group = 1)) +
  geom_line(stat = "identity", colour = "blue") +
  geom_point() +
  xlab("Year") +
  ylab("OA articles reported\n(cumulative sum)") +
  scale_y_continuous(limits = c(0,8000))
```

```{r}
iqr_df <-  my_apc %>%
  group_by(period) %>%
  summarise(iqr = mean(euro))
```

Among all articles, fees amounted to `r sum(my_apc$euro)` € including VAT, the average payment was `r mean(my_apc$euro)` € and the median value `r median(my_apc$euro)` € . Figure 2 shows the large price variation among the articles. The disclosed publication fees ranged from `r min(my_apc$euro)` € to `r max(my_apc$euro)` €. However, the average price paid varied somewhat during the period 2011 and 2014 (`r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - `r my_apc %>% filter(period == 2014) %>% summarise(mean(euro)) %>% round()` €). We also observe that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc))  *100`%) of the publication fees were paid in accordance with the DFG price cap of 2,000 €. Whereas related open data initiatives in Austria and the United Kingdom reported a large share of spending for hybrid journal articles, the situation in Germany is different: only  `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` articles in hybrid journals were reported by `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` research institutions, accounting for `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of the overall payments.

```{r payments_year, fig.width=6, fig.height=3,fig.cap="Payments per year"}
ggplot(my_apc, aes(factor(period), euro)) +
  geom_boxplot() +
  xlab("Year") +
  ylab("Publishing fee in EURO(log scale)") +
  scale_y_continuous(limits = c(0,8000))
```

The number of APC payments per institutions varied considerably (see Table 3). With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall submissions. In contrast, the two universities of technology, TU Clausthal and TU Ilmenau, who recently begun to set up support structures for fee-based open access journal articles, shared payments made for four articles each.

```{r results='asis'}
tt <- my_apc %>%
  group_by(institution) %>%
  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), Median = median(euro), Min = min(euro), Max = max(euro)) %>%
  arrange(desc(Articles))

colnames(tt) <- c("Institutions", "Articles funded", "Total", "Mean", "Median", "Minimum", "Maximum")
tt <- format(tt,big.mark=',')
pander::pandoc.table(tt, split.cells = 7, caption = 'Institutions self-reporting expenditures for open access publications (in €)', justify = 'lrrrrrr')
```

## CrossRef indexing{.unnumbered}

Along with the price information, participating institutions were required to identify funded articles by their DOI. They were reported for `r my_apc %>% filter(!is.na(doi)) %>% nrow()` out of `r nrow(my_apc)` articles. Of those, `r my_apc %>% filter(indexed_in_crossref == TRUE) %>% nrow()` were indexed in CrossRef, representing `r my_apc %>% filter(indexed_in_crossref == TRUE) %>% nrow() / my_apc %>% nrow() * 100`  % of all funded publications. The reasons why articles identified by a DOI were not registered with CrossRef differed. Some journals were not indexed by CrossRef at the time of our study but by the DOI agencies DataCite (Journal of new frontiers in spatial concepts published by KIT Scientific Publishing) and Medra (DIE ERDE: Journal of the Geographical Society of Berlin). In other cases, either the DOI did not refer to the full text despite the fact that the journal was indexed on a regular basis (compare <http://doi.org/10.1186/1471-2105-13-S19-S7> with <http://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-13-S19-S7>) or the resource type could not be retrieved, although the DOI resolves (<http://doi.org/10.1186/s12885-015-1795-7>).

## Cost data by publisher and journal{.unnumbered}

```{r}
apc_publisher <- my_apc
apc_publisher$publisher <- factor (apc_publisher$publisher, levels = c(rownames(data.frame(rev(sort(table(apc_publisher$publisher)))))))

levels(apc_publisher$publisher)[11:length(levels(apc_publisher$publisher))] <- "other"
```

We used the DOI to automatically fetch publisher and journal names for each article from the CrossRef REST API. Table 4 shows the top ten publishers in terms of payments made that represent `r apc_publisher %>% filter(!publisher == "other") %>% nrow() / my_apc %>% nrow() * 100` % of the spending for publication fees. In total, payments were made to  `r my_apc %>% group_by(publisher) %>% summarize(n()) %>% nrow()` publishing houses. In comparison with data from the UK, full open access publishers have a greater share on total spending. @Pinfield_2015, for instance, reported remarkably lower numbers for the open access publishers MPDI AG, Copernicus GmbH, and Hindawi Publishing.

```{r results='asis'}
tt <- apc_publisher %>%
  group_by(publisher) %>%
  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), Median = median(euro), Min = min(euro), Max = max(euro))

colnames(tt) <- c("Publisher", "Articles funded", "Total", "Mean", "Median", "Minimum", "Maximum")
tt <- format(tt,big.mark=',')
pander::pandoc.table(tt, split.cells = 7, caption = 'Publication fees paid per publisher (in €)', justify = 'lrrrrrr')
```

Most of the funding of publication fees in Germany went to the publisher Springer Science + Business Media, especially profiting from the merge with the former full open access publisher BioMed Central. In contrast, other established publishing houses such as Elsevier and Wiley-Blackwell rank lower, presumably because they mostly publish hybrid journals, which were not well represented in our data-set at the time of the study. Table 4 also illustrates the variation across and within publishers, which confirms earlier findings [@Pinfield_2015].

```{r results='asis'}
apc_journal_full_title <- my_apc
apc_journal_full_title$journal_full_title <- factor (apc_journal_full_title$journal_full_title, levels = c(rownames(data.frame(rev(sort(table(apc_journal_full_title$journal_full_title)))))))

levels(apc_journal_full_title$journal_full_title)[11:length(levels(apc_journal_full_title$journal_full_title))] <- "other"

tt <- apc_journal_full_title %>%
  group_by(journal_full_title) %>%
  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), Median = median(euro), Min = min(euro), Max = max(euro))

colnames(tt) <- c("Journal", "Articles funded", "Total", "Mean", "Median", "Minimum", "Maximum")
tt <- format(tt,big.mark=',')
pander::pandoc.table(tt, split.cells = 7, caption = 'Publication fees paid per journal (in €)', justify = 'lrrrrrr')
```

Prices also varied within single journals. Based on the number of articles paid for, Table 5 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals. Payments to these ten journals represent `r apc_journal_full_title %>% filter(!journal_full_title == "other") %>% nrow() / my_apc %>% nrow() * 100` % of all payments. In the case of Atmospheric Chemistry and Physics Discussions, the price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.

The data-set finally confirms the leading role of "mega-journals" in open access publishing, including the multidisciplinary PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline. In general, an estimated 14 out of more than 10,000 journals registered in DOAJ in 2015 accounted for up to 15–20 % of all articles published in full open access journals [@Bj_rk_2015].

# Discussion{.unnumbered}

The Open APC initiative extends existing methods to disclose spending on open access publication fees. Our workflow benefits from openly available tools and the social coding platforms GitHub, both of which are well established and suited to increase transparency in research [@Peng_2011; @Ram_2013]. For 99 % of the articles, CrossRef provided bibliographic information, which substantially contributed to a uniform data-set about formerly fragmented payments made for open access articles.

Although CrossRef disambiguates journal titles and publisher names and is therefore an authority-controlled source for open access journal literature, derivations from CrossRef metadata curation as well as the context of aggregation must be made clear. In particular, problems persist on how to deal with name changes and ongoing mergers. For example, the publisher Public Library of Science (PLOS) has changed its acronym from "PLoS" to "PLOS", which CrossRef metadata reflects from 2015 onward. We therefore normalized all PLOS journal titles in order to secure unique reference to these journals. Another publisher affected is "The Optical Society," formerly "Optical Society of America". Because the ownership of publishing houses can become combined, dealing with mergers is also essential to make cost data comparable. Jisc data, for instance, differentiate between the full open access publisher BioMed Central and the traditional publisher Springer, concluding that "traditional publishing houses" lean on the hybrid model [@Pinfield_2015]. This stands in stark contrast to our approach, in which CrossRef metadata reflects the merger of BioMed Central and Springer, resulting in Springer Science + Business Media to be the best represented publisher for articles in full open access journals in the Open APC data-set. Another approach for ensuring unique reference, but that we have not evaluated yet, is to use CrossRef's identifiers for journals and publishers instead.[^19]

Because of the dynamic landscape of academic publishing and its representation in CrossRef's data curation efforts, it is important to consider the time-frame of metadata aggregation. In our case, we re-used metadata shortly after the data submission. However, for some cost analysis -- for instance to prepare negotiations with publishers on future schemes to fund open access journal articles -- it could be more feasible to re-normalize the complete data. The rossRef API provides incremental metadata updates that can be used to assess the current potential of future funding. While licensing information is incompletely covered in the CrossRef index so far, and therefore not analyzed in our study, the growing importance of facilitating text mining may result in more and more publishers sharing this information with CrossRef in the future.

Participation is voluntary. Therefore, not all institutions in Germany that provide central funding of publication fees contribute cost data to this initiative. In a qualitative survey that also asked why German institutions are reluctant to share their cost data through the Open APC initiative one institution feared that increase in transparency would allow publishers to adjust prices in their favor. Others pointed out that the workload to produce such a data-set could be too extensive [@deppe_2015].

While there may still be institutions that have no overview of their APC spending, we would like to emphasize that reporting data that is already available within an institution to the Open APC initiative should not lead to much additional work. The central incentive of this initiative is to make it as easy as possible to submit data to it. Being able to combine that data into one standardized data-set increases transparency and comparability, and gives institutions a better understanding of the overall development of open access publishing. In particular, we cannot see the harm of increased transparency; in fact, not knowing how much is spent is undoubtedly a disadvantage in dealing with publishers.

Extending the data template to include information about funders or whether special agreements with publishers applied as suggested by @Pinfield_2015, would even increase the efforts needed to participate. However, with the growing demand for action in areas like the large-scale transition of toll-access journals to open access [@Schimmer_Geschuhn_Vogler], an updated data template could help institutions to better comply with these policy developments in the future. From our experience, another barrier to participate is the lack of skills in version control: the data submission itself is not always made directly by the institutions. Instead, they sent files to Bielefeld University Library with the request to make them available on GitHub on their behalf.

Future work needs to focus on analyzing the cost data. Of particular interest are questions concerning the coverage of central funding schemes in comparison with open access publication output in general, and the use of other means to cover publication fees in particular. The Open APC initiative does not cover personal budgets or make price reductions explicit, but studies suggest that there is a possible gray area [@BJ_RK_2012; @Bj_rk_2015; @Lawson_waiver_2015]. Existing study designs could be re-applied to examine the relationship between price on the one hand, and indexing coverage, journal prestige or management costs on the other [@Walters_2011; @Bj_rk_scientometrics_2015, @Pinfield_2015]. This, in turn, helps to address the central question of future business models in scholarly publishing from an international perspective [@Pinfield_2015].

<!-- # Conclusion{.unnumbered}

The purpose of the Open APC initiative is to gain better insight on spending for open access. The main operational idea—to make data submission easy by building upon existing data and by using popular open tools and data bases—has proven to be a feasible approach and to scale well. Starting with a few institutions, the integration of further data contributors has not been a problem. While we use article metadata reported by publishers to CrossRef, we think it's best to collect information on how much has been spent at the sources: the institutions paying for APCs. We look forward to more data contributors to the initiative and to adapt this scheme to new analyses, additional data, or international sources. The Open APC initiative can serve as an example of the value that open data, open source software, and the knowledge and skills of tech-savvy librarians add to modern tasks of libraries.-->

# Acknowledgment{.unnumbered}

We thank Andrea Hacker and Ada-Charlotte Regelmann for valuable comments on the first draft of this paper. We also thank Christoph Broschinski, Vitali Peil, and Dirk Pieper, the members of the DINI working group "Electronic publishing", and all data contributors[^20] of the Open APC initiative.

# References{.unnumbered}

[^1]: Comprehensive list of business and revenue models of open access journals: <http://oad.simmons.edu/oadwiki/OA_journal_business_models>

[^2]: <https://github.com/openapc/openapc-de>

[^2a]: Project summary available at <https://github.com/OpenAPC/openapc-de>. The initiative is now supported by the INTACT project: http://www.intact-project.org/.

[^2b]: However, it should be noted that the administrative costs of processing invoices for a large number of articles are not measured in this project.

[^3]: Guidelines for the funding program can be found here: <http://www.dfg.de/formulare/12_20/>

[^3a]: We are sure that some universities with DFG-funded publication funds pay for APCs higher than 2,000 EUR. They have to make sure they don't use the funder's money for these articles, and this maybe explains why some institutions are reluctant to report on these APCs. Scatter plots of all APC payments show that there are only very few APCs over 2,000 EUR.

[^4b]: These might include article processing, impact, rejection rates, management and investment, and profit margins. See @Van_Noorden_2013 for a general discussion and @Gumpenberger_2012 and @Bj_rk_scientometrics_2015 for discussions of journal impact and quality.

[^5]: CrossRed: Text and Data Mining for Researchers: <http://tdmsupport.crossref.org/researchers/>

[^6]: rOpenSci: <https://ropensci.org/>

[^7]: Europe PubMed Central RESTful Web Service: <https://europepmc.org/RestfulWebService>

[^8]: Thomson Reuters Article Match Retrieval Service: <http://wokinfo.com/directlinks/amrfaq/>

[^9]: <https://doaj.org/faq#metadata>

[^10]: <http://rmarkdown.rstudio.com/>

[^10c]: <http://www.dfg.de/en/dfg_profile/alliance/index.html>

[^11]: <http://dini.de/english/ag0/e-pub0/>

[^12]: <http://openapc.github.io>

[^13]: <https://jekyllrb.com/>

[^13a]: <http://opendatacommons.org/licenses/odbl/1-0/>

[^14]: <https://lists.uni-bielefeld.de/mailman2/cgi/unibi/listinfo/open-apc>

[^15]: <https://github.com/OpenAPC/openapc-de/wiki>

[^16]: <https://github.com/OpenAPC/openapc-de/issues>

[^16a]: The data is openly available on GitHub. The following analysis is based on version 2.1.13 of the dataset, available at <https://github.com/OpenAPC/openapc-de/tree/v2.1.13>.

[^17]: <http://contributor-covenant.org/>

[^18]: <https://github.com/CrossRef/rest-api-doc/blob/master/rest_api.md>

[^19]: We would like to thank Martin Fenner for pointing this out to us. See also the CrossRef API documentation <https://github.com/CrossRef/rest-api-doc/blob/master/rest_api.md>

[^20]: <https://github.com/OpenAPC/openapc-de#contributors>

[^22]: <http://dx.doi.org/10.4119/UNIBI/UB.2014.18>

[^23]: <https://github.com/OpenAPC/openapc-de/wiki/Handreichung-Dateneingabe>

### 2016-04-07 article.Rmd
minor typo

@@
 Yet, how and to what extent these activities are effective in terms of the number of supported research articles and associated costs remains under debate.
-This paper describes the Open APC initiative[^2], in which German universities and research organization share spendings on open access publication fees, and how it is currently implemented.
+This paper describes the Open APC initiative[^2], in which German universities and research organization share spending on open access publication fees, and how it is currently implemented.
 More specifically, it addresses three problem areas when studying the economical effects of open access publishing: fragmentation of open access funding, variable pricing schemes and the comparison across research institutions.

### 2016-05-19 R/aggregation.r
re-aggregate open apc dataset and store as tabular json

@@ -1,0 +1,41 @@
+## parse crossref
+
+## load apc data from release 2.4.2
+apc <- read.csv("https://raw.githubusercontent.com/OpenAPC/openapc-de/74d4d442a674f8af9dda522faebba25e1f5656cc/data/apc_de.csv", 
+                header = TRUE, sep = ",")
+
+### preprocess
+apc$doi <- tolower(apc$doi)
+apc$doi <- gsub("dx.doi.org/", "", apc$doi)
+
+### 1. step exclude NA
+apc_doi <- apc[!is.na(apc$doi),]
+
+### 2. fetch metadata from CrossRef with ropensci::rcrossref client and parser for the tdm xml
+source("R/cr_parse.r")
+apc_cr <- plyr::ldply(apc_doi$doi, plyr::failwith(f = cr_parse))
+#### backup
+jsonlite::stream_out(apc_cr, file("data/cr-apc.json"))
+
+### 3. repeat for dois not found because of possible time outs
+doi_lack <- apc_doi[!tolower(apc_doi$doi) %in% tolower(apc_cr$doi),]
+doi_miss <- plyr::ldply(doi_lack$doi, plyr::failwith(f = cr_parse))
+
+apc_cr <- rbind(apc_cr, doi_miss)
+### 4. merge with cost info
+require(dplyr)
+apc_cr$doi <- tolower(apc_cr$doi)
+apc_euro <- dplyr::select(apc, institution, period, euro, is_hybrid, doi)
+apc_cr_euro <- dplyr::left_join(apc_cr, apc_euro, by = "doi")
+variables_needed <- c("doi", "euro", "institution", "publisher", "journal_full_title", "period", "license_ref", "is_hybrid")
+apc_short <- dplyr::select(apc, one_of(variables_needed)) %>% 
+  filter(!doi %in% apc_cr_euro$doi)
+apc_cr_all <- dplyr::select(apc_cr_euro, one_of(variables_needed)) %>% rbind(., apc_short)
+sum(apc_cr_all$euro)
+jsonlite::stream_out(apc_cr_all, file("data/cr-apc-all.json"))
+
+### 5. get agencies for non-crossref dois
+dois_ac <- doi_lack[!doi_lack$doi %in% doi_miss$doi,]
+# backup 
+jsonlite::stream_out(dois_ac, file("data/missing_dois.json"))
+                        

### 2016-05-19 article.Rmd
starting to imporve alignment of  research question and methods with the results we reported in the first draft

--- 2016-04-07 article.Rmd
@@
 
-In this paper, we present the technical workflow of the Open APC initiative.
-We describe how the data-set is curated and which tools are used in order to produce, disseminate, and preserve the Open APC spending data.
-Presenting our results, we will particularly discuss how and to what extent the use of the CrossRef index accommodates the demand of disambiguated bibliographic information about journals and publishers when reporting about funding of open access journal articles.
-CrossRef, a Digital Object Identifiers (DOI) registration agency for scholarly literature, associate these persistent links with metadata CrossRef members such as publishers and research societies contribute.<!--  Our analysis of the data contributed so far illustrates possible ways of using this data-set.
-We are able to show the importance of reusable open data and encourage other possible data contributors to also publish their payment information through this initiative.
-The goal is to strengthen the insight on open access funding, enabling institutions to better plan for a transition to open access.
--->
+Drawing on self-reported cost data, we asked to what extend German universities and research institutions covered publication fees for open access journal articles over the last years.
+More specifically, we addressed three questions that arose while studying fee-based open access publishing:
+
+- how to gather bibliographic metadata about sponsored open access journal articles,
+- how is institutional publication fee spending distributed over publishers and journal titles,
+- what is the share of sponsored articles in open access journals versus hybrid journals.
 
@@
 
-The major goal of the Open APC initiative is to gain insights into expenses for publication fees by collecting data directly from the institutions paying on behalf of the authors they support.
-These institutions can report best on the most important part: How much they have spent for each article.
-For this aim, the Open APC initiative applies open data methods, designed around the idea of sharing the collected data as permissive as possible.
-At its core is a data pipeline to collect and to process the contributed data.
-In this section, we discuss the general idea of this pipeline and describe its stages: data submission (by spending institutions), merging contributions, re-using data, preserving data, and engaging with our communities to increase participation.
+The study examined cost information provided by the Open APC initiative.
+Under this umbrella, German universities and research institutions shared details of how much open access journals charged authors as open data.
+We analysed the data-set that was released on 13 May 2016 and is stored as version v2.4.2 in the corresponding GitHub repository.
+At the time of our study, the data-set reported the amount institutions paid for each article including value added tax, and the year an article was invoiced.
+Bibliographic metadata consisted of the Digital Object Identifier (DOI) to uniquely identify an article, the publisher, and information about the journal titles.
+The Open APC initiative automatically retrieved this publisher and journal information from CrossRef, a DOI minting agency for scholarly literature.
+Furthermore, the data-set differentiated whether an open access article was published in a subscription based journal ("hybrid journal").
 
-## General idea{.unnumbered}
+For gathering normalized publisher and journal titles, we re-aggregated the bibliographic metadata from CrossRef using the DOI on 19 May 2016 because of two reasons: First, CrossRef provides incremental metadata updates to reflect the dynamic landscape of academic publishing including ongoing mergers of publishing houses or name changes.
+A prominent example in this regard is the merger of the two large publishing houses Springer Business + Media and Nature Publishing that now operate under "Springer Nature".
+In contrast to this, the Open APC initiative normalized publisher and journal titles based on CrossRef shortly after the institution had provided its cost data.
+Therefore, re-aggregating the data-set ensures unique reference to currently existing publishers and journals at the expense of information about to which publisher a journal belonged while a sponsored article was published.
 
-Our approach drew on general open data guidelines that state to "keep things simple" and to "engage early and engage often" with data providers and potential users [@OpenDataHandbook].
-We therefore chose a simple data scheme, built our pipeline around the popular social coding platform GitHub, and re-used information from bibliographic indexes.
+Secondly, CrossRef is not the only registration agency for DOI's, but also DataCite and Medra mint DOIs for scholarly work.
+While obtaining bibliographic information from CrossRef, we also examined the number of publications for which each of the DOI registration agency provided metadata during our study.
+As a client, we used the R package rcrossref [@rcrossref] developed and maintained by the rOpenSci initiative.[^6] With the function `cr_cn`, we requested the linking resource `application/vnd.crossref.unixsd+xml`, which main function is to support text mining activities.[^5] The advantage of this XML format was that it distinguishes full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals.
+It also contained license information and disambiguated publisher information, thus avoiding confusion about licensing and naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014].
+In cases where no bibliographic information could be obtained, we used the Open APC values.
 
-We followed good practice of other open data projects to guarantee that the data is as open as possible.
-From the beginning, we emphasized that this also includes information on submissions -- date of submission, contributors, etc.
--- and open file formats.
-Because institutions should be allowed to self-report data and updates at different times, tracking initial submissions and data-set updates is particularly important.
-
-All data and documentation reside, therefore, on GitHub.
-Git, a distributed version control system, powers GitHub and allows people to collaborate on software projects.
-Git keeps a log of changes made in the source code and manages to synchronize local copies of the very same software repository.
-Because of its distributed and social characteristics, Git in general and GitHub in particular are suitable for researchers to mutually curate and share research artifacts including data-sets, analyses, or visualizations [@Ram_2013].
-
-In addition to GitHub, bibliographic indexes are an essential part of our data pipeline.
-We use CrossRef to normalize bibliographic metadata.
-This ensures automatic authority control for journal and publisher titles, which are the most appropriate levels of aggregations when analyzing expenditures for open access articles [@Pinfield_2015].
-We also check automatically the indexing status of each article in Europe PubMed Central, a large bibliographic index for life-science literature, the multidisciplinary database Web of Science, and the open access source Directory of Open Access Journals (DOAJ).
-
-## Data Submission{.unnumbered}
-
-Participants use a template to self-report data (see Table 1).
-Data structured in this way enables the fetch of additional data from external sources, and ensures that the merge into a single data-set succeeds.
-By collecting as little data as possible from the institutions, this approach allows for changes and the addition of new fields if required.
-
-|Variable               |Description                              |Required  |
-|:----------------------|:----------------------------------------|:-------- |
-|`institution`          |Top-level organisation e.g. MPG          | mandatory|
-|`period` 		          |Year of APC payment (YYYY)               | mandatory|
-|`euro` 		            |The amount paid in EURO (incl. VAT)      | mandatory|
-|`doi`                  |Digital Object Identifier                | mandatory|
-|`is_hybrid`            |Published in a access journal?           | mandatory|
-|`publisher`          	|Name of publication house                | optional |
-|`journal_full_title`   |Full name of periodical                  | optional |
-|`issn` 		            |International Standard Serial Number.    | optional |
-
-Table: Mandatory and optional data elements for disclosing funding of open access publication fees
-
-The data scheme reflects how the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc [@Lawson_data_2015] organized data about expenses for open access publication fees.
-The project requests items that are already present at the research institutions for internal reporting purposes [@Pinfield_2015].
-Guidance[^23] available to the participants furthermore refers to the principles of *tidy data* [@tidy-data].
-This responds to issues encountered in the United Kingdom, when the inexperienced use of spreadsheet software like Excel lead to misaligned tables [@woodward_2014].
-The principle of tidy data intends to reduce data cleaning efforts before statistical analyses.
-The three principles of tidy data state that, firstly, each variable must form a column, secondly, each observation must form a row, and, finally, each type of observational unit must form a table.
-To prevent messy data, it is also crucial that unknown values are not left empty.
-In our case, we use the R convention `NA` for handling those values.
-
-Data contributors collect the data in different systems.
-After preparing the data, it must be exported in the csv standard, a plain text file format for tabular data supported by most spreadsheet software that can be easily logged with version control systems like Git.
-
-To illustrate a sample data contribution displaying mandatory information in the csv format:
-
-```
-"Hannover U",2013,1241.02,"10.1371/journal.pone.0063501",FALSE
-```
-
-Contributors store the csv file into a folder that also contains a README file.
-Written in plain text or Markdown, the README contains information about the data-set and the contributing institution.
-The submission itself is facilitated through GitHub's pull request mechanism: contributors fork the initiative's data, add their contributions, upload their modified copy to their repository, and request importing their data into the initiative's repository by alerting the maintainers through sending a "pull request".
-
-## Data curation{.unnumbered}
-
-<!--Bielefeld University Library acts as curator and merges the contributed spreadsheets into one general data-set.--> By using open software, we, as curators, retrieve additional data elements based on the article's DOI.
-This avoids manual cleaning efforts the other initiatives were faced with.
-Because the DOI is at the center of the data curation, data normalization starts with a check of the DOI columns for possible duplicates and white space.
-CrossRef is then queried for article metadata matching the particular DOI.
-
-CrossRef provides several APIs to search for bibliographic information it indexes during DOI registration.
-In our case, we query the CrossRef REST API[^18] by using content negotiation.
-As resource type, we request the format `application/vnd.crossref.unixsd+xml`, which main function is to support text mining activities.[^5] The advantage of this XML format is that it distinguishes full and abbreviated journal titles as well as the media types of ISSNs, the International Standard Serial Number used to identify journals.
-It also contains license information and disambiguated publisher information, thus avoiding confusion about licensing and naming of publisher houses [@woodward_2014].
-
-As a client, we use the R package `rcrossref` [@rcrossref] developed and maintained by the rOpenSci initiative.[^6] With the function `cr_cn`, the client supports all linking types.
-Table 2 summarizes the data elements we retrieve:
-
-|Source     	|Data element  |Description                     |
-|:--------------|:-------------------|:------------------------------------|
-|CrossRef   	|`publisher` |Title of Publisher             |
-|CrossRef   	|`journal_full_title` |Full Title of Journal  |
-|CrossRef   	|`issn` |International Standard Serial Numbers (collapsed) |
-|CrossRef   	|`issn_print` |ISSN print |
-|CrossRef   	|`issn_electronic`  |ISSN electronic        |
-|CrossRef   	|`license_ref`  |License of the article     |
-|CrossRef   	|`indexed_in_crossref`  |Indexed in CrossRef? (logical)    |
-|EuropePMC    	|`pmid`  |PubMed ID                 |
-|EuropePMC    	|`pmcid` |PubMed Central ID         |
-|Web of Science |`ut` |Web of Science record ID             |
-|DOAJ           |`doaj` |Is the journal indexed in the DOAJ? (logical)    |
-
-Table: Sources used to automatically enrich the Open APC data-set
-
-In addition to CrossRef, the indexing status for each article in Europe PubMed Central and the Web of Science is checked.
-Europe PubMed Central, one of the largest database for life science literature, offers a public RESTful web services to access more than 24 million records and 870,000 deposited open access articles [@epmc_2014].
-Information from the Web of Science, which provides indexing of open access journal literature [@Walters_2011], is retrieved through the Thomson Reuters Article Match Retrieval Service[^8].
-This web interface is available to Bielefeld University Library as part of its Web of Science subscription.
-Finally, an automatic match with the DOAJ, a comprehensive and openly available registry of open access journals, is performed.
-
-After disambiguating and enriching the data, we append the rows into a single csv file that records all expenses.
-The data additions are logged with Git and pushed to the source code repository on GitHub.
-
-## Re-use{.unnumbered}
-
-We use the main README file of our data repository as general guide to the Open APC data collection and to present sample analyses.
-The README itself is written in R Markdown, an authoring format that allows for the combination of the R code, results, and text within one document.[^10] Compiled with knitr [@knitr], the README reflects new data submissions after every build.
-Sample results include tables summarizing the number of supported articles and costs per institution as well as figures that illustrate the distribution of publication fees.
-
-A blog hosted on GitHub presents all new data contributions.[^12] Since it is technically based on Jekyll,[^13] a static site generator, we can use the same development cycle we have for producing the README files.
-We write blog posts, which include information about the providing institution and the contributed data-set, in R Markdown, build the posts with `knitr` and, then, log and deploy the files to GitHub with Jekyll, making it possible to reproduce the analytical steps.
-
-## Preservation{.unnumbered}
-
-The data-set itself is licensed under the Open Database License (ODbL) v1.0,[^13a] permitting re-use under conditions of attribution and share-alike.
-To preserve the collected data, a GitLab installation hosted by Bielefeld University and administrated by Bielefeld University Library, mirrors a copy of the GitHub repository and is accessible via a DOI.[^22]
-
-To provide reference points to particular data contributions, unique version names refer to data submissions or other changes in the source code, which will be then distributed as releases through GitHub.
-This approach allows others to not only to refer, but also to re-use or self-archive particular snapshots of the data.
-
-## Engagement{.unnumbered}
-
-Besides technical measures, social aspects are crucial to make an open data initiative successful.
-For the Open APC initiative, we focus on engaging librarians and collaborating within existing networks.
-In Germany, the working group "Elektronisches Publizieren" of the Deutsche Initiative für Netzwerkinformation (DINI) coordinates these efforts.[^11] The Open APC initiative aims at increasing participation also through the use of social media, workshops, and regular community calls.
-
-A GitHub wiki[^15] shares guidance and interim reports.
-If participants want to report bugs or propose new functionality, GitHub's "issues" mechanism involves users while keeping track of the discussion.
-To ensure a constructive environment for all participants, we refer to the code of conduct from Contributor Covenant.[^17]
-
-<!--From our experience, the process described in this chapter has played an important role in getting many institutions to participate in voluntarily sharing information on their expenses for APCs.
-The data scheme has been revised at the end of 2015, and transformation to a new scheme has happened without any complications.
-Further adaption, e.g.
-adding subject classification if a suitable source can be identified, is possible.
-We encourage institutions and researchers to further work on the data pipeline while preserving its openness and the low entry barrier for new data contributors.-->
+After summarizing publication fee spending per institution, publisher and journal, we compared the share of hybrid journal articles in the Open APC data-set with those of Wellcome Trust, FWF and Jisc.
+For this aim, we systematically reviewed the proportions for each year from the accompanying literature.
 

### 2016-05-25 article.Rmd
research questions and methods re-tooled, calculations need to be addressed

@@
 ---
-title: 'Disclosing funding sources for open access publication fees: the Open APC initiative'
+title: 'Disclosing funding sources for open access publication fees: the Open APC
+  initiative'
 author:
-- Najko Jahn, Bielefeld University Library, Bielefeld University
-- <a href="http://orcid.org/0000-0002-5111-2788">Marco Tullney</a>, Technische Informationsbibliothek (TIB) - German National Library of Science and Technology, Hannover, Germany
-date: '7 April 2016'
+- Najko Jahn, Bielefeld University Library, Bielefeld University (TIB) - German National
+  Library of Science and Technology, Hannover, Germany
+- <a href="http://orcid.org/0000-0002-5111-2788">Marco Tullney</a>, Technische Informationsbibliothek
+date: "7 April 2016"
 output:
@@
 csl: frontiers.csl
-abstract: Publication fees in open access publishing hold a prominent place on the agenda of researchers, policy-makers, and academic publishers.
-This paper contributes to the evolving empirical basis on open access funding.
-It describes the Open APC initiative, in which German universities and research organizations share their expenditures for publication fees.
-As method, the initiative uses existing open data tools to aggregate and disseminate institutional spending on open access publication fees.
-In total, 29 German research organizations self-reported funding of 6,279 open access journal articles, which amounted to 8,039,339 €.
-The average payment for each article was 1,280 €, and the median payment 1,209 €.
-Our data-set comprises only 53 articles in hybrid journals.
-With an indexing coverage of 99 %, the findings reveal that the DOI agency CrossRef provides both comprehensive bibliographic coverage of the funded open access journal literature and disambiguated names of journal titles and publishing houses.
-We show that authority control of these bibliographic information is particularly relevant for the comparative study of the economical effects of open access publishing.
+abstract: Publication fees in open access publishing hold a prominent place on the
+  agenda of researchers, policy-makers, and academic publishers. This paper contributes
+  to the evolving empirical basis on open access funding. It describes the Open APC
+  initiative, in which German universities and research organizations share their
+  expenditures for publication fees. As method, the initiative uses existing open
+  data tools to aggregate and disseminate institutional spending on open access publication
+  fees. In total, 29 German research organizations self-reported funding of 6,279
+  open access journal articles, which amounted to 8,039,339 €. The average payment
+  for each article was 1,280 €, and the median payment 1,209 €. Our data-set comprises
+  only 53 articles in hybrid journals. With an indexing coverage of 99 %, the findings
+  reveal that the DOI agency CrossRef provides both comprehensive bibliographic coverage
+  of the funded open access journal literature and disambiguated names of journal
+  titles and publishing houses. We show that authority control of these bibliographic
+  information is particularly relevant for the comparative study of the economical
+  effects of open access publishing.
 ---
@@
 
-Drawing on self-reported cost data, we asked to what extend German universities and research institutions covered publication fees for open access journal articles over the last years.
-More specifically, we addressed three questions that arose while studying fee-based open access publishing:
-
-- how to gather bibliographic metadata about sponsored open access journal articles,
-- how is institutional publication fee spending distributed over publishers and journal titles,
-- what is the share of sponsored articles in open access journals versus hybrid journals.
+The aim of this study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
+Drawing on self-reported data that German institutions openly shared within the Open APC initiative, we focused our analysis on the amount of institutional publication fee spending, and compared these findings with that of related efforts in Austria and the UK.
+We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.
 
@@
 
-The study examined cost information provided by the Open APC initiative.
-Under this umbrella, German universities and research institutions shared details of how much open access journals charged authors as open data.
-We analysed the data-set that was released on 13 May 2016 and is stored as version v2.4.2 in the corresponding GitHub repository.
-At the time of our study, the data-set reported the amount institutions paid for each article including value added tax, and the year an article was invoiced.
-Bibliographic metadata consisted of the Digital Object Identifier (DOI) to uniquely identify an article, the publisher, and information about the journal titles.
-The Open APC initiative automatically retrieved this publisher and journal information from CrossRef, a DOI minting agency for scholarly literature.
-Furthermore, the data-set differentiated whether an open access article was published in a subscription based journal ("hybrid journal").
+We used cost data released by the Open APC initiative on May 13, 2016, to analyse expenditures on publication fees by German universities and research organisations.
+In addition to administrative data about the amount paid per article including value added tax and the year of invoicing, the Open APC initiative has begun to share information about whether an article was published in a fully or hybrid open access journal.
+Bibliographic metadata recorded in the data-set consisted of the Digital Object Identifier (DOI) to uniquely identify an article, and information about publisher and journal titles that were automatically retrieved from CrossRef when available.
 
-For gathering normalized publisher and journal titles, we re-aggregated the bibliographic metadata from CrossRef using the DOI on 19 May 2016 because of two reasons: First, CrossRef provides incremental metadata updates to reflect the dynamic landscape of academic publishing including ongoing mergers of publishing houses or name changes.
-A prominent example in this regard is the merger of the two large publishing houses Springer Business + Media and Nature Publishing that now operate under "Springer Nature".
-In contrast to this, the Open APC initiative normalized publisher and journal titles based on CrossRef shortly after the institution had provided its cost data.
-Therefore, re-aggregating the data-set ensures unique reference to currently existing publishers and journals at the expense of information about to which publisher a journal belonged while a sponsored article was published.
+Data collection in this study involved obtaining cost data from related open data initiatives and retrieving bibliographic metadata updates.
+To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed  the openly available spreadsheets provided by the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
+All data-sets were openly available through a figshare collection.
+We obtained the size  of funding for all publications and fully open access journal articles.
+The Euro foreign exchange reference rates from the European Central bank were consulted to convert costs reported in GBP to Euro.
 
-Secondly, CrossRef is not the only registration agency for DOI's, but also DataCite and Medra mint DOIs for scholarly work.
-While obtaining bibliographic information from CrossRef, we also examined the number of publications for which each of the DOI registration agency provided metadata during our study.
-As a client, we used the R package rcrossref [@rcrossref] developed and maintained by the rOpenSci initiative.[^6] With the function `cr_cn`, we requested the linking resource `application/vnd.crossref.unixsd+xml`, which main function is to support text mining activities.[^5] The advantage of this XML format was that it distinguishes full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals.
-It also contained license information and disambiguated publisher information, thus avoiding confusion about licensing and naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014].
+To provide a current view on open access publishing, DOIs reported in the Open APC data-set served us as a basis to re-collect bibliographic metadata from CrossRef on May 19, 2016.
+Since the beginning of data normalisation using CrossRef by the Open APC initiative in 2014, CrossRef has provided incremental metadata updates in which ongoing mergers of publishing houses or name changes are reflected.
+A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing announced on May 6, 2015, that now operate under Springer Nature.
+In contrast to this, the Open APC initiative gathered publisher and journal titles shortly after an institution provided its expenditures.
+
+As a client to access the CrossRef web service,  we used the R package rcrossref [@rcrossref] developed and maintained by the rOpenSci initiative.[^6] With the function `cr_cn`, we requested the linking resource `application/vnd.crossref.unixsd+xml` used  to support text mining activities.[^5] At the time of our study, this XML format distinguished full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals.
+It also contained disambiguated publisher information, thus avoiding confusion about naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014].
 In cases where no bibliographic information could be obtained, we used the Open APC values.
+Since CrossRef is not the only registration agency for DOI's, but also the agencies DataCite and Medra mint DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
 
-After summarizing publication fee spending per institution, publisher and journal, we compared the share of hybrid journal articles in the Open APC data-set with those of Wellcome Trust, FWF and Jisc.
-For this aim, we systematically reviewed the proportions for each year from the accompanying literature.
+- statistical calculations
 
@@
 # load data
-my_apc <- read.csv("data/apc_de.csv", header = TRUE, sep =",", stringsAsFactors = FALSE)
+my_apc <- jsonlite::stream_in(file("data/cr-apc-all.json"), verbose = FALSE)
 
@@
 library(dplyr)
-# remove Bielefelds Zero value
-my_apc <- filter(my_apc, euro > 0)
+## remove Bielefelds Zero value
+my_apc <- dplyr::filter(my_apc, euro > 0)
 # remove intech book chapters
-my_apc <-  filter(my_apc, !is.na(journal_full_title))
-# clean publisher titles, because metadata could not be fetched
-my_apc$publisher[grep("BioMed Central", my_apc$publisher, ignore.case = T)] <- "Springer Science + Business Media"
-my_apc$publisher[grep("Impact Journals, LLC", my_apc$publisher, ignore.case = T)] <- "Impact Journals LLC"
-my_apc$publisher[grep("e-Century Publishing Corporation", my_apc$publisher, ignore.case = T)] <- "e-Century Publ. Corp."
-my_apc$publisher[grep("Academic Journals", my_apc$publisher, ignore.case = T)] <- "Academic Journals"
+my_apc <-  dplyr::filter(my_apc, !is.na(journal_full_title))
+
 # fix plos name change in journal titles
 my_apc$journal_full_title <- gsub("PLoS", "PLOS", my_apc$journal_full_title)
-my_apc$journal_full_title <- gsub("PloS","PLOS", my_apc$journal_full_title)
-my_apc$journal_full_title <- gsub("PlOS","PLOS", my_apc$journal_full_title)
-my_apc$journal_full_title <- gsub("Sustainibility","Sustainability", my_apc$journal_full_title)
-my_apc$journal_full_title <- gsub("Surgeal","Surgical", my_apc$journal_full_title)
-my_apc$journal_full_title <- gsub("Oncotarget","OncoTarget", my_apc$journal_full_title)
-my_apc$journal_full_title <- gsub("BMC cancer","BMC Cancer", my_apc$journal_full_title)
+my_apc$journal_full_title <- gsub("PLOS ONE", "PLOS ONE", my_apc$journal_full_title)
+
+# remove Austrian FWF data
+my_apc <- dplyr::filter(my_apc, !institution == "FWF - Austrian Science Fund")
+# remove data from 2016
+my_apc <- dplyr::filter(my_apc, !period == "2016")
 
@@
 
-On February 19th 2016,[^16a] the Open APC initiative covered `r nrow(my_apc)` articles, whose publication fees were centrally paid by `r length(unique(my_apc$institution))` German universities and research institutions.
-The number of supported open access journal articles, which were reported to this initiative, grew over the years (see Figure 1).
+On May 13, 2016,[^16a] the Open APC initiative covered `r nrow(my_apc)` articles whose publication fees were centrally paid by `r length(unique(my_apc$institution))` German universities and research institutions until 2015.
+The number of supported open access journal articles grew over the years (see Figure 1).
 While one institution disclosed `r filter(my_apc, period == min(my_apc$period)) %>% nrow()` payments made in `r as.character(min(my_apc$period))`, the majority shared their expenditures from 2013 onwards.
 With `r max(apc_time$articles)` articles, the year `r apc_time[apc_time$articles == max(apc_time$articles),"period"]` was best represented in our data-set.
-Because of a time lag between payments made and reporting them to the Open APC initiative, only `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institution were able to partly contribute their cost data for 2015 at the time of this analysis.
+However, `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institutions contributed their cost data for 2015 at the time of this analysis, suggesting that there exits a time lag between payments made and reporting these spending to the Open APC initiative.
+
+```{r histm, fig.width=6, fig.height=2.5,fig.cap="Histogram APC"}
+require(ggplot2)
+## http://www.noamross.net/blog/2013/11/20/formatting-plots-for-pubs.html
+science_theme = theme(panel.grid.major = element_line(size = 0.5, color = "grey"), 
+    axis.line.x = element_line(size = 0.7, color = "black"),
+    axis.line.y = element_line(size = 0.7, color = "black"),
+    text = element_text(size = 14))
+
+ggplot(my_apc, aes(euro)) + 
+  geom_histogram(binwidth = 250, colour ="black", boundary = 0) + 
+  scale_x_continuous(limits = c(0,8000)) +
+  xlab("Publication fee in €") +
+  ylab("Articles") +
+  theme_bw() + science_theme
+
+```
 
@@
   summarize(articles = n())
+
+## http://www.noamross.net/blog/2013/11/20/formatting-plots-for-pubs.html
+science_theme = theme(panel.grid.major = element_line(size = 0.5, color = "grey"), 
+    axis.line.x = element_line(size = 0.7, color = "black"),
+    axis.line.y = element_line(size = 0.7, color = "black"),
+    text = element_text(size = 14))
+
 ggplot(apc_time, aes(factor(period), cumsum(articles), group = 1)) +
@@
   ylab("OA articles reported\n(cumulative sum)") +
-  scale_y_continuous(limits = c(0,8000))
+  scale_y_continuous(limits = c(0,8000)) +
+  theme_bw() + science_theme
+ggsave("figure_1.pdf", height=4, width=6)
+
 ```
@@
 ggplot(my_apc, aes(factor(period), euro)) +
-  geom_boxplot() +
+  geom_boxplot(outlier.size = 0.5) +
   xlab("Year") +
-  ylab("Publishing fee in EURO(log scale)") +
-  scale_y_continuous(limits = c(0,8000))
+  ylab("Publication fee paid in EURO") +
+  scale_y_continuous(limits = c(0,8000)) + 
+  theme_bw() + science_theme
+ggsave("figure_2.pdf", height=4, width=6)
+
 ```
@@
 They were reported for `r my_apc %>% filter(!is.na(doi)) %>% nrow()` out of `r nrow(my_apc)` articles.
-Of those, `r my_apc %>% filter(indexed_in_crossref == TRUE) %>% nrow()` were indexed in CrossRef, representing `r my_apc %>% filter(indexed_in_crossref == TRUE) %>% nrow() / my_apc %>% nrow() * 100`  % of all funded publications.
+Of those, `my_apc %>% filter(indexed_in_crossref == TRUE) %>% nrow()` were indexed in CrossRef, representing `my_apc %>% filter(indexed_in_crossref == TRUE) %>% nrow() / my_apc %>% nrow() * 100`  % of all funded publications.
 The reasons why articles identified by a DOI were not registered with CrossRef differed.
@@
 apc_publisher <- my_apc
-apc_publisher$publisher <- factor (apc_publisher$publisher, levels = c(rownames(data.frame(rev(sort(table(apc_publisher$publisher)))))))
-
-levels(apc_publisher$publisher)[11:length(levels(apc_publisher$publisher))] <- "other"
+publisher_ <- names(rev(sort(table(apc_publisher$publisher)))[1:10])
+apc_publisher$publisher <- ifelse(apc_publisher$publisher %in% publisher_, apc_publisher$publisher, "Other")
+# SORT FACTOR
+apc_publisher$publisher <- factor(apc_publisher$publisher, levels=c(publisher_, "Other")) 
 ```
@@
 We used the DOI to automatically fetch publisher and journal names for each article from the CrossRef REST API.
-Table 4 shows the top ten publishers in terms of payments made that represent `r apc_publisher %>% filter(!publisher == "other") %>% nrow() / my_apc %>% nrow() * 100` % of the spending for publication fees.
+Table 4 shows the top ten publishers in terms of payments made that represent `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the spending for publication fees.
 In total, payments were made to  `r my_apc %>% group_by(publisher) %>% summarize(n()) %>% nrow()` publishing houses.
@@
   group_by(publisher) %>%
-  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), Median = median(euro), Min = min(euro), Max = max(euro))
+  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), SD = sd(euro), Median = median(euro), Min = min(euro), Max = max(euro)) %>%
+  mutate(range_apc = paste(Min, "--", Max)) %>%
+  select(-Min, -Max)
 
-colnames(tt) <- c("Publisher", "Articles funded", "Total", "Mean", "Median", "Minimum", "Maximum")
+colnames(tt) <- c("Journal", "Articles funded", "Total", "Mean", "SD", "Median", "Range")
 tt <- format(tt,big.mark=',')
@@
 ```{r results='asis'}
-apc_journal_full_title <- my_apc
-apc_journal_full_title$journal_full_title <- factor (apc_journal_full_title$journal_full_title, levels = c(rownames(data.frame(rev(sort(table(apc_journal_full_title$journal_full_title)))))))
+apc_journal <- my_apc
 
-levels(apc_journal_full_title$journal_full_title)[11:length(levels(apc_journal_full_title$journal_full_title))] <- "other"
+journal_ <- names(rev(sort(table(apc_journal$journal_full_title)))[1:10])
+apc_journal$journal_full_title <- ifelse(apc_journal$journal_full_title %in% journal_, 
+                                         apc_journal$journal_full_title, "Other")
 
-tt <- apc_journal_full_title %>%
+# SORT FACTOR
+apc_journal$journal_full_title <- factor(apc_journal$journal_full_title, levels=c(journal_, "Other")) 
+tt <- apc_journal %>%
   group_by(journal_full_title) %>%
-  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), Median = median(euro), Min = min(euro), Max = max(euro))
+  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), SD = sd(euro), Median = median(euro), Min = min(euro), Max = max(euro)) %>%
+  mutate(range_apc = paste(Min, "--", Max)) %>%
+  select(-Min, -Max)
 
-colnames(tt) <- c("Journal", "Articles funded", "Total", "Mean", "Median", "Minimum", "Maximum")
+colnames(tt) <- c("Journal", "Articles funded", "Total", "Mean", "SD", "Median", "Range")
 tt <- format(tt,big.mark=',')
@@
 Based on the number of articles paid for, Table 5 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals.
-Payments to these ten journals represent `r apc_journal_full_title %>% filter(!journal_full_title == "other") %>% nrow() / my_apc %>% nrow() * 100` % of all payments.
+Payments to these ten journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of all payments.
 In the case of Atmospheric Chemistry and Physics Discussions, the price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.

### 2016-05-25 R/aggregation.r

@@ -1,3 +1,7 @@
+#############################################
+# use script from home directory of project # 
+#############################################
+
 ## parse crossref
 
 ## load apc data from release 2.4.2
@@ -11,13 +15,16 @@ apc$doi <- gsub("dx.doi.org/", "", apc$doi)
 ### 1. step exclude NA
 apc_doi <- apc[!is.na(apc$doi),]
 
-### 2. fetch metadata from CrossRef with ropensci::rcrossref client and parser for the tdm xml
+### 2. fetch metadata from CrossRef with ropensci::rcrossref client and parser 
+### for the tdm xml
 source("R/cr_parse.r")
 apc_cr <- plyr::ldply(apc_doi$doi, plyr::failwith(f = cr_parse))
 #### backup
 jsonlite::stream_out(apc_cr, file("data/cr-apc.json"))
 
-### 3. repeat for dois not found because of possible time outs
+### 3. repeat for dois not found because of time outs and white space we found
+apc_doi$doi <- trimws(apc_doi$doi)
+apc$doi <- trimws(apc$doi)
 doi_lack <- apc_doi[!tolower(apc_doi$doi) %in% tolower(apc_cr$doi),]
 doi_miss <- plyr::ldply(doi_lack$doi, plyr::failwith(f = cr_parse))
 
@@ -27,15 +34,32 @@ require(dplyr)
 apc_cr$doi <- tolower(apc_cr$doi)
 apc_euro <- dplyr::select(apc, institution, period, euro, is_hybrid, doi)
 apc_cr_euro <- dplyr::left_join(apc_cr, apc_euro, by = "doi")
-variables_needed <- c("doi", "euro", "institution", "publisher", "journal_full_title", "period", "license_ref", "is_hybrid")
+variables_needed <- c("doi", "euro", "institution", "publisher", 
+                      "journal_full_title", "period", "license_ref", "is_hybrid")
 apc_short <- dplyr::select(apc, one_of(variables_needed)) %>% 
   filter(!doi %in% apc_cr_euro$doi)
-apc_cr_all <- dplyr::select(apc_cr_euro, one_of(variables_needed)) %>% rbind(., apc_short)
-sum(apc_cr_all$euro)
+# no member and journal ids available because articles were not indexed by crossref
+apc_short$member_id <- NA
+apc_short$journal_id <- NA
[…]

### 2016-05-25 R/cr_parse.r

@@ -1,5 +1,5 @@
 cr_parse <- function(doi) {
-  
+
   # fetch XML through rcrossref
   tryCatch({
     doc <- rcrossref::cr_cn(doi, "crossref-tdm")

### 2016-05-25 article.Rmd
research question paragraph and methods completed

@@
 The aim of this study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-Drawing on self-reported data that German institutions openly shared within the Open APC initiative, we focused our analysis on the amount of institutional publication fee spending, and compared these findings with that of related efforts in Austria and the UK.
+Drawing on self-reported data that German institutions openly shared through the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and British initiatives.
 We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.
@@
 We used cost data released by the Open APC initiative on May 13, 2016, to analyse expenditures on publication fees by German universities and research organisations.
-In addition to administrative data about the amount paid per article including value added tax and the year of invoicing, the Open APC initiative has begun to share information about whether an article was published in a fully or hybrid open access journal.
-Bibliographic metadata recorded in the data-set consisted of the Digital Object Identifier (DOI) to uniquely identify an article, and information about publisher and journal titles that were automatically retrieved from CrossRef when available.
+In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully or hybrid open access journal as well as the recorded DOI.
 
-Data collection in this study involved obtaining cost data from related open data initiatives and retrieving bibliographic metadata updates.
-To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed  the openly available spreadsheets provided by the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
-All data-sets were openly available through a figshare collection.
-We obtained the size  of funding for all publications and fully open access journal articles.
-The Euro foreign exchange reference rates from the European Central bank were consulted to convert costs reported in GBP to Euro.
+Data collection in this study involved obtaining cost data from related open data initiatives.
+To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed  the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
+For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles.
+In the case of FWF, we gathered the cost information from accompanying reports.
+We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.
 
-To provide a current view on open access publishing, DOIs reported in the Open APC data-set served us as a basis to re-collect bibliographic metadata from CrossRef on May 19, 2016.
-Since the beginning of data normalisation using CrossRef by the Open APC initiative in 2014, CrossRef has provided incremental metadata updates in which ongoing mergers of publishing houses or name changes are reflected.
-A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing announced on May 6, 2015, that now operate under Springer Nature.
-In contrast to this, the Open APC initiative gathered publisher and journal titles shortly after an institution provided its expenditures.
+Instead of re-using publisher and journal information contained in the Open APC data-set, we fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOI's.
+Although the Open APC initative gathered metadata representing publishers and journals from CrossRef as well, this information was  retrieved at the time when the participating institutions submitted the cost information.
+However, CrossRef regularly updates metadata to represent ongoing mergers of publishing houses or name changes.
+A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that now operates as Springer Nature.
+To reflect these changes in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set.
 
-As a client to access the CrossRef web service,  we used the R package rcrossref [@rcrossref] developed and maintained by the rOpenSci initiative.[^6] With the function `cr_cn`, we requested the linking resource `application/vnd.crossref.unixsd+xml` used  to support text mining activities.[^5] At the time of our study, this XML format distinguished full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals.
+As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access CrossRef's REST API.
+We requested the XML-based format `application/vnd.crossref.unixsd+xml`, which distinguished full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals.
 It also contained disambiguated publisher information, thus avoiding confusion about naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014].
@@
 Since CrossRef is not the only registration agency for DOI's, but also the agencies DataCite and Medra mint DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
-
-- statistical calculations
 

### 2016-05-26 article.Rmd
why we not enrich the other data-sets with crossref

@@
 
-We used cost data released by the Open APC initiative on May 13, 2016, to analyse expenditures on publication fees by German universities and research organisations.
-In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully or hybrid open access journal as well as the recorded DOI.
-
-Data collection in this study involved obtaining cost data from related open data initiatives.
-To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed  the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
-For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles.
-In the case of FWF, we gathered the cost information from accompanying reports.
-We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.
+We analysed self-reported cost data released by the Open APC initiative on May 13, 2016, to assess institutional spending on open access publication fees in Germany.
+In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully or hybrid open access journal as well as the recorded DOI from the data-set.
 
 Instead of re-using publisher and journal information contained in the Open APC data-set, we fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOI's.
-Although the Open APC initative gathered metadata representing publishers and journals from CrossRef as well, this information was  retrieved at the time when the participating institutions submitted the cost information.
+Although the Open APC initiative gathered metadata representing publishers and journals from CrossRef as well, this information was  retrieved at the time when the participating institutions submitted the cost information.
 However, CrossRef regularly updates metadata to represent ongoing mergers of publishing houses or name changes.
@@
 As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access CrossRef's REST API.
-We requested the XML-based format `application/vnd.crossref.unixsd+xml`, which distinguished full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals.
+We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals, were distinguished.
 It also contained disambiguated publisher information, thus avoiding confusion about naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014].
@@
 Since CrossRef is not the only registration agency for DOI's, but also the agencies DataCite and Medra mint DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
+
+Data collection in this study also involved obtaining cost data from related open data initiatives.
+To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
+For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles.
+In the case of FWF, we gathered the cost information from accompanying reports.
+We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.
+Our comparison between the open data initiatives focussed on the last two years 2014 and 2015.Since Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the exchange rates of the full two-years as we could not determine the actual dates of spending.
+We excluded Jisc's spending in 2014 because full currency information in the data-set was not available.
+
+Data collection methods of the Open APC initiative and those of the others differed in some aspects.
+For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institution were required to report, publication identifier in the Wellcome Trust data might also be added through automated compliance checks.
+For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided only to compare German spending with that of other initiatives, but not its size and distribution over publisher and journal titles, as well as the indexing coverage in CrossRef.
+However, we made use of the literature to discuss our findings in relation to the other initiatives.
 

### 2016-05-29 article.Rmd
new draft discussion

@@
 
-The Open APC initiative extends existing methods to disclose spending on open access publication fees.
-Our workflow benefits from openly available tools and the social coding platforms GitHub, both of which are well established and suited to increase transparency in research [@Peng_2011; @Ram_2013].
-For 99 % of the articles, CrossRef provided bibliographic information, which substantially contributed to a uniform data-set about formerly fragmented payments made for open access articles.
+In Germany, institutional spending on open access publication fees has increased over the years.
+With a share of …, the majority of open access articles German institutions reported to the Open APC initiative were published in fully open access journals.
+This presumably reflects the DFG funding policy which excludes the support of articles published in hybrid open access journals.
+The DFG has been financially supporting the implementation of central publication funds at more than 30 German universities since 2009.
+However, reviewing self-reported cost data from funders or countries that also support hybrid open access journals or open access books revealed a smaller proportions of payments in favor of articles in fully open access journals.
+Since open access publication fee spending is generally fragmented, we cannot answer whether German researchers avoid opting for open access when publishing in hybrid journals or use other budgets to pay publication fees required to make their work open access through these kind of journals.
 
-Although CrossRef disambiguates journal titles and publisher names and is therefore an authority-controlled source for open access journal literature, derivations from CrossRef metadata curation as well as the context of aggregation must be made clear.
-In particular, problems persist on how to deal with name changes and ongoing mergers.
-For example, the publisher Public Library of Science (PLOS) has changed its acronym from "PLoS" to "PLOS", which CrossRef metadata reflects from 2015 onward.
-We therefore normalized all PLOS journal titles in order to secure unique reference to these journals.
-Another publisher affected is "The Optical Society," formerly "Optical Society of America".
-Because the ownership of publishing houses can become combined, dealing with mergers is also essential to make cost data comparable.
-Jisc data, for instance, differentiate between the full open access publisher BioMed Central and the traditional publisher Springer, concluding that "traditional publishing houses" lean on the hybrid model [@Pinfield_2015].
-This stands in stark contrast to our approach, in which CrossRef metadata reflects the merger of BioMed Central and Springer, resulting in Springer Science + Business Media to be the best represented publisher for articles in full open access journals in the Open APC data-set.
-Another approach for ensuring unique reference, but that we have not evaluated yet, is to use CrossRef's identifiers for journals and publishers instead.[^19]
+In our study, CrossRef thoroughly indexed the analysed open access journal articles.
+Metadata representing publisher and journal titles could be gathered for 99 % of the self-reported article and successfully merged with the Open APC cost data.
+Using metadata from CrossRef, therefore, could reduce the extensive validation work of bibliographic information in self-reported data-sets on open access publication fees provided that the reporting of the DOI along with the expenses is made mandatory.
+Drawing on CrossRef would also increase the comparability of cost data for future negotiations with publishers on open access agreements, and the open access spending between open data initiatives that apply the same reporting standards, as its metadata represent the dynamic landscape of academic publishing in terms of ongoing mergers of publishing houses or name changes.
 
-Because of the dynamic landscape of academic publishing and its representation in CrossRef's data curation efforts, it is important to consider the time-frame of metadata aggregation.
-In our case, we re-used metadata shortly after the data submission.
-However, for some cost analysis -- for instance to prepare negotiations with publishers on future schemes to fund open access journal articles -- it could be more feasible to re-normalize the complete data.
-The rossRef API provides incremental metadata updates that can be used to assess the current potential of future funding.
-While licensing information is incompletely covered in the CrossRef index so far, and therefore not analyzed in our study, the growing importance of facilitating text mining may result in more and more publishers sharing this information with CrossRef in the future.
+Another advantage of self-reported data-sets on the article-level is that they enable researchers and practitioners alike to study in which open access journals researchers from one institution actually publish, and to compare these findings with that of other universities or research organizations.
+For instance, our study revealed that the size of publication fee spending differed among the institutions with the  Max Planck Society accounting for almost 39 % of the overall articles.
+Many universities and research organization reported remarkably lower number of supported open access articles to the Open APC initiative.
+Using self-reported data, therefore, contributes to the understanding about how much and to what extent spending on open access publishing varies on the institutional level.
+This is particularly relevant given the increasingly important role open access publishing plays in recent negotiations between German universities and research organizations forming consortia on the one side and publishers on the other about financing scholarly publishing in future [allianz].
 
-Participation is voluntary.
+Our study is limited in some respects.
+One is that we cannot assess whether publishers and journals granted publication fee discounts seeing that the Open APC initiative does not track this kind of information.
+However, the large price ranges of particular journals suggests that varying pricing schemes are in place.
+Adding to this complexity, it seems that some institutions only paid parts of the publication fee.
+Take for instance the journal Nature Communication.
+Charges reported in our sample ranged between 2000 €, the DFG price cap, and 4.403 €.
+Such co-payments that involve several budgets are a discussed strategy to sustain publication funds at German universities [Fournier].
+In another case, one university included its charges for participating in the German SCOAP consortia and presumably divided the sum by the articles published in SCOAP journals.
+In this consortia managed by the German National Library of Science and Technology payments were not directly made on the article level.
+Instead, subscription costs between a participating library and a publisher were reconciled, and the reduction transferred to the consortia to finance publications in SCOAP journals.
+
+It must also be noted that reporting to the Open APC initiative is voluntary.
 Therefore, not all institutions in Germany that provide central funding of publication fees contribute cost data to this initiative.
@@
 Others pointed out that the workload to produce such a data-set could be too extensive [@deppe_2015].
+As there no registry of institutional open access funds or similar support structures exists, we cannot assess how many German universities and research organizations do not share their spendings on open access publication fees.
 
-While there may still be institutions that have no overview of their APC spending, we would like to emphasize that reporting data that is already available within an institution to the Open APC initiative should not lead to much additional work.
-The central incentive of this initiative is to make it as easy as possible to submit data to it.
-Being able to combine that data into one standardized data-set increases transparency and comparability, and gives institutions a better understanding of the overall development of open access publishing.
-In particular, we cannot see the harm of increased transparency; in fact, not knowing how much is spent is undoubtedly a disadvantage in dealing with publishers.
+Our analysis on how institutional spending per articles was distributed over publishers and journal titles shows that open access publishing is diverse and concentrated at the same time.
+While we were able to identify 139 individual publishing houses that were supported by the German universities and research organizations, the distribution is highly skewed.
+92 % of open access publication fee spending went to ten publishers, confirming a general high concentration of few publishers in current academic publishing.
+However, our study could not confirm that publications in open access journals owned by traditional publishing houses account for most of the spending on publication fees.
+Rather, open access publishers such as Public Library of Science (PloS), Copernicus GmbH or MPDI AG rank higher in our study than in the analyses of cost data in the UK.
+The study finally confirms the importance of mega-journals that account for a large share of all articles in fully open access journals.
+Conclusion
 
-Extending the data template to include information about funders or whether special agreements with publishers applied as suggested by @Pinfield_2015, would even increase the efforts needed to participate.
-However, with the growing demand for action in areas like the large-scale transition of toll-access journals to open access [@Schimmer_Geschuhn_Vogler], an updated data template could help institutions to better comply with these policy developments in the future.
-From our experience, another barrier to participate is the lack of skills in version control: the data submission itself is not always made directly by the institutions.
-Instead, they sent files to Bielefeld University Library with the request to make them available on GitHub on their behalf.
+# Conclusion{.unnumbered}
 
-Future work needs to focus on analyzing the cost data.
-Of particular interest are questions concerning the coverage of central funding schemes in comparison with open access publication output in general, and the use of other means to cover publication fees in particular.
-The Open APC initiative does not cover personal budgets or make price reductions explicit, but studies suggest that there is a possible gray area [@BJ_RK_2012; @Bj_rk_2015; @Lawson_waiver_2015].
-Existing study designs could be re-applied to examine the relationship between price on the one hand, and indexing coverage, journal prestige or management costs on the other [@Walters_2011; @Bj_rk_scientometrics_2015, @Pinfield_2015].
-This, in turn, helps to address the central question of future business models in scholarly publishing from an international perspective [@Pinfield_2015].
-
-<!-- # Conclusion{.unnumbered}
-
-The purpose of the Open APC initiative is to gain better insight on spending for open access.
+<!-- The purpose of the Open APC initiative is to gain better insight on spending for open access.
 The main operational idea—to make data submission easy by building upon existing data and by using popular open tools and data bases—has proven to be a feasible approach and to scale well.

### 2016-05-30 article.Rmd
some minor improvements

@@
 The aim of this study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-Drawing on self-reported data that German institutions openly shared through the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and British initiatives.
+Drawing on self-reported data from the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and UK initiatives.
 We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.
@@
 
-Instead of re-using publisher and journal information contained in the Open APC data-set, we fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOI's.
+Instead of re-using publisher and journal information contained in the Open APC data-set, we fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOIs.
 Although the Open APC initiative gathered metadata representing publishers and journals from CrossRef as well, this information was retrieved at the time when the participating institutions submitted the cost information.
-However, CrossRef regularly updates metadata to represent ongoing mergers of publishing houses or name changes.
-A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that now operates as Springer Nature.
-To reflect these changes in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set.
+The Open APC initiative tracked the date of data submission and the contributors for every data submission with Git, a version control system, increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub to be transparent over time.
+However, CrossRef regularly updated metadata to represent ongoing mergers of publishing houses or name changes.
+A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that operated as Springer Nature at the time of our study.
+To reflect these dynamics in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set.
 
 As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access CrossRef's REST API.
-We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals, were distinguished.
-It also contained disambiguated publisher information, thus avoiding confusion about naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014].
+We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the ISSN media types, the International Standard Serial Number used to identify journals, were distinguished.
+It also contained normalised publisher information, thus avoiding confusion about naming of publishing houses other studies were faced with when working with self-reported data [@woodward_2014].
 In cases where no bibliographic information could be obtained, we used the Open APC values.
-Since CrossRef is not the only registration agency for DOI's, but also the agencies DataCite and Medra mint DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
+Since CrossRef was not the only registration agency for DOIs, but also the agencies DataCite and Medra minted DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
 
 Data collection in this study also involved obtaining cost data from related open data initiatives.
-To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
+To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we reviewed the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
 For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles.
@@
 We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.
-Our comparison between the open data initiatives focussed on the last two years 2014 and 2015.Since Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the exchange rates of the full two-years as we could not determine the actual dates of spending.
+Our comparison between the open data initiatives focussed on the last two years 2014 and 2015.
+Because Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
 We excluded Jisc's spending in 2014 because full currency information in the data-set was not available.
@@
 Data collection methods of the Open APC initiative and those of the others differed in some aspects.
-For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institution were required to report, publication identifier in the Wellcome Trust data might also be added through automated compliance checks.
+For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institution were required to report, publication identifier in the Wellcome Trust data was also added through automated compliance checks.
+Our first screening of the data-sets revealed that some articles lacked a DOI.
 For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided only to compare German spending with that of other initiatives, but not its size and distribution over publisher and journal titles, as well as the indexing coverage in CrossRef.
-However, we made use of the literature to discuss our findings in relation to the other initiatives.
+However, we made use of the literature to discuss our findings in relation to these Austrian and UK initiatives.
 
@@
 
-```{r histm, fig.width=6, fig.height=2.5,fig.cap="Histogram APC"}
-require(ggplot2)
-## http://www.noamross.net/blog/2013/11/20/formatting-plots-for-pubs.html
-science_theme = theme(panel.grid.major = element_line(size = 0.5, color = "grey"), 
-    axis.line.x = element_line(size = 0.7, color = "black"),
-    axis.line.y = element_line(size = 0.7, color = "black"),
-    text = element_text(size = 14))
-
-ggplot(my_apc, aes(euro)) + 
-  geom_histogram(binwidth = 250, colour ="black", boundary = 0) + 
-  scale_x_continuous(limits = c(0,8000)) +
-  xlab("Publication fee in €") +
-  ylab("Articles") +
-  theme_bw() + science_theme
-
-```
-
 ```{r articles_year, fig.width=6, fig.height=2.5,fig.cap="Growth of Open APC Initiative"}
@@
 ggsave("figure_1.pdf", height=4, width=6)
+
+```
+
+```{r histm, fig.width=6, fig.height=2.5,fig.cap="Histogram APC"}
+require(ggplot2)
+## http://www.noamross.net/blog/2013/11/20/formatting-plots-for-pubs.html
+science_theme = theme(panel.grid.major = element_line(size = 0.5, color = "grey"), 
+    axis.line.x = element_line(size = 0.7, color = "black"),
+    axis.line.y = element_line(size = 0.7, color = "black"),
+    text = element_text(size = 14))
+
+ggplot(my_apc, aes(euro)) + 
+  geom_histogram(binwidth = 250, colour ="black", boundary = 0) + 
+  scale_x_continuous(limits = c(0,8000)) +
+  xlab("Publication fee in Euro") +
+  ylab("Articles") +
+  theme_bw() + science_theme
 
@@
 
-The data-set finally confirms the leading role of "mega-journals" in open access publishing, including the multidisciplinary PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline.
-In general, an estimated 14 out of more than 10,000 journals registered in DOAJ in 2015 accounted for up to 15–20 % of all articles published in full open access journals [@Bj_rk_2015].
-
 # Discussion{.unnumbered}
 
-In Germany, institutional spending on open access publication fees has increased over the years.
-With a share of …, the majority of open access articles German institutions reported to the Open APC initiative were published in fully open access journals.
+In Germany, institutional spending on open access publication fees has increased over the years, confirming the general trend of publication fees in academic publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
+With a share of 99%, the majority of open access articles German institutions reported to the Open APC initiative were published in fully open access journals.
 This presumably reflects the DFG funding policy which excludes the support of articles published in hybrid open access journals.
 The DFG has been financially supporting the implementation of central publication funds at more than 30 German universities since 2009.
-However, reviewing self-reported cost data from funders or countries that also support hybrid open access journals or open access books revealed a smaller proportions of payments in favor of articles in fully open access journals.
-Since open access publication fee spending is generally fragmented, we cannot answer whether German researchers avoid opting for open access when publishing in hybrid journals or use other budgets to pay publication fees required to make their work open access through these kind of journals.
+However, in reviewing self-reported cost data from funders or countries that also support hybrid open access journals or open access books, we revealed smaller proportions of payments in favour of articles in fully open access journals.
+Because open access publication fee spending is fragmented, we cannot answer whether German researchers avoided opting for open access when publishing in hybrid journals or used other budgets to pay publication fees required to make their work open access through these kind of journals.
 
-In our study, CrossRef thoroughly indexed the analysed open access journal articles.
-Metadata representing publisher and journal titles could be gathered for 99 % of the self-reported article and successfully merged with the Open APC cost data.
-Using metadata from CrossRef, therefore, could reduce the extensive validation work of bibliographic information in self-reported data-sets on open access publication fees provided that the reporting of the DOI along with the expenses is made mandatory.
+In our study, CrossRef thoroughly indexed open access journal articles in the Open APC data-set.
+We could gather metadata representing publisher and journal titles for 99 % articles, and successfully merged these information with the Open APC cost data.
+Using metadata from CrossRef, therefore, reduces the extensive validation work of bibliographic information provided that the reporting of the DOI along with the expenses is made mandatory.
 Drawing on CrossRef would also increase the comparability of cost data for future negotiations with publishers on open access agreements, and the open access spending between open data initiatives that apply the same reporting standards, as its metadata represent the dynamic landscape of academic publishing in terms of ongoing mergers of publishing houses or name changes.
 
-Another advantage of self-reported data-sets on the article-level is that they enable researchers and practitioners alike to study in which open access journals researchers from one institution actually publish, and to compare these findings with that of other universities or research organizations.
+Another advantage of self-reported data-sets on the article-level to disclose spending on open access publication fees is that they enable researchers and practitioners alike to study in which open access journals researchers from one institution actually publish, and to compare these findings with that of other universities or research organizations.
 For instance, our study revealed that the size of publication fee spending differed among the institutions with the  Max Planck Society accounting for almost 39 % of the overall articles.
@@
 Using self-reported data, therefore, contributes to the understanding about how much and to what extent spending on open access publishing varies on the institutional level.
-This is particularly relevant given the increasingly important role open access publishing plays in recent negotiations between German universities and research organizations forming consortia on the one side and publishers on the other about financing scholarly publishing in future [allianz].
+This is particularly relevant given the increasingly important role open access publishing plays in recent negotiations between German universities and research organizations forming consortia on the one side and publishers on the other about financing scholarly publishing in future [@allianz].
 
-Our study is limited in some respects.
+This study is limited in some respects.
 One is that we cannot assess whether publishers and journals granted publication fee discounts seeing that the Open APC initiative does not track this kind of information.
-However, the large price ranges of particular journals suggests that varying pricing schemes are in place.
-Adding to this complexity, it seems that some institutions only paid parts of the publication fee.
+However, the large price ranges of particular journals suggests that varying pricing schemes were in place.
+Adding to this complexity, it is likely that some institutions only paid parts of the publication fee.
 Take for instance the journal Nature Communication.
 Charges reported in our sample ranged between 2000 €, the DFG price cap, and 4.403 €.
-Such co-payments that involve several budgets are a discussed strategy to sustain publication funds at German universities [Fournier].
+Such co-payments that involve several budgets were a proposed strategy to sustain publication funds at German universities [@Fournier_2013].
 In another case, one university included its charges for participating in the German SCOAP consortia and presumably divided the sum by the articles published in SCOAP journals.
-In this consortia managed by the German National Library of Science and Technology payments were not directly made on the article level.
+In this national consortia, managed by the German National Library of Science and Technology, payments were not directly made per article.
 Instead, subscription costs between a participating library and a publisher were reconciled, and the reduction transferred to the consortia to finance publications in SCOAP journals.
@@
 Therefore, not all institutions in Germany that provide central funding of publication fees contribute cost data to this initiative.
-In a qualitative survey that also asked why German institutions are reluctant to share their cost data through the Open APC initiative one institution feared that increase in transparency would allow publishers to adjust prices in their favor.
+In a qualitative survey, asking why German institutions are reluctant to share their cost data through the Open APC initiative, one institution feared that increase in transparency would allow publishers to adjust prices in their favour.
 Others pointed out that the workload to produce such a data-set could be too extensive [@deppe_2015].
-As there no registry of institutional open access funds or similar support structures exists, we cannot assess how many German universities and research organizations do not share their spendings on open access publication fees.
+As no registry of institutional open access funds or similar support structures exists, we cannot assess the number of non-participants in Germany.
 
@@
 Rather, open access publishers such as Public Library of Science (PloS), Copernicus GmbH or MPDI AG rank higher in our study than in the analyses of cost data in the UK.
-The study finally confirms the importance of mega-journals that account for a large share of all articles in fully open access journals.
-Conclusion
+
+This study finally confirms the leading role of "mega-journals" in open access publishing, including the multidisciplinary PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline.
+In general, an estimated 14 out of more than 10,000 journals registered in DOAJ in 2015 accounted for up to 15–20 % of all articles published in full open access journals [@Bj_rk_2015].
 

### 2016-05-31 article.Rmd
streamline intrduction first try

@@
 
-Publication fees, often paid by funders or universities, are a widely discussed open access business model.
+<!-- Publication fees, often paid by funders or universities, are a widely discussed business model for open access journals.
 Yet, how and to what extent these activities are effective in terms of the number of supported research articles and associated costs remains under debate.
-This paper describes the Open APC initiative[^2], in which German universities and research organization share spendings on open access publication fees, and how it is currently implemented.
-More specifically, it addresses three problem areas when studying the economical effects of open access publishing: fragmentation of open access funding, variable pricing schemes and the comparison across research institutions.
-Such an approach extends methods and improves data collection activities for researchers and practitioners, as well as contribute to a better understanding of factors affecting the analysis of publication fees in open access publishing.
+This study contributes to a growing empirical base about institutional spending on open access publication fees.
+More specifically, it uses self-reported cost data from German universities and research organisations to examine how much they spent, and how the so funded open access journal articles were distributed over publishers and journal titles using bibliographic information from CrossRef.
+Such an approach extends methods and improves data collection activities for researchers and practitioners alike, as well as contribute to a better understanding of factors affecting the analysis of publication fees in open access publishing.
+-->
 
@@
 However, collecting information about what was funded is in most cases difficult.
-One reason why payments made for open access journal publications are often hard to track is that, on the one hand, they are fragmented across the budgets of funding agencies, research institutions, and libraries, and, on the other hand, taken from personal budgets.
-Furthermore, open access funding mostly exists in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
+One reason why payments made for open access journal publications are often hard to track is that they are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if no institutional support exists, are taken from personal budgets.
+Institutional open access funding mostly exists in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
 Personal budgets stand in contrast with those support structures and are likely used to cover low price publication fees [@Solomon_2011; @Bj_rk_2015].
-Along with the fragmentation of payments, funding for open access publications lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
+
+Along with the fragmentation of payments, another problem is that funding for open access publications lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014].
+This situation is similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
 It also remains unclear which factors contribute to price formation.[^4b] While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
 Other factors leading to a complex landscape of variable pricing schemes [@Pinfield_2015] include submission or page charges [@BJ_RK_2012].
-Hybrid journals substantially add to this complexity, because comprehensive offset systems to avoid paying for the same article twice, through subscription and publication fee, are rare, which, in turn, leads to the phenomena of "double dipping" in scholarly publishing [@Pinfield_2015].
+
+Hybrid journals substantially add to the complexity of open access funding.
+These journals, allowing articles to be published immediately as open access after a charge was paid, rely both on subscriptions and publication fees as revenue sources, which, in turn, leads to the phenomena of "double dipping" in academic publishing [@Pinfield_2015].
+Although the uptake of open access through hybrid journals were described as low [@Solomon_2012; @Bj_rk_learned_2014], this model has gained attention through recent science policies, notably because of open access policies from the UK [@Pinfield_Review_2015].
 
 This complex situation of fee-based open access publishing creates difficulties for researchers and practitioners alike.
-Because of fragmented payments, the extent of funding remains unclear.
-To increase transparency, some research funders have begun collecting and disclosing expenditures for open access journal articles as open data.
-As per definition, open data is data that "can be freely used, modified, and shared by anyone for any purpose" [@OpenDataHandbook].
-Therefore, opening up information about the funding of open access journal articles promises to enhance the discussion about current and future business models in academic publishing.
-To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc], who both released data on publication fees they had funded.
-The British not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
+Research addressing the question about how much was spent on open access journal articles either surveyed authors [@soap], collected price information from journal websites [@Walters_2011, @Solomon_2012], or used self-reported data [@Pinfield_2015].
+In the latter case, some research funders have begun collecting and disclosing expenditures for open access journal articles as open data.
+
+To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc].
+The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
 Disclosed as publicly available spreadsheets, these data-sets self-report expenditures along with bibliographic information, including title, journal and publisher, persistent identifier to the publisher's version, and a link to a deposit in a subject repository.
@@
 
-The open access landscape in Germany, which is the focus of this paper, shares the general problems of in-transparency regarding funding schemes and costs as discussed above.
-The Deutsche Forschungsgemeinschaft (DFG), the largest research funder in Germany, has been encouraging open access publishing since years.
-It launched its "Open-Access Publishing" program in 2009 that has strongly influenced the support of open access publication fees through funds managed by university libraries.[^3] With this program, the DFG aims to help universities to establish support structures for publishing in open access journals where authors are requested to pay a publication fee.
-To reduce administrative burdens, grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
-Examples include central invoicing schemes and related agreements between university libraries and publishers.
-Grantees are also required to report the institutional publication output and their fees paid for open access journal articles to the DFG on a regular basis, and to present the university-wide strategy to sustain the funds when DFG’s initial support runs out lately in 2019.
+In examining self-reported institutional payments made for open access articles in Germany, this paper seeks to contribute to the evolving empirical evidence on questions about open access publication fees as business model in academic publishing.
+In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has been encouraging open access publishing since years.
+The DFG launched its "Open-Access Publishing" program in 2009.
+This programme has strongly influenced how university libraries support open access publication fees through funds.[^3] To reduce administrative burdens, grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
+They include central invoicing schemes and related agreements between university libraries and publishers.
+Grantees are also required to report the institutional publication output and their publication fee spending to the DFG on a regular basis, and to present the university-wide strategy to sustain the funds when DFG’s initial support runs out lately in 2019.
+
 The DFG enforces a set of criteria grantees have to comply with, leading to similar implementations for supporting open access publishing across German universities: these criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding VAT) [@Fournier_2013].
@@
 
-The growing share of articles published in fee-based open access journals in recent years has led to calls for an unified approach towards funding of publication fees.
-The Allianz der Wissenschaftsorganisationen[^10c], representing all major research organizations in Germany, thus marks transparency as a major means to sustain an "adequate open access publication system" [@allianz].
-However, there are various ways to achieve this goal.
-The existing approaches in Austria and the United Kingdom have one institution in charge to collect and analyze the data.
-The history of the Open APC initiative is rather bottom-up: In May 2014, Bielefeld University Library began to share its expenditures for publication fees.
-The library put its approach to the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11] as a basis for discussion, and invited others to participate.
-Reflecting the increasing demand for publicly available data, contributions from Universität Regensburg and Universität Hannover followed soon after.
-As of writing, 29 universities and research institutes voluntary reported their data to the Open APC initiative to be included into a unified data-set of all expenditures.
+The evolving institutional support structures to cover open access publication fees has led to calls for an unified approach towards supporting open access journal publishing.
+The Allianz der Wissenschaftsorganisationen[^10c], representing all major research organizations in Germany, marks in this regard transparency as one way to sustain an "adequate open access publication system" [@allianz].
+Reflecting Austrian and UK initiatives to share institutional spending on open access publication fees as open data, as well as professional discussions on open access publishing, Bielefeld University Library began to openly share its payments in May 2014.
+After putting its approach to the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11] as a basis for discussion, the library invited others to participate under the umbrella of the Open APC initiative, and contributions from other German research institutions followed soon after.
 
-The aim of this study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-Drawing on self-reported data from the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and UK initiatives.
+Drawing on self-reported data from this open data initiative, the aim of the study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
+The analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and UK initiatives.
 We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.

### 2016-05-31 article.Rmd
back-uping

@@
 To cover these fees, authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
-However, collecting information about what was funded is in most cases difficult.
-One reason why payments made for open access journal publications are often hard to track is that they are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if no institutional support exists, are taken from personal budgets.
-Institutional open access funding mostly exists in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
+Yet, how and to what extent these activities are effective in terms of the number of supported research articles and associated costs remains under debate.
+One reason why studying what was funded is in most cases difficult is that payments are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if  institutional support is limited, are taken from personal budgets.
+Recent research indicate that institutional open access funding mostly exists in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
 Personal budgets stand in contrast with those support structures and are likely used to cover low price publication fees [@Solomon_2011; @Bj_rk_2015].
 
-Along with the fragmentation of payments, another problem is that funding for open access publications lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014].
-This situation is similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
+Along with the fragmentation of payments, another problem in the empirical study of open access journals is that funding for open access publications lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
 It also remains unclear which factors contribute to price formation.[^4b] While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
@@
 This complex situation of fee-based open access publishing creates difficulties for researchers and practitioners alike.
-Research addressing the question about how much was spent on open access journal articles either surveyed authors [@soap], collected price information from journal websites [@Walters_2011, @Solomon_2012], or used self-reported data [@Pinfield_2015].
-In the latter case, some research funders have begun collecting and disclosing expenditures for open access journal articles as open data.
+Research addressing the question about how much was spent on open access journal articles either surveyed authors [@soap], collected price information from journal websites [@Walters_2011; @Solomon_2012], or used self-reported data [@Pinfield_2015].
 
+-> summary of empirical findings
+
+<!-- In the latter case, some research funders have begun collecting and disclosing expenditures for open access journal articles as open data.
 To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc].
@@
 In the case of the Wellcome Trust, crowd-sourcing data cleaning activities through a Google spreadsheet in combination with checks against bibliographic sources massively improved the spending data (see comments in @wellcome_apc).
+-->
 
-In examining self-reported institutional payments made for open access articles in Germany, this paper seeks to contribute to the evolving empirical evidence on questions about open access publication fees as business model in academic publishing.
+In examining self-reported institutional payments made for open access articles in Germany, this paper seeks to contribute to the evolving empirical evidence on questions about open access publication fees.
 In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has been encouraging open access publishing since years.
@@
 
-Instead of re-using publisher and journal information contained in the Open APC data-set, we fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOIs.
+We fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOIs.
 Although the Open APC initiative gathered metadata representing publishers and journals from CrossRef as well, this information was retrieved at the time when the participating institutions submitted the cost information.
@@
 A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that operated as Springer Nature at the time of our study.
-To reflect these dynamics in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set.
+To reflect these dynamics in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set instead of re-using publisher and journal information contained in the Open APC data-set.
 
@@
 In cases where no bibliographic information could be obtained, we used the Open APC values.
-Since CrossRef was not the only registration agency for DOIs, but also the agencies DataCite and Medra minted DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
+Because CrossRef was not the only registration agency for DOIs, but also the agencies DataCite and Medra minted DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
 
 Data collection in this study also involved obtaining cost data from related open data initiatives.
-To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we reviewed the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
+To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we reviewed the openly available data-sets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
 For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles.
@@
 For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided only to compare German spending with that of other initiatives, but not its size and distribution over publisher and journal titles, as well as the indexing coverage in CrossRef.
-However, we made use of the literature to discuss our findings in relation to these Austrian and UK initiatives.
 

### 2016-06-01 article.Rmd
introduction

@@
 
+## General Background{.unnumbered}
+
 <!-- Publication fees, often paid by funders or universities, are a widely discussed business model for open access journals.
@@
 The rise of open access journals matches the increasing relevance of publication fees in academic publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-To cover these fees, authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
-Yet, how and to what extent these activities are effective in terms of the number of supported research articles and associated costs remains under debate.
-One reason why studying what was funded is in most cases difficult is that payments are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if  institutional support is limited, are taken from personal budgets.
-Recent research indicate that institutional open access funding mostly exists in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
-Personal budgets stand in contrast with those support structures and are likely used to cover low price publication fees [@Solomon_2011; @Bj_rk_2015].
+To cover these fees, sometimes also referred to as article-processing charges (APC), authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
+Yet, how and to what extent these research support activities are effective in terms of the number of supported articles and associated costs remains under debate.
 
-Along with the fragmentation of payments, another problem in the empirical study of open access journals is that funding for open access publications lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
+One reason why studying how much was spent on publication fees is in most cases difficult is that payments for these charges are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
+Another key problem for the empirical study of institutional spending on publication fees is that funding for open access publications lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
 It also remains unclear which factors contribute to price formation.[^4b] While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
 Other factors leading to a complex landscape of variable pricing schemes [@Pinfield_2015] include submission or page charges [@BJ_RK_2012].
+Hybrid journals substantially add to this complexity of open access funding [@Kingsley_2014].
+These journals, allowing articles to be published immediately as open access after a charge was paid, rely both on subscriptions and publication fees as revenue sources.
+Although the uptake of open access through hybrid journals was described as lower and more expensive compared to that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014], this model has gained attention through recent science policies, notably because of open access policies from the UK [@Pinfield_Review_2015].
 
-Hybrid journals substantially add to the complexity of open access funding.
-These journals, allowing articles to be published immediately as open access after a charge was paid, rely both on subscriptions and publication fees as revenue sources, which, in turn, leads to the phenomena of "double dipping" in academic publishing [@Pinfield_2015].
-Although the uptake of open access through hybrid journals were described as low [@Solomon_2012; @Bj_rk_learned_2014], this model has gained attention through recent science policies, notably because of open access policies from the UK [@Pinfield_Review_2015].
+Previous empirical studies asking how much was being spent for publishing in open access journals confirmed that authors make use of funding or institutional support for covering open access publication fees.
+In 2010, the SOAP project carried out a comprehensive survey and asked 9,645 authors from various disciplines how they financed publication fees [@soap].
+Of the respondents, 12 % had used personal budgets.
+By contrast, most had access to institutional support structures or research funding.
+These findings are consistent with that of other surveys, adding that sources of funding exists mostly in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
+Personal budgets, on the other hand, are likely used to cover low price publication fees [@Solomon_2011; @Bj_rk_2015].
 
-This complex situation of fee-based open access publishing creates difficulties for researchers and practitioners alike.
-Research addressing the question about how much was spent on open access journal articles either surveyed authors [@soap], collected price information from journal websites [@Walters_2011; @Solomon_2012], or used self-reported data [@Pinfield_2015].
+Asked about the amount spent on publication fees, 22,977  respondents in the SOAP survey mostly reported prices between 501 and 1000 € (12,9%), and between 1001 and 3000 € (9.9 %).
+However, more than half of the authors answering this question had not to pay a publication fee when publishing in open access journals, and 14 % did not report the amount spent.
+While examining the indexing coverage of fully open access journals in subject-specific databases, @Walters_2011 collected publication fees as journal characteristic in 2009.
+This study reported an average fee of 923 $.
+Journals indexed in the Web of Science charged 1,171 $ on average.
+@Solomon_2012 calculated an similar average of 906 $ for publishing in fully open access journals.
+However, they found a large price variation between 8 to 3,900 $.
+They suggested to cluster fully open access journals using publication fees into several groups.
+In this regard, while one set consists of costly high-impact journals (2000 $ - 4000 $), a large number of journals charge 1500 $ - 2000 $.
+Journals with broad scope like PLOS ONE form another segment charging between 1000 - 1500 $, just as mid-priced journals from commercial publishers (500 - 1000 $), society journals (below 500 $), and low-priced journals (below 200 $).
 
--> summary of empirical findings
+Complementary to survey data or the consultation of journal web pages for determining spending on publication fees, some research funders and research performing institutions have recently begun to disclose their expenditures on publication fees as open data.
+To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc].
+The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
+The latter was examined by @Pinfield_2015.
+Although the average spending on publication fees remained stable across the universities, they found large price variations, as well as variations in the number of articles the analysed UK universities supported between 2007 - 2014.
 
-<!-- In the latter case, some research funders have begun collecting and disclosing expenditures for open access journal articles as open data.
+<!-- This complex situation of fee-based open access publishing creates difficulties for researchers and practitioners alike.
+Some European initiatives have therefore begun to collect and disclose their expenditures on open access journal articles as open data.
 To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc].
@@
 Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates.
-In the case of the Wellcome Trust, crowd-sourcing data cleaning activities through a Google spreadsheet in combination with checks against bibliographic sources massively improved the spending data (see comments in @wellcome_apc).
--->
+In the case of the Wellcome Trust, crowd-sourcing data cleaning activities through a Google spreadsheet in combination with checks against bibliographic sources massively improved the spending data (see comments in @wellcome_apc).-->
 
-In examining self-reported institutional payments made for open access articles in Germany, this paper seeks to contribute to the evolving empirical evidence on questions about open access publication fees.
+## Central funding for publication fees in Germany{.unnumbered}
+
+This paper focuses on how much German universities and research organization spent on open access publication fees.
 In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has been encouraging open access publishing since years.
@@
 After putting its approach to the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11] as a basis for discussion, the library invited others to participate under the umbrella of the Open APC initiative, and contributions from other German research institutions followed soon after.
+
+## Research question{.unnumbered}
 

### 2016-06-01 article.Rmd
improving intro about german situation

@@
 The rise of open access journals matches the increasing relevance of publication fees in academic publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-To cover these fees, sometimes also referred to as article-processing charges (APC), authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
+To cover publication fees, also referred to as article-processing charges (APC), authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
 Yet, how and to what extent these research support activities are effective in terms of the number of supported articles and associated costs remains under debate.
@@
 One reason why studying how much was spent on publication fees is in most cases difficult is that payments for these charges are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
-Another key problem for the empirical study of institutional spending on publication fees is that funding for open access publications lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
-It also remains unclear which factors contribute to price formation.[^4b] While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
-Other factors leading to a complex landscape of variable pricing schemes [@Pinfield_2015] include submission or page charges [@BJ_RK_2012].
+Another key problem is that funding for open access journals using publication fees lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
+It also remains unclear which factors contribute to price formation.[^4b] While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts, and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
+Other factors leading to variable pricing schemes include submission or page charges [@BJ_RK_2012].
 Hybrid journals substantially add to this complexity of open access funding [@Kingsley_2014].
@@
 
-Previous empirical studies asking how much was being spent for publishing in open access journals confirmed that authors make use of funding or institutional support for covering open access publication fees.
-In 2010, the SOAP project carried out a comprehensive survey and asked 9,645 authors from various disciplines how they financed publication fees [@soap].
-Of the respondents, 12 % had used personal budgets.
-By contrast, most had access to institutional support structures or research funding.
+Previous empirical studies asking how much was being spent for publishing in open access journals confirmed that authors make use of funding or institutional support for covering publication fees.
+In 2010, the SOAP project carried out a comprehensive survey in which 9,645 authors from various disciplines provided insights into how they financed publication fees [@soap].
+The majority of the respondents used institutional support structures or research funding.
+By contrast, 12 % paid publication fees individually.
 These findings are consistent with that of other surveys, adding that sources of funding exists mostly in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
@@
 
-Asked about the amount spent on publication fees, 22,977  respondents in the SOAP survey mostly reported prices between 501 and 1000 € (12,9%), and between 1001 and 3000 € (9.9 %).
-However, more than half of the authors answering this question had not to pay a publication fee when publishing in open access journals, and 14 % did not report the amount spent.
-While examining the indexing coverage of fully open access journals in subject-specific databases, @Walters_2011 collected publication fees as journal characteristic in 2009.
-This study reported an average fee of 923 $.
-Journals indexed in the Web of Science charged 1,171 $ on average.
+Asked about the amount being spent on publication fees, 22,977 SOAP survey respondents reported prices between 501 and 1000 € (12,9%), and between 1,001 and 3,000 € (9.9 %).
+However, more than half of the authors answering this question had not to pay a publication fee to get publish in an open access journal, and 14 % did not report the amount spent.
+While examining the indexing coverage of fully open access journals in bibliographic databases in 2009, @Walters_2011  also provided publication fees as journal characteristics.
+Their study reported an average publication fee of 923 $, and for Journals indexed in the Web of Science 1,171 $ on average.
 @Solomon_2012 calculated an similar average of 906 $ for publishing in fully open access journals.
-However, they found a large price variation between 8 to 3,900 $.
-They suggested to cluster fully open access journals using publication fees into several groups.
-In this regard, while one set consists of costly high-impact journals (2000 $ - 4000 $), a large number of journals charge 1500 $ - 2000 $.
-Journals with broad scope like PLOS ONE form another segment charging between 1000 - 1500 $, just as mid-priced journals from commercial publishers (500 - 1000 $), society journals (below 500 $), and low-priced journals (below 200 $).
+However, they found a large price variance across journals and articles.
+Accordingly, they suggested to cluster fully open access journals using publication fees into several groups.
+In this regard, while one set consisted of costly high-impact journals (2,000 $ - 4,000 $), a large number of journals charged 1,500 $ - 2,000 $.
+Journals with broad scope like PLOS ONE formed another segment charging between 1,000 - 1,500 $, just as mid-priced journals from commercial publishers (500 - 1,000 $), society journals (below 500 $), and low-priced journals (below 200 $).
 
-Complementary to survey data or the consultation of journal web pages for determining spending on publication fees, some research funders and research performing institutions have recently begun to disclose their expenditures on publication fees as open data.
+Complementary to survey data or the consultation of journal web pages for determining spending on publication fees, some research funders and research performing institutions have recently begun to disclose their spending on publication fees as open data.
 To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc].
 The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
-The latter was examined by @Pinfield_2015.
-Although the average spending on publication fees remained stable across the universities, they found large price variations, as well as variations in the number of articles the analysed UK universities supported between 2007 - 2014.
+Disclosed as publicly available spreadsheets, these data-sets self-report payments made along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
+Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates.
+Parts of Jisc cost data was examined by @Pinfield_2015.
+Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as variations in the number of articles UK universities supported between 2007 - 2014.
 
@@
 The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
-Disclosed as publicly available spreadsheets, these data-sets self-report expenditures along with bibliographic information, including title, journal and publisher, persistent identifier to the publisher's version, and a link to a deposit in a subject repository.
+Disclosed as publicly available spreadsheets, these data-sets self-report payments made along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version, and a link to a deposit in a subject repository.
 Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates.
@@
 This paper focuses on how much German universities and research organization spent on open access publication fees.
-In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has been encouraging open access publishing since years.
-The DFG launched its "Open-Access Publishing" program in 2009.
-This programme has strongly influenced how university libraries support open access publication fees through funds.[^3] To reduce administrative burdens, grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
+In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced through its "Open-Access Publishing" program how universities manage institutional support for publication fees.[^3] Before the DFG started to pay for centrally funded publication fees on a pro rata basis in 2011, and similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012], only few central funds existed [@eppelin_2012].
+The DFG has enforced a set of criteria grantees have to comply with, leading to similar implementations for supporting open access publishing across German universities: these criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding value added tax).
+Grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
 They include central invoicing schemes and related agreements between university libraries and publishers.
-Grantees are also required to report the institutional publication output and their publication fee spending to the DFG on a regular basis, and to present the university-wide strategy to sustain the funds when DFG’s initial support runs out lately in 2019.
+[@Fournier_2013].
 
-The DFG enforces a set of criteria grantees have to comply with, leading to similar implementations for supporting open access publishing across German universities: these criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding VAT) [@Fournier_2013].
 Research institutes organized in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this funding program, contributing to the diversity of schemes in Germany.
-In response, some organizations have adopted similar processes to support authors.
+But in response, some organizations have adopted similar processes to support authors.
 The Max-Planck-Gesellschaft operates their long-lasting open access activities, including handling spending and publisher agreements centrally, through the Max Planck Digital Library [@Schimmer_2013; @Sikora_2015], while the Leibniz-Gemeinschaft set up a dedicated open access fund in 2016.
@@
 The evolving institutional support structures to cover open access publication fees has led to calls for an unified approach towards supporting open access journal publishing.
-The Allianz der Wissenschaftsorganisationen[^10c], representing all major research organizations in Germany, marks in this regard transparency as one way to sustain an "adequate open access publication system" [@allianz].
+The Allianz der Wissenschaftsorganisationen[^10c], a science policy board representing all major research organizations in Germany, marks cost transparency as one way to sustain an "adequate open access publication system" [@allianz].
 Reflecting Austrian and UK initiatives to share institutional spending on open access publication fees as open data, as well as professional discussions on open access publishing, Bielefeld University Library began to openly share its payments in May 2014.
-After putting its approach to the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11] as a basis for discussion, the library invited others to participate under the umbrella of the Open APC initiative, and contributions from other German research institutions followed soon after.
+After engaging with the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11] other German institutions joined under the umbrella of the Open APC initiative soon after.
 
@@
 
-Drawing on self-reported data from this open data initiative, the aim of the study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-The analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and UK initiatives.
+The aim of the study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
+Using self-reported cost data from the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and UK initiatives.
 We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.

### 2016-06-02 article.Rmd
cost review jisc and wellcome trust

@@
 
-One reason why studying how much was spent on publication fees is in most cases difficult is that payments for these charges are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
+One reason why studying how much is being spent on publication fees is in most cases difficult is that payments for these charges are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
 Another key problem is that funding for open access journals using publication fees lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
@@
 Parts of Jisc cost data was examined by @Pinfield_2015.
-Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as variations in the number of articles UK universities supported between 2007 - 2014.
+Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as a varying number of articles UK universities supported between 2007 - 2014.
 

### 2016-06-02 data/cost_data_uk/README.Rmd
cost review jisc and wellcome trust

# Log: Review of publication fees being spent by FWF, Wellcome Trust and JISC

## Data Sources

All data files were downloaded on 2 June 2016.

### FWF

- Katharina Rieck, Doris Haslinger, Sasa Meischke-Ilic, Ünzüle Kirindi-Hentschel, & Falk Reckling. (2016). Austrian Science Fund (FWF) Publication Cost Data 2015. Figshare. http://doi.org/10.6084/m9.figshare.3180166

- Falk Reckling, & Katharina Rieck. (2015). Austrian Science Fund (FWF) Publication Cost Data 2014. Figshare. http://doi.org/10.6084/m9.figshare.1378610

### JISC

- Stuart Lawson. (2015). APC data for 25 UK higher education institutions - 2014. Figshare. http://doi.org/10.6084/m9.figshare.1305596.v5

- Stuart Lawson. (2016). APC data for 27 UK higher education institutions in 2015. Figshare. http://doi.org/10.6084/m9.figshare.1507481.v4

### Wellcome Trust

- Robert Kiley. (2016). Wellcome Tust/COAF spend on open access publishing (article processing charges) - 2014-15. Figshare. http://doi.org/10.6084/m9.figshare.3118936.v1

- Robert Kiley. (2015). Wellcome Trust open access (OA) spend and compliance monitoring: 2013-14. Figshare. http://doi.org/10.6084/m9.figshare.1321361.v5

## Pre-processing

Before loading cost data into the R console, we removed windows encoding from the Jisc data files. We also removed currency symbols from the Wellcome Trust spreadsheets and converted them to csv files with LibreOffice.

Pre-processed files were stored in this folder.

## Converting currencies

To summarize  Wellcome Trust's and Jisc's spending, we converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank. Our comparison between the open data initiatives focussed on the last two years 2014 and 2015. Because Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.

## Cost aggregation

Note, we used the accompanying data report to determine FWF publication fee spending

### JISC 2014

```{r}
jisc_14 <- read.csv("jisc_2014.csv", header = TRUE, sep =",")
jisc_14$APC <- as.numeric(as.character(jisc_14$APC.paid....including.VAT.if.charged))

# convert to average euro exchange rate
jisc_14$APC <- jisc_14$APC * 1.2411

apc_sum <- aggregate(jisc_14$APC, list(jisc_14$Type.of.publication), sum, na.rm = TRUE)
apc_n <- as.data.frame(table(jisc_14$Type.of.publication))

tt <- data.frame(apc_n, apc_sum[,c("x")])

colnames(tt) <- c("Type", "Freq", "Fees paid in Euro")

tt$`Mean Fee` <- tt$`Fees paid in Euro` / tt$Freq

tt
sum(tt$`Fees paid in Euro`)

write.csv(tt, "aggregated_jisc_2014.csv")
```

### JISC 2015

```{r}
jisc_15 <- read.csv("jisc_2015.csv", header = TRUE, sep =",")

jisc_15$APC <- as.numeric(as.character(jisc_15$APC.paid....including.VAT.if.charged))

# convert to average euro exchange rate
jisc_15$APC <- jisc_15$APC * 1.3785

apc_sum <- aggregate(jisc_15$APC, list(jisc_15$Type.of.publication), sum, na.rm = TRUE)

apc_n <- as.data.frame(table(jisc_15$Type.of.publication))

tt <- data.frame(apc_n, apc_sum[,c("x")])

colnames(tt) <- c("Type", "Freq", "Fees paid in Euro")

tt$`Mean Fee` <- tt$`Fees paid in Euro` / tt$Freq

tt

sum(tt$`Fees paid in Euro`)

write.csv(tt, "aggregated_jisc_2015.csv")
```

### wellcome 2013-14

```{r}
wellcome_13_14 <- read.csv("wellcome_13_14.csv", header = T, sep =",")
wellcome_13_14$APC <- wellcome_13_14$Cost......inc.VAT.when.charged.

# convert to average euro exchange rate
wellcome_13_14$APC <- wellcome_13_14$APC * 1.2094

apc_n <- data.frame(table(wellcome_13_14$Journal.Type))
apc_sum <- aggregate(wellcome_13_14$APC, list(wellcome_13_14$Journal.Type), sum, na.rm = TRUE)
apc_mean <- aggregate(wellcome_13_14$APC, list(wellcome_13_14$Journal.Type), mean, na.rm = TRUE)

tt <- data.frame(apc_n, apc_sum[,c("x")], apc_mean[,c("x")])

colnames(tt) <- c("Type", "Freq", "Fees paid in Euro", "Mean Fee paid in Euro")

tt

sum(tt$`Fees paid in GBP`)

write.csv(tt, "aggregated_wellcome_13_14.csv")
```

### wellcome 2014 - 2014

```{r}
wellcome_14_15 <- read.csv("wellcome_14_15.csv", header = T, sep =",")
wellcome_14_15$APC <- wellcome_14_15$Cost......inc.VAT.when.charged.

# convert to average euro exchange rate
wellcome_14_15$APC <- wellcome_14_15$APC * 1.3099

apc_n <- data.frame(table(wellcome_14_15$Journal.Type))
apc_sum <- aggregate(wellcome_14_15$APC, list(wellcome_14_15$Journal.Type), sum, na.rm = TRUE)
apc_mean <- aggregate(wellcome_14_15$APC, list(wellcome_14_15$Journal.Type), mean, na.rm = TRUE)

tt <- data.frame(apc_n, apc_sum[,c("x")], apc_mean[,c("x")])

colnames(tt) <- c("Type", "Freq", "Fees paid in Euro", "Mean Fee paid in Euro")

tt

sum(tt$`Fees paid in GBP`)

write.csv(tt, "aggregated_wellcome_14_14.csv")
```

### 2016-06-02 article.Rmd
aggregated sum reviewed and compared with open apc
adding subsection in results about the price comparision

@@
 We also observe that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc))  *100`%) of the publication fees were paid in accordance with the DFG price cap of 2,000 €.
-Whereas related open data initiatives in Austria and the United Kingdom reported a large share of spending for hybrid journal articles, the situation in Germany is different: only  `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` articles in hybrid journals were reported by `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` research institutions, accounting for `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of the overall payments.
 
@@
 ```
+
+## Comparision of related cost data-sets
+
+Table 2 compares Open APC spending data with that of the Austrian FWF, as well as with Jisc's and Wellcome Trust's expenditures.
+Prices were converted according to the average Euro exchange rate of the examined periods, and gathered for both fully and hybrid open access journals.
+The comparison reveals that the Open APC initiative lacked cost information about hybrid journals, whereas the related Austrian and UK open data initiatives could report a large share of spending for these journals between 2014 and 2015.
+This situation likely reflected the DFG-inspired funding policy of most participating institutions that excluded support for publications in hybrid journals.
+Over the years 2005 – 2015, `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` German universities and research institutions  reported `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` hybrid journal articles to the Open APC initiative, accounting for `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of the overall payments.
+In terms of the number of supported articles and the amount being spent on publication fees, by contrast, the Open APC data-set provided the most comprehensive price information for fully open access journals compared to what the Austrian and UK initiatives had reported.
+
+```{r results='asis'}
+tt <- read.csv("data/aggregates_fees_paid.csv", header = TRUE, sep =",")
+colnames(tt) <- c("Cost data-set", "Journal Type", "Articles funded", "Total costs in Euro", "Mean")
+pander::pandoc.table(tt,split.cells = 5 , big.mark = ",", caption = 'Comparision of cost data per period and journal type  (in €)', justify = 'lrrrr')
+```
+
+Comparison of average prices suggests that publishing in hybrid journal was more expensive than in fully open access journals.
+Price differentials between these two categories were also reported earlier, concluding that prices for fully open access journals were generally lower [@Solomon_2012; @Pinfield_2015].
+In 2014 and 2015, the mean price for fully open access journals calculated from all data-sets was below the DFG price cap of 2,000 €.
 

### 2016-06-02 article.Rmd
minor fixes comparision

@@
 Some European initiatives have therefore begun to collect and disclose their expenditures on open access journal articles as open data.
-To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc].
+To our knowledge, the first research funders providing such data were the Wellcome Trust [@fwf_apc_13] and the Austrian Science Fund FWF [@wellcome_apc_12].
 The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
@@
 
-Data collection in this study also involved obtaining cost data from related open data initiatives.
-To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we reviewed the openly available data-sets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc.
-For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles.
-In the case of FWF, we gathered the cost information from accompanying reports.
-We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.
+Data collection also involved obtaining cost data from related open data initiatives.
+To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we reviewed the openly available data-sets from the the Austrian Science Fund FWF [@fwf_apc_14; @fwf_apc_15], Jisc [@jisc_14; @jisc_15] and the Wellcome Trust [@wellcome_apc_13_14; @wellcome_apc_14_15].
+For analysis, we obtained the overall publication fee spending to support fully and hybrid open access journal articles.
+In the case of FWF, we gathered the cost information from the accompanying spending reports.
+We used the spreadsheet data to calculate  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.
 Our comparison between the open data initiatives focussed on the last two years 2014 and 2015.
 Because Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
-We excluded Jisc's spending in 2014 because full currency information in the data-set was not available.
+We excluded articles from the analysis for which neither price information nor the journal type could be obtained.
+In the case of Jisc's 2014 data [@jisc_14], 4.861.772 € being spent on 2,812 publications were not eligible for further analysis because no publication type was given in the data-set.
 

### 2016-06-03 article.Rmd
retooled introduction

@@
 The rise of open access journals matches the increasing relevance of publication fees in academic publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-To cover publication fees, also referred to as article-processing charges (APC), authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
+To cover these fees, also referred to as article-processing charges (APC), authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
 Yet, how and to what extent these research support activities are effective in terms of the number of supported articles and associated costs remains under debate.
 
-One reason why studying how much is being spent on publication fees is in most cases difficult is that payments for these charges are fragmented across the budgets of funding agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
-Another key problem is that funding for open access journals using publication fees lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
-It also remains unclear which factors contribute to price formation.[^4b] While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts, and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
+One reason why the study of open access journals using publication fees is in most cases difficult is that total spending on publication fees is fragmented across the budgets of grant agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
+Asking 9,645 authors from various disciplines how they financed publication fees, a comprehensive survey in 2010 revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.
+By contrast, 12 % paid publication fees individually [@soap].
+These findings are consistent with that of other studies, adding that sources of funding mostly exists in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
+Personal budgets, on the other hand, are likely used to cover low price publication fees [@Solomon_2011; @Bj_rk_2015].
+
+Another key problem for the study of open access is that funding for open access journals using publication fees lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
+However, empirical studies examining publication fees gathered from journal websites report similar findings.
+While the average publication fee reported ranged between 904 $ [@Solomon_2012] and 923 $ [@Walters_2011], considerable price variances across journals and publishers were observed.
+Accordingly, @Solomon_2012 suggested to cluster fully open access journal using publication fees into several groups.
+In descending order, high-impact journals charged most, followed by biomedicine journals from commercial publishers, large multi-disciplinary journals, and mid-price ranged journals from commercial publishers covering a large spectrum of disciplines.
+Lower priced journals were published by academic societies and by publishers from low-income countries.
+
+It remains unclear which other factors contribute to this varying pricing.
+Generally, these might include article processing, impact, rejection rates, management and investment, and profit margins [@Van_Noorden_2013].
+While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts, and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
 Other factors leading to variable pricing schemes include submission or page charges [@BJ_RK_2012].
+
 Hybrid journals substantially add to this complexity of open access funding [@Kingsley_2014].
@@
 
-Previous empirical studies asking how much was being spent for publishing in open access journals confirmed that authors make use of funding or institutional support for covering publication fees.
-In 2010, the SOAP project carried out a comprehensive survey in which 9,645 authors from various disciplines provided insights into how they financed publication fees [@soap].
-The majority of the respondents used institutional support structures or research funding.
-By contrast, 12 % paid publication fees individually.
-These findings are consistent with that of other surveys, adding that sources of funding exists mostly in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
-Personal budgets, on the other hand, are likely used to cover low price publication fees [@Solomon_2011; @Bj_rk_2015].
-
-Asked about the amount being spent on publication fees, 22,977 SOAP survey respondents reported prices between 501 and 1000 € (12,9%), and between 1,001 and 3,000 € (9.9 %).
-However, more than half of the authors answering this question had not to pay a publication fee to get publish in an open access journal, and 14 % did not report the amount spent.
-While examining the indexing coverage of fully open access journals in bibliographic databases in 2009, @Walters_2011  also provided publication fees as journal characteristics.
-Their study reported an average publication fee of 923 $, and for Journals indexed in the Web of Science 1,171 $ on average.
-@Solomon_2012 calculated an similar average of 906 $ for publishing in fully open access journals.
-However, they found a large price variance across journals and articles.
-Accordingly, they suggested to cluster fully open access journals using publication fees into several groups.
-In this regard, while one set consisted of costly high-impact journals (2,000 $ - 4,000 $), a large number of journals charged 1,500 $ - 2,000 $.
-Journals with broad scope like PLOS ONE formed another segment charging between 1,000 - 1,500 $, just as mid-priced journals from commercial publishers (500 - 1,000 $), society journals (below 500 $), and low-priced journals (below 200 $).
-
-Complementary to survey data or the consultation of journal web pages for determining spending on publication fees, some research funders and research performing institutions have recently begun to disclose their spending on publication fees as open data.
-To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc].
+To address the problems of fragmented spending on publication fees and in-transparency about what was paid, some European research funders and research performing institutions have recently begun to disclose their spending on publication fees as open data.
+To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc_12] and the Austrian Science Fund FWF [@fwf_apc_13].
 The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
 Disclosed as publicly available spreadsheets, these data-sets self-report payments made along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
-Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates.
+Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates [@woodward_2014].
 Parts of Jisc cost data was examined by @Pinfield_2015.
@@
 This paper focuses on how much German universities and research organization spent on open access publication fees.
-In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced through its "Open-Access Publishing" program how universities manage institutional support for publication fees.[^3] Before the DFG started to pay for centrally funded publication fees on a pro rata basis in 2011, and similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012], only few central funds existed [@eppelin_2012].
-The DFG has enforced a set of criteria grantees have to comply with, leading to similar implementations for supporting open access publishing across German universities: these criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding value added tax).
+In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced how universities manage institutional support for publication fees.[^3] Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" program  in 2011, and similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012], only few central funds existed [@eppelin_2012].
+Because the DFG has enforced a set of criteria grantees have to comply with, similar implementations for supporting open access publishing across German universities exists [@Fournier_2013]: these criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding value added tax).
 Grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
-They include central invoicing schemes and related agreements between university libraries and publishers.
-[@Fournier_2013].
+They include central invoicing schemes and related agreements between university libraries and publishers [@Fournier_2013].
 
@@
 
-[^4b]: These might include article processing, impact, rejection rates, management and investment, and profit margins.
-See @Van_Noorden_2013 for a general discussion and @Gumpenberger_2012 and @Bj_rk_scientometrics_2015 for discussions of journal impact and quality.
-
 [^5]: CrossRed: Text and Data Mining for Researchers: <http://tdmsupport.crossref.org/researchers/>

### 2016-06-03 article.Rmd
adding springer nature analysis and remove tables and figures from the manuscript

@@
 In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced how universities manage institutional support for publication fees.[^3] Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" program  in 2011, and similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012], only few central funds existed [@eppelin_2012].
-Because the DFG has enforced a set of criteria grantees have to comply with, similar implementations for supporting open access publishing across German universities exists [@Fournier_2013]: these criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding value added tax).
+Because the DFG has enforced a set of criteria grantees have to comply with, similar implementations for supporting open access publishing across German universities exists [@Fournier_2013].
+These criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding value added tax)[^3].
 Grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
@@
 my_apc$journal_full_title <- gsub("PLoS", "PLOS", my_apc$journal_full_title)
-my_apc$journal_full_title <- gsub("PLOS ONE", "PLOS ONE", my_apc$journal_full_title)
+my_apc$journal_full_title <- gsub("Scientific Repor.", "Scientific Reports", my_apc$journal_full_title)
 
@@
 
-On May 13, 2016,[^16a] the Open APC initiative covered `r nrow(my_apc)` articles whose publication fees were centrally paid by `r length(unique(my_apc$institution))` German universities and research institutions until 2015.
+After excluding payments being made for non-journal articles and in 2016, we retrieved `r nrow(my_apc)` articles for which the publication fee was centrally paid by `r length(unique(my_apc$institution))` German universities and research institutions until 2015.
 The number of supported open access journal articles grew over the years (see Figure 1).
@@
 
-```{r articles_year, fig.width=6, fig.height=2.5,fig.cap="Growth of Open APC Initiative"}
+**Figure 1: Growth of Open APC Initiative**
+
+```{r articles_year, results = 'hide'}
 apc_time <- my_apc %>%
@@
     text = element_text(size = 14))
-
+pdf("figure_1.pdf", height=4, width=6) 
 ggplot(apc_time, aes(factor(period), cumsum(articles), group = 1)) +
@@
   theme_bw() + science_theme
-ggsave("figure_1.pdf", height=4, width=6)
+dev.off()
 
@@
 
-```{r histm, fig.width=6, fig.height=2.5,fig.cap="Histogram APC"}
+```{r histm, results = 'hide'}
 require(ggplot2)
@@
     text = element_text(size = 14))
-
+pdf("figure_2.pdf", height=4, width=6)
 ggplot(my_apc, aes(euro)) + 
@@
   theme_bw() + science_theme
-
+dev.off()
 ```
@@
 Among all articles, fees amounted to `r sum(my_apc$euro)` € including VAT, the average payment was `r mean(my_apc$euro)` € and the median value `r median(my_apc$euro)` € .
-Figure 2 shows the large price variation among the articles.
+
+**Figure 2: Distribution of institutional spending on publication fees by German research organisations**
+
+**Figure 3: Institutional spending on publication fees by German research organisations per year**
+
+Figure 3 presents the large price variation among the articles.
 The disclosed publication fees ranged from `r min(my_apc$euro)` € to `r max(my_apc$euro)` €.
@@
 
-```{r payments_year, fig.width=6, fig.height=3,fig.cap="Payments per year"}
+```{r payments_year, results = 'hide'}
+pdf("figure_3.pdf", height=4, width=6)
 ggplot(my_apc, aes(factor(period), euro)) +
@@
   theme_bw() + science_theme
-ggsave("figure_2.pdf", height=4, width=6)
-
+dev.off()
 ```
 
-The number of APC payments per institutions varied considerably (see Table 3).
+The number of APC payments per institutions varied considerably (see Table 1).
 With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall submissions.
-In contrast, the two universities of technology, TU Clausthal and TU Ilmenau, who recently begun to set up support structures for fee-based open access journal articles, shared payments made for four articles each.
+In contrast, the  universities of technology, TU Clausthal, TU Dortmund TU Ilmenau that recently begun to set up support structures for fee-based open access journal articles, self-reported low numbers of payments.
 
-```{r results='asis'}
+**Table 1: Institutional spending on open access publications (in €)**
+
+```{r}
 tt <- my_apc %>%
   group_by(institution) %>%
-  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), Median = median(euro), Min = min(euro), Max = max(euro)) %>%
+  summarise(Articles = n(), Total = sum(euro), Mean = mean(euro), SD = sd(euro), Median = median(euro), Min = min(euro), Max = max(euro)) %>%
+  mutate(range_apc = paste(Min, "--", Max)) %>%
+  select(-Min, -Max) %>%
   arrange(desc(Articles))
 
-colnames(tt) <- c("Institutions", "Articles funded", "Total", "Mean", "Median", "Minimum", "Maximum")
-tt <- format(tt,big.mark=',')
-pander::pandoc.table(tt, split.cells = 7, caption = 'Institutions self-reporting expenditures for open access publications (in €)', justify = 'lrrrrrr')
+colnames(tt) <- c("Institution", "Articles funded", "Total", "Mean", "SD", "Median", "Min-Max")
+table_1 <- format(tt,big.mark=',')
+#table_1 <- pander::pandoc.table(tt, split.cells = 7, caption = 'Publication fees paid per publisher (in €)', justify = 'lrrrrrr')
+
 ```
@@
 
-```{r results='asis'}
+```{r results='hide'}
 tt <- read.csv("data/aggregates_fees_paid.csv", header = TRUE, sep =",")
 colnames(tt) <- c("Cost data-set", "Journal Type", "Articles funded", "Total costs in Euro", "Mean")
-pander::pandoc.table(tt,split.cells = 5 , big.mark = ",", caption = 'Comparision of cost data per period and journal type  (in €)', justify = 'lrrrr')
+table_2 <- tt
+#table_2 <- pander::pandoc.table(tt,split.cells = 5 , big.mark = ",", caption = 'Comparision of cost data per period and journal type  (in €)', justify = 'lrrrr')
 ```
@@
 
-Along with the price information, participating institutions were required to identify funded articles by their DOI.
-They were reported for `r my_apc %>% filter(!is.na(doi)) %>% nrow()` out of `r nrow(my_apc)` articles.
-Of those, `my_apc %>% filter(indexed_in_crossref == TRUE) %>% nrow()` were indexed in CrossRef, representing `my_apc %>% filter(indexed_in_crossref == TRUE) %>% nrow() / my_apc %>% nrow() * 100`  % of all funded publications.
-The reasons why articles identified by a DOI were not registered with CrossRef differed.
-Some journals were not indexed by CrossRef at the time of our study but by the DOI agencies DataCite (Journal of new frontiers in spatial concepts published by KIT Scientific Publishing) and Medra (DIE ERDE: Journal of the Geographical Society of Berlin).
-In other cases, either the DOI did not refer to the full text despite the fact that the journal was indexed on a regular basis (compare <http://doi.org/10.1186/1471-2105-13-S19-S7> with <http://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-13-S19-S7>) or the resource type could not be retrieved, although the DOI resolves (<http://doi.org/10.1186/s12885-015-1795-7>).
+```{r}
+doi_miss <- jsonlite::stream_in(file("data/missing_dois_agency.json"), verbose = FALSE)
+doi_miss[doi_miss$doi_agency == "", "doi_ageny"] <- NA
+doi_miss <- doi_miss[!doi_miss$institution == "FWF - Austrian Science Fund",]
+
+cr <- my_apc %>% 
+  filter(!doi %in% doi_miss$doi) %>%
+  filter(doi = !is.na(doi))
+
+agencies <- doi_miss %>% 
+  group_by(doi_agency) %>% 
+  summarise(n()) %>%
+  rbind(c("CrossRef", nrow(cr)),.)
+
+agencies <- data.frame(table(unlist(doi_miss$doi_agency)))
+
+```
+
+To identify the sponsored open access journal articles, as well as to gather bibliographic metadata, DOIs were a mandatory part of the Open APC initiative's data collection activities.
+The participating institutions reported DOIs for `r my_apc %>% filter(!is.na(doi)) %>% nrow()`  out of `r nrow(my_apc)` articles.
+using these DOIs, we retrieved additional metadata from CrossRef for `r nrow(cr)` publications,  representing `r nrow(cr) / nrow(my_apc) * 100`  % of the total article volume.
+Articles for which no metadata could be obtained, were either registered with the DOI agency DataCite (10 articles) or Medra (two articles).
+For eight articles, our parser could not gather the XML resource, although these publications were registered with CrossRef.
+Seven DOIs reported to the Open APC initiative did not resolve.
 
@@
 
+**Table 2: Publication fees paid per publisher (in €)**
+
 ```{r results='asis'}
@@
 colnames(tt) <- c("Journal", "Articles funded", "Total", "Mean", "SD", "Median", "Range")
-tt <- format(tt,big.mark=',')
-pander::pandoc.table(tt, split.cells = 7, caption = 'Publication fees paid per publisher (in €)', justify = 'lrrrrrr')
+table_2 <- format(tt,big.mark=',')
+# pander::pandoc.table(tt, split.cells = 7, caption = 'Publication fees paid per publisher (in €)', justify = 'lrrrrrr')
+
+## springer nature merger
+springer_nat <- my_apc %>% 
+  filter(publisher == "Springer Nature")
+
+springer <- springer_nat %>% filter(member_id == "297")
+nature <- springer_nat %>% filter(member_id == "339")
 ```
 
-Most of the funding of publication fees in Germany went to the publisher Springer Science + Business Media, especially profiting from the merge with the former full open access publisher BioMed Central.
-In contrast, other established publishing houses such as Elsevier and Wiley-Blackwell rank lower, presumably because they mostly publish hybrid journals, which were not well represented in our data-set at the time of the study.
-Table 4 also illustrates the variation across and within publishers, which confirms earlier findings [@Pinfield_2015].
+Most of the funding of publication fees in Germany went to the publisher Springer Nature, most llikeky profiting from the merge with the open access publisher BioMed Central in 2008, and in 2015 between the well-established publishers Springer Science + Business Media and Nature Publishing Group.
+Using the CrossRef-Member-ID instead of the publisher name, we were able to differentiate between journals  formely published by Springer Science + Business Media and Nature Publishing Group.
+In terms of articles, the majority of payments made were for publications in journal that were formerly associated with Springer Science + Business Media.
+Springer Science + Business Media journals accounted for `r nrow(springer)`articles, representing `r nrow(springer) / nrow(springer_nat) * 100` of the total Springer Nature article volume in the spending data, and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % of the amount spent.
+Median publication fee spending slightly differed between Springer Science + Business Media (`r median(springer$euro)` €) and Nature Publishing (`r median(nature$euro)` €).
+However, price variation was higher for Nature Publishing journals (SD = `r sd(nature$euro)`€) compared with that of the former Springer Science + Business Media titles (SD = `r sd(nature$euro)` €).
+
+In contrast to Springer Nature, other well-established publishing houses such as Elsevier and Wiley-Blackwell rank lower in our analysis, presumably because they mostly published hybrid journals, a journal type that was not well represented in our data-set at the time of this study.
+Table 4 also illustrates the variation across and within publishers, which confirms earlier findings [@Solomon_2012; @Pinfield_2015].
+
+**Table 3: Publication fees paid per journal (in €)**
 
@@
 colnames(tt) <- c("Journal", "Articles funded", "Total", "Mean", "SD", "Median", "Range")
-tt <- format(tt,big.mark=',')
-pander::pandoc.table(tt, split.cells = 7, caption = 'Publication fees paid per journal (in €)', justify = 'lrrrrrr')
+table_3 <- format(tt,big.mark=',')
+#pander::pandoc.table(tt, split.cells = 7, caption = 'Publication fees paid per journal (in €)', justify = 'lrrrrrr')
 ```

### 2016-06-03 article.Rmd
finishing the work for today

@@
 
-One reason why the study of open access journals using publication fees is in most cases difficult is that total spending on publication fees is fragmented across the budgets of grant agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
+One reason why the study of how much was spent is in most cases difficult is that total publication fee spending is fragmented across the budgets of grant agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
 Asking 9,645 authors from various disciplines how they financed publication fees, a comprehensive survey in 2010 revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.
@@
 
-Another key problem for the study of open access is that funding for open access journals using publication fees lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
-However, empirical studies examining publication fees gathered from journal websites report similar findings.
-While the average publication fee reported ranged between 904 $ [@Solomon_2012] and 923 $ [@Walters_2011], considerable price variances across journals and publishers were observed.
+Another key problem in this regard is that funding for open access journals using publication fees lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
+Empirical studies examining publication fees so far gathered price information from author surveys [@soap], or journal websites.
+Using the latter method, two studies investigating journals across multiple disciplines observed similar averages that ranged between 904 $ [@Solomon_2012] and 923 $ [@Walters_2011], as well as considerable price variances across journals and publishers.
 Accordingly, @Solomon_2012 suggested to cluster fully open access journal using publication fees into several groups.
-In descending order, high-impact journals charged most, followed by biomedicine journals from commercial publishers, large multi-disciplinary journals, and mid-price ranged journals from commercial publishers covering a large spectrum of disciplines.
+In descending order, these are high-impact journals, followed by biomedicine journals from commercial publishers, large multi-disciplinary journals, and mid-price ranged journals from commercial publishers covering a large spectrum of disciplines.
 Lower priced journals were published by academic societies and by publishers from low-income countries.
 
-It remains unclear which other factors contribute to this varying pricing.
+Nevertheless, it remains unclear which factors contribute to pricing in academic publishing.
 Generally, these might include article processing, impact, rejection rates, management and investment, and profit margins [@Van_Noorden_2013].
-While fixed prices for individual articles are common, agreements between publishers and institutions often provide discounts, and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
+While fixed prices for individual articles are common, agreements between publishers and institutions can lead to discounts, and publishers sometimes waive publication fees for authors from low-income countries [@BJ_RK_2012; @Lawson_waiver_2015].
 Other factors leading to variable pricing schemes include submission or page charges [@BJ_RK_2012].
@@
 These journals, allowing articles to be published immediately as open access after a charge was paid, rely both on subscriptions and publication fees as revenue sources.
-Although the uptake of open access through hybrid journals was described as lower and more expensive compared to that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014], this model has gained attention through recent science policies, notably because of open access policies from the UK [@Pinfield_Review_2015].
+Although the uptake of open access through hybrid journals was described as lower and more expensive compared to that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014; @Pinfield_2015], this model has gained attention through recent science policies, notably because of open access policies from the UK [@Pinfield_Review_2015].
 
-To address the problems of fragmented spending on publication fees and in-transparency about what was paid, some European research funders and research performing institutions have recently begun to disclose their spending on publication fees as open data.
+To address these problems of fragmented spending on publication fees and of in-transparency about what was paid, some European research funders and research performing institutions have recently begun to disclose their payment for publication as open data.
 To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc_12] and the Austrian Science Fund FWF [@fwf_apc_13].
 The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
-Disclosed as publicly available spreadsheets, these data-sets self-report payments made along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
-Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates [@woodward_2014].
-Parts of Jisc cost data was examined by @Pinfield_2015.
-Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as a varying number of articles UK universities supported between 2007 - 2014.
+Disclosed as publicly available spreadsheets, these data-sets self-report what was being spent along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
+Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates [@Neylon_2014; @woodward_2014].
+Parts of Jisc's cost data was examined by @Pinfield_2015.
+Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as a varying number of articles UK universities supported between 2007 - 2014, confirming the earlier studies based on price information from journal websites.
 
@@
 
-This paper focuses on how much German universities and research organization spent on open access publication fees.
+This paper focuses on how much German universities and research organisation spent on open access publication fees.
 In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced how universities manage institutional support for publication fees.[^3] Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" program  in 2011, and similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012], only few central funds existed [@eppelin_2012].
@@
 Grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
-They include central invoicing schemes and related agreements between university libraries and publishers [@Fournier_2013].
+They include central invoicing schemes and related agreements between university libraries and publishers, leading to discounts for the authors [@Fournier_2013].
 
-Research institutes organized in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this funding program, contributing to the diversity of schemes in Germany.
+Research institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this DFG funding program, contributing to the diversity of schemes in Germany.
 But in response, some organizations have adopted similar processes to support authors.
@@
 
-The evolving institutional support structures to cover open access publication fees has led to calls for an unified approach towards supporting open access journal publishing.
-The Allianz der Wissenschaftsorganisationen[^10c], a science policy board representing all major research organizations in Germany, marks cost transparency as one way to sustain an "adequate open access publication system" [@allianz].
-Reflecting Austrian and UK initiatives to share institutional spending on open access publication fees as open data, as well as professional discussions on open access publishing, Bielefeld University Library began to openly share its payments in May 2014.
+The evolving institutional support structures to cover open access publication fees has led to calls for an unified approach towards supporting open access journal publishing in Germany.
+The Allianz der Wissenschaftsorganisationen[^10c], a science policy board representing all major research organizations in Germany, marked cost transparency as one way to sustain an "adequate open access publication system" [@allianz].
+Reflecting Austrian and UK initiatives to share institutional spending on open access publication fees as open data, as well as professional discussions on open access publishing, Bielefeld University Library began to openly share its payment of publication fees in May 2014.
 After engaging with the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11] other German institutions joined under the umbrella of the Open APC initiative soon after.
@@
 The aim of the study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-Using self-reported cost data from the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and UK initiatives.
+Using self-reported cost data from the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these expenditure data with data from related Austrian and UK initiatives, both in terms of size and proportion of the different journal types fully and hybrid open access journal.
 We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.

### 2016-06-04 article.Rmd
replaced CrossRef with Crossref after their rebranding

@@
   only 53 articles in hybrid journals. With an indexing coverage of 99 %, the findings
-  reveal that the DOI agency CrossRef provides both comprehensive bibliographic coverage
+  reveal that the DOI agency Crossref provides both comprehensive bibliographic coverage
   of the funded open access journal literature and disambiguated names of journal
@@
 This study contributes to a growing empirical base about institutional spending on open access publication fees.
-More specifically, it uses self-reported cost data from German universities and research organisations to examine how much they spent, and how the so funded open access journal articles were distributed over publishers and journal titles using bibliographic information from CrossRef.
+More specifically, it uses self-reported cost data from German universities and research organisations to examine how much they spent, and how the so funded open access journal articles were distributed over publishers and journal titles using bibliographic information from Crossref.
 Such an approach extends methods and improves data collection activities for researchers and practitioners alike, as well as contribute to a better understanding of factors affecting the analysis of publication fees in open access publishing.
@@
 Using self-reported cost data from the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these expenditure data with data from related Austrian and UK initiatives, both in terms of size and proportion of the different journal types fully and hybrid open access journal.
-We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.
+We also asked how thoroughly self-reported articles were indexed in Crossref, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.
 
@@
 
-We fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOIs.
-Although the Open APC initiative gathered metadata representing publishers and journals from CrossRef as well, this information was retrieved at the time when the participating institutions submitted the cost information.
+We fetched bibliographic metadata for each article from Crossref on May 19, 2016, on the basis of the reported DOIs.
+Although the Open APC initiative gathered metadata representing publishers and journals from Crossref as well, this information was retrieved at the time when the participating institutions submitted the cost information.
 The Open APC initiative tracked the date of data submission and the contributors for every data submission with Git, a version control system, increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub to be transparent over time.
-However, CrossRef regularly updated metadata to represent ongoing mergers of publishing houses or name changes.
+However, Crossref regularly updated metadata to represent ongoing mergers of publishing houses or name changes.
 A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that operated as Springer Nature at the time of our study.
-To reflect these dynamics in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set instead of re-using publisher and journal information contained in the Open APC data-set.
+To reflect these dynamics in academic publishing, we decided to retrieve updated metadata from Crossref for the whole Open APC data-set instead of re-using publisher and journal information contained in the Open APC data-set.
 
-As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access CrossRef's REST API.
+As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access Crossref's REST API.
 We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the ISSN media types, the International Standard Serial Number used to identify journals, were distinguished.
@@
 In cases where no bibliographic information could be obtained, we used the Open APC values.
-Because CrossRef was not the only registration agency for DOIs, but also the agencies DataCite and Medra minted DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
+Because Crossref was not the only registration agency for DOIs, but also the agencies DataCite and Medra minted DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
 
@@
 Our first screening of the data-sets revealed that some articles lacked a DOI.
-For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided only to compare German spending with that of other initiatives, but not its size and distribution over publisher and journal titles, as well as the indexing coverage in CrossRef.
+For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided only to compare German spending with that of other initiatives, but not its size and distribution over publisher and journal titles, as well as the indexing coverage in Crossref.
 
@@
 
-## CrossRef indexing{.unnumbered}
+## Crossref indexing{.unnumbered}
 
@@
   summarise(n()) %>%
-  rbind(c("CrossRef", nrow(cr)),.)
+  rbind(c("Crossref", nrow(cr)),.)
 
@@
 The participating institutions reported DOIs for `r my_apc %>% filter(!is.na(doi)) %>% nrow()`  out of `r nrow(my_apc)` articles.
-using these DOIs, we retrieved additional metadata from CrossRef for `r nrow(cr)` publications,  representing `r nrow(cr) / nrow(my_apc) * 100`  % of the total article volume.
+using these DOIs, we retrieved additional metadata from Crossref for `r nrow(cr)` publications,  representing `r nrow(cr) / nrow(my_apc) * 100`  % of the total article volume.
 Articles for which no metadata could be obtained, were either registered with the DOI agency DataCite (10 articles) or Medra (two articles).
-For eight articles, our parser could not gather the XML resource, although these publications were registered with CrossRef.
+For eight articles, our parser could not gather the XML resource, although these publications were registered with Crossref.
 Seven DOIs reported to the Open APC initiative did not resolve.
@@
 
-We used the DOI to automatically fetch publisher and journal names for each article from the CrossRef REST API.
+We used the DOI to automatically fetch publisher and journal names for each article from the Crossref REST API.
 Table 4 shows the top ten publishers in terms of payments made that represent `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the spending for publication fees.
@@
 Most of the funding of publication fees in Germany went to the publisher Springer Nature, most llikeky profiting from the merge with the open access publisher BioMed Central in 2008, and in 2015 between the well-established publishers Springer Science + Business Media and Nature Publishing Group.
-Using the CrossRef-Member-ID instead of the publisher name, we were able to differentiate between journals  formely published by Springer Science + Business Media and Nature Publishing Group.
+Using the Crossref-Member-ID instead of the publisher name, we were able to differentiate between journals  formely published by Springer Science + Business Media and Nature Publishing Group.
 In terms of articles, the majority of payments made were for publications in journal that were formerly associated with Springer Science + Business Media.
@@
 
-In our study, CrossRef thoroughly indexed open access journal articles in the Open APC data-set.
+In our study, Crossref thoroughly indexed open access journal articles in the Open APC data-set.
 We could gather metadata representing publisher and journal titles for 99 % articles, and successfully merged these information with the Open APC cost data.
-Using metadata from CrossRef, therefore, reduces the extensive validation work of bibliographic information provided that the reporting of the DOI along with the expenses is made mandatory.
-Drawing on CrossRef would also increase the comparability of cost data for future negotiations with publishers on open access agreements, and the open access spending between open data initiatives that apply the same reporting standards, as its metadata represent the dynamic landscape of academic publishing in terms of ongoing mergers of publishing houses or name changes.
+Using metadata from Crossref, therefore, reduces the extensive validation work of bibliographic information provided that the reporting of the DOI along with the expenses is made mandatory.
+Drawing on Crossref would also increase the comparability of cost data for future negotiations with publishers on open access agreements, and the open access spending between open data initiatives that apply the same reporting standards, as its metadata represent the dynamic landscape of academic publishing in terms of ongoing mergers of publishing houses or name changes.
 
@@
 Starting with a few institutions, the integration of further data contributors has not been a problem.
-While we use article metadata reported by publishers to CrossRef, we think it's best to collect information on how much has been spent at the sources: the institutions paying for APCs.
+While we use article metadata reported by publishers to Crossref, we think it's best to collect information on how much has been spent at the sources: the institutions paying for APCs.
 We look forward to more data contributors to the initiative and to adapt this scheme to new analyses, additional data, or international sources.
@@
 
-[^19]: We would like to thank Martin Fenner for pointing this out to us. See also the CrossRef API documentation <https://github.com/CrossRef/rest-api-doc/blob/master/rest_api.md>
+[^19]: We would like to thank Martin Fenner for pointing this out to us. See also the Crossref API documentation <https://github.com/CrossRef/rest-api-doc/blob/master/rest_api.md>
 

### 2016-06-04 article.Rmd
text review

@@
 
-Research institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this DFG funding program, contributing to the diversity of schemes in Germany.
+Research institutes organized in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this DFG funding program, contributing to the diversity of schemes in Germany.
 But in response, some organizations have adopted similar processes to support authors.
@@
 We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the ISSN media types, the International Standard Serial Number used to identify journals, were distinguished.
-It also contained normalised publisher information, thus avoiding confusion about naming of publishing houses other studies were faced with when working with self-reported data [@woodward_2014].
+It also contained normalized publisher information, thus avoiding confusion about naming of publishing houses other studies were faced with when working with self-reported data [@woodward_2014].
 In cases where no bibliographic information could be obtained, we used the Open APC values.
@@
 Because Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
-We excluded articles from the analysis for which neither price information nor the journal type could be obtained.
-In the case of Jisc's 2014 data [@jisc_14], 4.861.772 € being spent on 2,812 publications were not eligible for further analysis because no publication type was given in the data-set.
+We excluded articles from the analysis for which either cost information or  journal type was missing.
+In the case of Jisc's 2014 data [@jisc_14], we could not analyze 4.861.772 € spent on 2,812 publications because no publication type was given in the data-set.
 
-Data collection methods of the Open APC initiative and those of the others differed in some aspects.
-For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institution were required to report, publication identifier in the Wellcome Trust data was also added through automated compliance checks.
+The data collection methods of the Open APC initiative and those of the other initiatives differ in some aspects.
+For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, in the Wellcome Trust data a publication identifier was added as part of the automated compliance checks.
 Our first screening of the data-sets revealed that some articles lacked a DOI.
-For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided only to compare German spending with that of other initiatives, but not its size and distribution over publisher and journal titles, as well as the indexing coverage in Crossref.
+For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided to only compare German APC spending with that reported by other initiatives.
+We did not compare the overall size, the distribution over publishers and journal titles, or the indexing coverage in Crossref.
 

### 2016-06-04 article.Rmd
whitespace
€, Euro, EUR -> €

@@
 In the case of FWF, we gathered the cost information from the accompanying spending reports.
-We used the spreadsheet data to calculate  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.
+We used the spreadsheet data to calculate Wellcome Trust's and Jisc's spending, and converted the prices from GBP to € in accordance with the average € foreign exchange reference rates provided by the European Central Bank.
 Our comparison between the open data initiatives focussed on the last two years 2014 and 2015.
@@
   scale_x_continuous(limits = c(0,8000)) +
-  xlab("Publication fee in Euro") +
+  xlab("Publication fee in €") +
   ylab("Articles") +
@@
   xlab("Year") +
-  ylab("Publication fee paid in EURO") +
+  ylab("Publication fee paid in €") +
   scale_y_continuous(limits = c(0,8000)) + 
@@
 Table 2 compares Open APC spending data with that of the Austrian FWF, as well as with Jisc's and Wellcome Trust's expenditures.
-Prices were converted according to the average Euro exchange rate of the examined periods, and gathered for both fully and hybrid open access journals.
+Prices were converted according to the average € exchange rate of the examined periods, and gathered for both fully and hybrid open access journals.
 The comparison reveals that the Open APC initiative lacked cost information about hybrid journals, whereas the related Austrian and UK open data initiatives could report a large share of spending for these journals between 2014 and 2015.
@@
 tt <- read.csv("data/aggregates_fees_paid.csv", header = TRUE, sep =",")
-colnames(tt) <- c("Cost data-set", "Journal Type", "Articles funded", "Total costs in Euro", "Mean")
+colnames(tt) <- c("Cost data-set", "Journal Type", "Articles funded", "Total costs in €", "Mean")
 table_2 <- tt
@@
 
-[^3a]: We are sure that some universities with DFG-funded publication funds pay for APCs higher than 2,000 EUR.
+[^3a]: We are sure that some universities with DFG-funded publication funds pay for APCs higher than 2,000 €.
 They have to make sure they don't use the funder's money for these articles, and this maybe explains why some institutions are reluctant to report on these APCs.
-Scatter plots of all APC payments show that there are only very few APCs over 2,000 EUR.
+Scatter plots of all APC payments show that there are only very few APCs over 2,000 €.
 

### 2016-06-04 article.Rmd
minor fixes

@@
 
-One reason why the study of how much was spent is in most cases difficult is that total publication fee spending is fragmented across the budgets of grant agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
+One reason why the study on how much was being spent is in most cases difficult is that total publication fee spending is fragmented across the budgets of grant agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
 Asking 9,645 authors from various disciplines how they financed publication fees, a comprehensive survey in 2010 revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.
@@
 Another key problem in this regard is that funding for open access journals using publication fees lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
-Empirical studies examining publication fees so far gathered price information from author surveys [@soap], or journal websites.
-Using the latter method, two studies investigating journals across multiple disciplines observed similar averages that ranged between 904 $ [@Solomon_2012] and 923 $ [@Walters_2011], as well as considerable price variances across journals and publishers.
+Empirical studies examining publication fees so far obtained price estimates by surveying authors [@soap], or from journal websites.
+Using the latter method, two studies investigating journals across a broad range of disciplines calculated similar price averages that ranged between 904 $ [@Solomon_2012] and 923 $ [@Walters_2011], as well as considerable price variances across journals and publishers.
 Accordingly, @Solomon_2012 suggested to cluster fully open access journal using publication fees into several groups.
-In descending order, these are high-impact journals, followed by biomedicine journals from commercial publishers, large multi-disciplinary journals, and mid-price ranged journals from commercial publishers covering a large spectrum of disciplines.
+In descending order, these are high-impact journals, followed by biomedicine journals from commercial publishers, large multi-disciplinary journals, and mid-price journals from commercial publishers covering a large spectrum of disciplines.
 Lower priced journals were published by academic societies and by publishers from low-income countries.
@@
 
-Hybrid journals substantially add to this complexity of open access funding [@Kingsley_2014].
+Hybrid journals substantially add to this complexity of open access funding [@Kingsley_2014; @Pinfield_2015].
 These journals, allowing articles to be published immediately as open access after a charge was paid, rely both on subscriptions and publication fees as revenue sources.
-Although the uptake of open access through hybrid journals was described as lower and more expensive compared to that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014; @Pinfield_2015], this model has gained attention through recent science policies, notably because of open access policies from the UK [@Pinfield_Review_2015].
+Although the uptake of open access through hybrid journals was described as lower and more expensive compared to that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014], this model has gained attention through recent science policies, notably because of open access policies from the UK [@Pinfield_Review_2015].
 
-To address these problems of fragmented spending on publication fees and of in-transparency about what was paid, some European research funders and research performing institutions have recently begun to disclose their payment for publication as open data.
+To address these problems of fragmented spending on publication fees and of in-transparency about what was being paid, some European research funders and research performing institutions have recently begun to disclose their payment for publication as open data.
 To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc_12] and the Austrian Science Fund FWF [@fwf_apc_13].
@@
 Disclosed as publicly available spreadsheets, these data-sets self-report what was being spent along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
-Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates [@Neylon_2014; @woodward_2014].
+Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers to the full text [@Neylon_2014; @woodward_2014].
 Parts of Jisc's cost data was examined by @Pinfield_2015.
-Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as a varying number of articles UK universities supported between 2007 - 2014, confirming the earlier studies based on price information from journal websites.
+Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as a varying number of articles UK universities supported between 2007 - 2014, confirming earlier studies that collected price information from journal websites.
 
@@
 Grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
-They include central invoicing schemes and related agreements between university libraries and publishers, leading to discounts for the authors [@Fournier_2013].
+This includes central invoicing schemes, and related agreements between university libraries and publishers, leading to discounts for the authors [@Fournier_2013].
 
-Research institutes organized in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this DFG funding program, contributing to the diversity of schemes in Germany.
+Research institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this DFG funding program.
 But in response, some organizations have adopted similar processes to support authors.
@@
 The evolving institutional support structures to cover open access publication fees has led to calls for an unified approach towards supporting open access journal publishing in Germany.
-The Allianz der Wissenschaftsorganisationen[^10c], a science policy board representing all major research organizations in Germany, marked cost transparency as one way to sustain an "adequate open access publication system" [@allianz].
+The Allianz der Wissenschaftsorganisationen[^10c], a science policy board representing all major research organisations in Germany, marked price transparency as one way to sustain an "adequate open access publication system" [@allianz].
 Reflecting Austrian and UK initiatives to share institutional spending on open access publication fees as open data, as well as professional discussions on open access publishing, Bielefeld University Library began to openly share its payment of publication fees in May 2014.
@@
 The aim of the study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-Using self-reported cost data from the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these expenditure data with data from related Austrian and UK initiatives, both in terms of size and proportion of the different journal types fully and hybrid open access journal.
+Using self-reported cost data from the Open APC initiative, the analysis focused on the amount that was being spent on publication fees, and compared these expenditure with data from related Austrian and UK initiatives, both in terms of size and the proportion of articles being published in fully and hybrid open access journals.
 We also asked how thoroughly self-reported articles were indexed in Crossref, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.
@@
 
-We analysed self-reported cost data released by the Open APC initiative on May 13, 2016, to assess institutional spending on open access publication fees in Germany.
+We analysed self-reported cost data released by the Open APC initiative on May 13, 2016[^24] to assess institutional spending on open access publication fees in Germany.
 In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully or hybrid open access journal as well as the recorded DOI from the data-set.
@@
 We fetched bibliographic metadata for each article from Crossref on May 19, 2016, on the basis of the reported DOIs.
-Although the Open APC initiative gathered metadata representing publishers and journals from Crossref as well, this information was retrieved at the time when the participating institutions submitted the cost information.
-The Open APC initiative tracked the date of data submission and the contributors for every data submission with Git, a version control system, increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub to be transparent over time.
-However, Crossref regularly updated metadata to represent ongoing mergers of publishing houses or name changes.
+Although the Open APC initiative gathered metadata representing publishers and journals from Crossref as well, this information was retrieved at the time when the participating institutions submitted their expenses.
+The Open APC initiative kept track of the date when these data-sets were submitted with Git, a version control system, increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub to be transparent over time.
+However, during these data collection activities Crossref had regularly updated metadata to represent ongoing mergers of publishing houses.
 A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that operated as Springer Nature at the time of our study.
-To reflect these dynamics in academic publishing, we decided to retrieve updated metadata from Crossref for the whole Open APC data-set instead of re-using publisher and journal information contained in the Open APC data-set.
+To reflect these dynamics in academic publishing, we decided to retrieve updated metadata from Crossref for the whole Open APC data-set instead of re-using the historical publisher and journal information contained in the Open APC data-set.
 
-As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access Crossref's REST API.
+As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access Crossref's REST API[^25].
 We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the ISSN media types, the International Standard Serial Number used to identify journals, were distinguished.
-It also contained normalized publisher information, thus avoiding confusion about naming of publishing houses other studies were faced with when working with self-reported data [@woodward_2014].
+It also contained normalised publisher information, thus avoiding confusion about naming of publishing houses other studies were faced with when working with self-reported data [@woodward_2014].
 In cases where no bibliographic information could be obtained, we used the Open APC values.
@@
 Data collection also involved obtaining cost data from related open data initiatives.
-To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we reviewed the openly available data-sets from the the Austrian Science Fund FWF [@fwf_apc_14; @fwf_apc_15], Jisc [@jisc_14; @jisc_15] and the Wellcome Trust [@wellcome_apc_13_14; @wellcome_apc_14_15].
-For analysis, we obtained the overall publication fee spending to support fully and hybrid open access journal articles.
+To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we consulted the openly available data-sets from the the Austrian Science Fund FWF [@fwf_apc_14; @fwf_apc_15], Jisc [@jisc_14; @jisc_15] and the Wellcome Trust [@wellcome_apc_13_14; @wellcome_apc_14_15].
+For analysis, we obtained the overall publication fee spending on both fully and hybrid open access journal articles.
 In the case of FWF, we gathered the cost information from the accompanying spending reports.
@@
 Our comparison between the open data initiatives focussed on the last two years 2014 and 2015.
-Because Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
+Because Wellcome Trust's spending was reported for the financial periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
 We excluded articles from the analysis for which either cost information or journal type was missing.
-In the case of Jisc's 2014 data [@jisc_14], we could not analyze 4.861.772 € spent on 2,812 publications because no publication type was given in the data-set.
+In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to 4.861.772 € from the analysis because no publication type was given in the data-set.
 
-The data collection methods of the Open APC initiative and those of the other initiatives differ in some aspects.
+The data collection methods of the Open APC initiative and those of the other initiatives differed in some aspects.
 For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, in the Wellcome Trust data a publication identifier was added as part of the automated compliance checks.
 Our first screening of the data-sets revealed that some articles lacked a DOI.
-For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided to only compare German APC spending with that reported by other initiatives.
-We did not compare the overall size, the distribution over publishers and journal titles, or the indexing coverage in Crossref.
+For this reason, and as our main focus was institutional funding for publication fees in Germany, we decided to only compare German APC spending with that reported by other initiatives.
+We did not compare the distribution of spending over publishers and journal titles, or the indexing coverage in Crossref.
 
@@
 
-After excluding payments being made for non-journal articles and in 2016, we retrieved `r nrow(my_apc)` articles for which the publication fee was centrally paid by `r length(unique(my_apc$institution))` German universities and research institutions until 2015.
-The number of supported open access journal articles grew over the years (see Figure 1).
-While one institution disclosed `r filter(my_apc, period == min(my_apc$period)) %>% nrow()` payments made in `r as.character(min(my_apc$period))`, the majority shared their expenditures from 2013 onwards.
+After excluding payments being made for non-journal articles, as well as articles invoiced in 2016, we retrieved `r nrow(my_apc)` open access journal articles `r length(unique(my_apc$institution))` German universities and research institutions financially supported between 2005 and 2015.
+As illustrated in Figure 1, payments made for open access journal articles grew over the years.
+While one institution supported `r filter(my_apc, period == min(my_apc$period)) %>% nrow()` articles in `r as.character(min(my_apc$period))`, the majority shared their expenditure from 2013 onwards.
 With `r max(apc_time$articles)` articles, the year `r apc_time[apc_time$articles == max(apc_time$articles),"period"]` was best represented in our data-set.
-However, `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institutions contributed their cost data for 2015 at the time of this analysis, suggesting that there exits a time lag between payments made and reporting these spending to the Open APC initiative.
+However, the data-set contained `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institutions that contributed their cost data for 2015 at the time of this analysis, suggesting a time lag between payments made and reporting these spending to the Open APC initiative.
 
@@
 Among all articles, fees amounted to `r sum(my_apc$euro)` € including VAT, the average payment was `r mean(my_apc$euro)` € and the median value `r median(my_apc$euro)` € .
+Figure 2 presents the distribution of institutional spending on publications.
+We observed that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100`%) of the publication fees were paid in accordance with the DFG price cap of 2,000 €.
+Most payment on publications ranged between 1,000 - 1,250 €.
 
@@
 
-Figure 3 presents the large price variation among the articles.
-The disclosed publication fees ranged from `r min(my_apc$euro)` € to `r max(my_apc$euro)` €.
-However, the average price paid varied somewhat during the period 2011 and 2014 (`r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - `r my_apc %>% filter(period == 2014) %>% summarise(mean(euro)) %>% round()` €).
-We also observe that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100`%) of the publication fees were paid in accordance with the DFG price cap of 2,000 €.
+Figure 3 presents the large price variations.
+Publication fees that were paid by the German universities and research organisations ranged between `r min(my_apc$euro)` € and `r max(my_apc$euro)` €  (SD `r sd(my_apc$euro)`).
+However, the average price paid varied somewhat during the period 2011 and 2015 (`r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - `r my_apc %>% filter(period == 2015) %>% summarise(mean(euro)) %>% round()` €).
+Average publication fee spending per article in Germany was `r mean(my_apc$euro)` (Median = `r median(my_apc$euro)` €, SD = `r sd(my_apc$euro)` €).
 
@@
 With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall submissions.
-In contrast, the universities of technology, TU Clausthal, TU Dortmund TU Ilmenau that recently begun to set up support structures for fee-based open access journal articles, self-reported low numbers of payments.
+By contrast, the universities of technology, TU Clausthal, TU Dortmund TU Ilmenau that recently begun to set up support structures for fee-based open access journal articles, self-reported low numbers of payments.
 
@@
 
-Table 2 compares Open APC spending data with that of the Austrian FWF, as well as with Jisc's and Wellcome Trust's expenditures.
-Prices were converted according to the average € exchange rate of the examined periods, and gathered for both fully and hybrid open access journals.
-The comparison reveals that the Open APC initiative lacked cost information about hybrid journals, whereas the related Austrian and UK open data initiatives could report a large share of spending for these journals between 2014 and 2015.
-This situation likely reflected the DFG-inspired funding policy of most participating institutions that excluded support for publications in hybrid journals.
-Over the years 2005 – 2015, `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` German universities and research institutions reported `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` hybrid journal articles to the Open APC initiative, accounting for `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of the overall payments.
+Table 2 compares Open APC spending data with that of the Austrian FWF, as well as with Jisc's and Wellcome Trust's expenditure.
+Prices were converted according to the average Euro exchange rate of the examined periods, and gathered for both fully and hybrid open access journals.
+The comparison reveals that the Open APC initiative lacked cost information about hybrid journals, whereas the related Austrian and UK open data initiatives could report a large share of spending on these journals between 2014 and 2015.
+This situation presumably reflected the importance of the DFG as funding source for publication fees in Germany.
+The DFG funding policy excluded support for publications in hybrid journals.
+Over the years 2005 – 2015, `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` German universities and research institutions reported `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` hybrid journal articles to the Open APC initiative, representing `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of the overall supported article volume.
 In terms of the number of supported articles and the amount being spent on publication fees, by contrast, the Open APC data-set provided the most comprehensive price information for fully open access journals compared to what the Austrian and UK initiatives had reported.
@@
 Comparison of average prices suggests that publishing in hybrid journal was more expensive than in fully open access journals.
-Price differentials between these two categories were also reported earlier, concluding that prices for fully open access journals were generally lower [@Solomon_2012; @Pinfield_2015].
+Price differentials between these two categories were also reported earlier, indicating that prices for fully open access journals were typically lower [@Solomon_2012; @Pinfield_2015].
 In 2014 and 2015, the mean price for fully open access journals calculated from all data-sets was below the DFG price cap of 2,000 €.
@@
 
-To identify the sponsored open access journal articles, as well as to gather bibliographic metadata, DOIs were a mandatory part of the Open APC initiative's data collection activities.
+To identify publication fee spending on the article-level, as well as to gather bibliographic metadata, DOIs were a mandatory part of the Open APC initiative's data collection activities.
 The participating institutions reported DOIs for `r my_apc %>% filter(!is.na(doi)) %>% nrow()` out of `r nrow(my_apc)` articles.
-using these DOIs, we retrieved additional metadata from Crossref for `r nrow(cr)` publications, representing `r nrow(cr) / nrow(my_apc) * 100` % of the total article volume.
-Articles for which no metadata could be obtained, were either registered with the DOI agency DataCite (10 articles) or Medra (two articles).
-For eight articles, our parser could not gather the XML resource, although these publications were registered with Crossref.
+Using these DOIs, we retrieved additional metadata from Crossref for `r nrow(cr)` publications, representing `r nrow(cr) / nrow(my_apc) * 100` % of the total article volume.
+Articles for which no metadata could be obtained, were registered with the DOI agency DataCite (10 articles) or Medra (two articles).
+For eight articles, our parser could not gather the XML resource, although these publications were registered with Crossref at the time of our study.
 Seven DOIs reported to the Open APC initiative did not resolve.
@@
 We used the DOI to automatically fetch publisher and journal names for each article from the Crossref REST API.
-Table 4 shows the top ten publishers in terms of payments made that represent `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the spending for publication fees.
+Table 4 shows the top ten publishers in terms of payments made that represented `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall spending on publication fees.
 In total, payments were made to `r my_apc %>% group_by(publisher) %>% summarize(n()) %>% nrow()` publishing houses.
-In comparison with data from the UK, full open access publishers have a greater share on total spending.
+In comparison with data from the UK, some open access publishers have a greater share on total spending.
 @Pinfield_2015, for instance, reported remarkably lower numbers for the open access publishers MPDI AG, Copernicus GmbH, and Hindawi Publishing.
@@
 
-Most of the funding of publication fees in Germany went to the publisher Springer Nature, most llikeky profiting from the merge with the open access publisher BioMed Central in 2008, and in 2015 between the well-established publishers Springer Science + Business Media and Nature Publishing Group.
+Most of publication fee spending in Germany was on articles published in Springer Nature journals, most likely profiting from the merge with the open access publisher BioMed Central in 2008, and that between the well-established publishers Springer Science + Business Media and Nature Publishing Group in 2015.
 Using the Crossref-Member-ID instead of the publisher name, we were able to differentiate between journals formely published by Springer Science + Business Media and Nature Publishing Group.
-In terms of articles, the majority of payments made were for publications in journal that were formerly associated with Springer Science + Business Media.
-Springer Science + Business Media journals accounted for `r nrow(springer)`articles, representing `r nrow(springer) / nrow(springer_nat) * 100` of the total Springer Nature article volume in the spending data, and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % of the amount spent.
+In terms of articles, the majority of payments made were for publications in journals formerly associated with Springer Science + Business Media.
+Springer Science + Business Media journals accounted for `r nrow(springer)` articles, representing `r nrow(springer) / nrow(springer_nat) * 100`% of the overall Springer Nature article volume in the Open APC data, and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % in terms of the amount being spent.
 Median publication fee spending slightly differed between Springer Science + Business Media (`r median(springer$euro)` €) and Nature Publishing (`r median(nature$euro)` €).
-However, price variation was higher for Nature Publishing journals (SD = `r sd(nature$euro)`€) compared with that of the former Springer Science + Business Media titles (SD = `r sd(nature$euro)` €).
+However, price variation was higher for Nature Publishing journals (SD = `r sd(nature$euro)`€) compared with that of the former Springer Science + Business Media titles (SD = `r sd(springer$euro)` €).
 
-In contrast to Springer Nature, other well-established publishing houses such as Elsevier and Wiley-Blackwell rank lower in our analysis, presumably because they mostly published hybrid journals, a journal type that was not well represented in our data-set at the time of this study.
-Table 4 also illustrates the variation across and within publishers, which confirms earlier findings [@Solomon_2012; @Pinfield_2015].
+In contrast to Springer Nature, other well-established publishing houses such as Elsevier and Wiley-Blackwell rank lower in our analysis.
+Table 2, furthermore, illustrates the price variations across and within publishers, which confirms earlier findings [@Solomon_2012; @Pinfield_2015].
 
@@
 Prices also varied within single journals.
-Based on the number of articles paid for, Table 5 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals.
-Payments to these ten journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of all payments.
-In the case of Atmospheric Chemistry and Physics Discussions, the price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
+Based on the number of articles paid for, Table 3 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals.
+Payments to these journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall expenditure.
+In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
 
@@
 [^23]: <https://github.com/OpenAPC/openapc-de/wiki/Handreichung-Dateneingabe>
+
+[^24]: https://github.com/OpenAPC/openapc-de/tree/v2.4.3
+
+[^25]: <https://github.com/CrossRef/rest-api-doc/blob/master/rest_api.md>

### 2016-06-04 article.Rmd
retooled discussion, enough for today

@@
 To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc_12] and the Austrian Science Fund FWF [@fwf_apc_13].
-The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
+The not-for-profit company Jisc followed by collecting data from UK universities [@Lawson_data_2015].
 Disclosed as publicly available spreadsheets, these data-sets self-report what was being spent along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
-Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers to the full text [@Neylon_2014; @woodward_2014].
+Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers to the full text including the Digital Object Identifier (DOI) [@Neylon_2014; @woodward_2014].
 Parts of Jisc's cost data was examined by @Pinfield_2015.
@@
 Disclosed as publicly available spreadsheets, these data-sets self-report payments made along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version, and a link to a deposit in a subject repository.
-Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates.
+Curatorial efforts
+focused on the disambiguation of publisher and journal titles as well as on detecting duplicates.
 In the case of the Wellcome Trust, crowd-sourcing data cleaning activities through a Google spreadsheet in combination with checks against bibliographic sources massively improved the spending data (see comments in @wellcome_apc).-->
@@
 
-This paper focuses on how much German universities and research organisation spent on open access publication fees.
-In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced how universities manage institutional support for publication fees.[^3] Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" program in 2011, and similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012], only few central funds existed [@eppelin_2012].
+This paper focuses on how much German universities and research organisations spent on open access publication fees.
+In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced how universities manage institutional support for these charges.
+Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" program in 2011, and similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012], only few central funds existed [@eppelin_2012].
 Because the DFG has enforced a set of criteria grantees have to comply with, similar implementations for supporting open access publishing across German universities exists [@Fournier_2013].
@@
 Grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
-This includes central invoicing schemes, and related agreements between university libraries and publishers, leading to discounts for the authors [@Fournier_2013].
+This includes central invoicing schemes and memberships, among others, agreed between university libraries and publishers that, in many cases, lead to discounts [@Fournier_2013].
 
@@
 Payments to these journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall expenditure.
-In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
+The multidisciplinary PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline, were also well represented in the Open APC data-set.
+In general, an estimated 14 out of more than 10,000 journals registered in DOAJ in 2015 accounted for up to 15–20 % of all articles published in full open access journals [@Bj_rk_2015] In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
 
@@
 
-In Germany, institutional spending on open access publication fees has increased over the years, confirming the general trend of publication fees in academic publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-With a share of 99%, the majority of open access articles German institutions reported to the Open APC initiative were published in fully open access journals.
-This presumably reflects the DFG funding policy which excludes the support of articles published in hybrid open access journals.
-The DFG has been financially supporting the implementation of central publication funds at more than 30 German universities since 2009.
-However, in reviewing self-reported cost data from funders or countries that also support hybrid open access journals or open access books, we revealed smaller proportions of payments in favour of articles in fully open access journals.
-Because open access publication fee spending is fragmented, we cannot answer whether German researchers avoided opting for open access when publishing in hybrid journals or used other budgets to pay publication fees required to make their work open access through these kind of journals.
+In Germany, institutional spending on publication fees charged by open access journals has increased over the years.
+These findings are consistent with the general trend of publication fees as revenue source for open access publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
+They also reflect the increasing importance of providing central support to cover these charges [@Fournier_2013; @Kingsley_2014].
+Similar to investigations of the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending size varies across the German universities and research organisations included in the Open APC initiative.
+With a proportion of 39 % of the overall number of articles, the Max Planck Society, a large non-university research organisation, supported most open access journal articles.
+This likely reflects the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed large open access agreements with publishers over the last decade [@Schimmer_2013, @Sikora_2015].
+Many universities and research organisations, on the other hand, reported a remarkably lower number of supported articles.
 
-In our study, Crossref thoroughly indexed open access journal articles in the Open APC data-set.
-We could gather metadata representing publisher and journal titles for 99 % articles, and successfully merged these information with the Open APC cost data.
-Using metadata from Crossref, therefore, reduces the extensive validation work of bibliographic information provided that the reporting of the DOI along with the expenses is made mandatory.
-Drawing on Crossref would also increase the comparability of cost data for future negotiations with publishers on open access agreements, and the open access spending between open data initiatives that apply the same reporting standards, as its metadata represent the dynamic landscape of academic publishing in terms of ongoing mergers of publishing houses or name changes.
-
-Another advantage of self-reported data-sets on the article-level to disclose spending on open access publication fees is that they enable researchers and practitioners alike to study in which open access journals researchers from one institution actually publish, and to compare these findings with that of other universities or research organizations.
-For instance, our study revealed that the size of publication fee spending differed among the institutions with the Max Planck Society accounting for almost 39 % of the overall articles.
-Many universities and research organization reported remarkably lower number of supported open access articles to the Open APC initiative.
-Using self-reported data, therefore, contributes to the understanding about how much and to what extent spending on open access publishing varies on the institutional level.
-This is particularly relevant given the increasingly important role open access publishing plays in recent negotiations between German universities and research organizations forming consortia on the one side and publishers on the other about financing scholarly publishing in future [@allianz].
+To answer data curation issues while unifying spending data raised by UK initiatives, re-using DOIs to gather bibliographic metadata from Crossref is a promising approach.
+In our study, Crossref thoroughly indexed open access journal articles disclosed in the Open APC data-set, providing publisher and journal titles for 99 % of the overall article volume.
+Making use of metadata from Crossref, therefore, reduces the extensive validation work of bibliographic information provided that the DOI is made available in the cost data, a requirement needed to be addressed also in other fields of the quantitative study of scholarly communication [@Haustein_2016].
+Drawing on Crossref would also increase the comparability of cost data to prepare future negotiations with publishers on open access agreements, because Crossref's metadata reflects the current landscape of academic publishing in terms of ongoing mergers of publishing houses or name changes.
+Comparative studies of publication fee spending using data on the article-level would also benefit from such an approach.
 
 This study is limited in some respects.
-One is that we cannot assess whether publishers and journals granted publication fee discounts seeing that the Open APC initiative does not track this kind of information.
-However, the large price ranges of particular journals suggests that varying pricing schemes were in place.
+One is that we cannot assess whether publishers and journals granted publication fee discounts because the Open APC initiative does not track this kind of information.
+However, large price ranges of particular journals suggests that varying pricing schemes are in place, an observation also made earlier [@Solomon_2012, @Pinfield_2015].
 Adding to this complexity, it is likely that some institutions only paid parts of the publication fee.
 Take for instance the journal Nature Communication.
-Charges reported in our sample ranged between 2000 €, the DFG price cap, and 4.403 €.
-Such co-payments that involve several budgets were a proposed strategy to sustain publication funds at German universities [@Fournier_2013].
+Charges reported to the Open APC initiative ranged between 2000 €, the DFG price cap, and 4.403 €.
+Although such payments made from several budgets are a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro rata amounts cannot be recognised as such from the Open APC data, leading to a possibly flawed representations of publication fee spending in Germany.
 In another case, one university included its charges for participating in the German SCOAP consortia and presumably divided the sum by the articles published in SCOAP journals.
-In this national consortia, managed by the German National Library of Science and Technology, payments were not directly made per article.
-Instead, subscription costs between a participating library and a publisher were reconciled, and the reduction transferred to the consortia to finance publications in SCOAP journals.
+In German SCOAP consortia, managed by the German National Library of Science and Technology, payments were not directly made per article.
+Instead, subscription costs between a participating library and a publisher were reconciled, and the reduction transferred to the SCOAP to finance publications in particle physics journals.
 
@@
 
-Our analysis on how institutional spending per articles was distributed over publishers and journal titles shows that open access publishing is diverse and concentrated at the same time.
+Our analysis on how institutional spending per articles was distributed over publishers and journal titles indicates that open access publishing is heterogeneous and concentrated at the same time.
 While we were able to identify 139 individual publishing houses that were supported by the German universities and research organizations, the distribution is highly skewed.
-92 % of open access publication fee spending went to ten publishers, confirming a general high concentration of few publishers in current academic publishing.
-However, our study could not confirm that publications in open access journals owned by traditional publishing houses account for most of the spending on publication fees.
+92 % of open access publication fee spending went to ten publishers, being consistent with a observed high concentration of few publishers in current academic publishing [@Larivi_re_2015].
+However, our study could not confirm that publications in open access journals owned by traditional publishing houses account for most of the spending on publication fees as observed by @Pinfield_2015.
 Rather, open access publishers such as Public Library of Science (PloS), Copernicus GmbH or MPDI AG rank higher in our study than in the analyses of cost data in the UK.
 
-This study finally confirms the leading role of "mega-journals" in open access publishing, including the multidisciplinary PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline.
-In general, an estimated 14 out of more than 10,000 journals registered in DOAJ in 2015 accounted for up to 15–20 % of all articles published in full open access journals [@Bj_rk_2015].
+One possible explanation why traditional publishers are less well represented in our study, is the lack of cost information about hybrid open access journals.
+99 % of the overall article volume German universities and research organisations financially supported were published in fully open access journals.
+This presumably reflects the DFG funding policy which excludes the support of articles published in hybrid open access journals through many university open access funds.
+However, reviewing self-reported cost data from funders or countries that also support hybrid open access journals or open access books, we observe smaller proportions of payments in favour of articles in fully open access journals.
+Because open access publication fee spending is fragmented, it remains open to speculation whether authors affiliated with universities and research organisations simply avoided opting for open access when publishing in hybrid journals for political or economical reasons, or used other budgets to pay publication fees required to make their work open access through these kind of journals.
 

### 2016-06-05 article.Rmd
updated: background section

@@
 
-One reason why the study on how much was being spent is in most cases difficult is that total publication fee spending is fragmented across the budgets of grant agencies, research institutions, and libraries, or, if support is limited, are taken from personal budgets.
+The analysis of APC spending is difficult.
+The total spending is fragmented across the budgets of grant agencies, research institutions, and libraries, or is covered by personal budgets.
 Asking 9,645 authors from various disciplines how they financed publication fees, a comprehensive survey in 2010 revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.

### 2016-06-05 article.Rmd
updated: central funding

@@
 In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced how universities manage institutional support for these charges.
-Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" program in 2011, and similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012], only few central funds existed [@eppelin_2012].
-Because the DFG has enforced a set of criteria grantees have to comply with, similar implementations for supporting open access publishing across German universities exists [@Fournier_2013].
-These criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 € (excluding value added tax)[^3].
-Grantees agree not only to reimburse the bills on behalf of the researchers they support, but also to look for ways to improve the handling of those financial transactions.
+Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" program in 2011, only a few central funds existed [@eppelin_2012].
+This is similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012].
+The DFG has enforced a set of criteria grantees have to comply with  [@Fournier_2013], leading to similar implementations for open access supporting across German universities.
+These criteria exclude sponsoring of articles in hybrid journals and the funding of articles whose publication fee exceeds 2,000 €[^3].
+Grantees agree not only to pay for APCs but also to find ways to improve the handling of those financial transactions.
 This includes central invoicing schemes and memberships, among others, agreed between university libraries and publishers that, in many cases, lead to discounts [@Fournier_2013].
 
-Research institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft are not eligible for this DFG funding program.
+Research institutes outside of universities, e.g.
+those institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft, are not eligible for this DFG funding program.
 But in response, some organizations have adopted similar processes to support authors.
@@
 Reflecting Austrian and UK initiatives to share institutional spending on open access publication fees as open data, as well as professional discussions on open access publishing, Bielefeld University Library began to openly share its payment of publication fees in May 2014.
-After engaging with the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11] other German institutions joined under the umbrella of the Open APC initiative soon after.
+After engaging with the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11], other German institutions joined under the umbrella of the Open APC initiative soon after.
 

### 2016-06-05 article.Rmd
further updates to analysis and discussion

@@
 The aim of the study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-Using self-reported cost data from the Open APC initiative, the analysis focused on the amount that was being spent on publication fees, and compared these expenditure with data from related Austrian and UK initiatives, both in terms of size and the proportion of articles being published in fully and hybrid open access journals.
+Using self-reported cost data from the Open APC initiative, the analysis focused on the amount that was being spent on publication fees, and compared these expenditures with data from related Austrian and UK initiatives, both in terms of size and the proportion of articles being published in fully and hybrid open access journals.
 We also asked how thoroughly self-reported articles were indexed in Crossref, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.
@@
 We analysed self-reported cost data released by the Open APC initiative on May 13, 2016[^24] to assess institutional spending on open access publication fees in Germany.
-In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully or hybrid open access journal as well as the recorded DOI from the data-set.
+In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully open access journal or in a hybrid journal as well as the DOI reported in the data-set.
 
@@
 In the case of FWF, we gathered the cost information from the accompanying spending reports.
-We used the spreadsheet data to calculate Wellcome Trust's and Jisc's spending, and converted the prices from GBP to € in accordance with the average € foreign exchange reference rates provided by the European Central Bank.
-Our comparison between the open data initiatives focussed on the last two years 2014 and 2015.
+We used the spreadsheet data to calculate Wellcome Trust's and Jisc's spending, and converted the prices from GBP to € in accordance with the average foreign exchange reference rates provided by the European Central Bank.
+Our comparison of the open data initiatives focussed on the last two years 2014 and 2015.
 Because Wellcome Trust's spending was reported for the financial periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
@@
 
-After excluding payments being made for non-journal articles, as well as articles invoiced in 2016, we retrieved `r nrow(my_apc)` open access journal articles `r length(unique(my_apc$institution))` German universities and research institutions financially supported between 2005 and 2015.
-As illustrated in Figure 1, payments made for open access journal articles grew over the years.
+After excluding payments being made for non-journal articles as well as articles invoiced in 2016, we retrieved `r nrow(my_apc)` open access journal articles `r length(unique(my_apc$institution))` German universities and research institutions financially supported between 2005 and 2015.
+As illustrated in Figure 1, payments made for open access journal articles increased over the years.
 While one institution supported `r filter(my_apc, period == min(my_apc$period)) %>% nrow()` articles in `r as.character(min(my_apc$period))`, the majority shared their expenditure from 2013 onwards.
@@
 With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall submissions.
-By contrast, the universities of technology, TU Clausthal, TU Dortmund TU Ilmenau that recently begun to set up support structures for fee-based open access journal articles, self-reported low numbers of payments.
+By contrast, several universities that only recently begun to set up support structures for fee-based open access journal articles reported low numbers of payments.
 
@@
 Comparison of average prices suggests that publishing in hybrid journal was more expensive than in fully open access journals.
-Price differentials between these two categories were also reported earlier, indicating that prices for fully open access journals were typically lower [@Solomon_2012; @Pinfield_2015].
+Price differences between these two categories were also reported earlier, indicating that prices for fully open access journals were typically lower [@Solomon_2012; @Pinfield_2015].
 In 2014 and 2015, the mean price for fully open access journals calculated from all data-sets was below the DFG price cap of 2,000 €.
@@
 In comparison with data from the UK, some open access publishers have a greater share on total spending.
-@Pinfield_2015, for instance, reported remarkably lower numbers for the open access publishers MPDI AG, Copernicus GmbH, and Hindawi Publishing.
+@Pinfield_2015, for instance, reported remarkably lower numbers for the open access publishers MPDI, Copernicus, and Hindawi.
 
@@
 Payments to these journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall expenditure.
-The multidisciplinary PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline, were also well represented in the Open APC data-set.
-In general, an estimated 14 out of more than 10,000 journals registered in DOAJ in 2015 accounted for up to 15–20 % of all articles published in full open access journals [@Bj_rk_2015] In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
+The multidisciplinary journal PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline, were also well represented in the Open APC data-set.
+In general, an estimated 14 out of more than 10,000 journals registered in DOAJ in 2015 accounted for up to 15–20 % of all articles published in fully open access journals [@Bj_rk_2015] In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
 
@@
 With a proportion of 39 % of the overall number of articles, the Max Planck Society, a large non-university research organisation, supported most open access journal articles.
-This likely reflects the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed large open access agreements with publishers over the last decade [@Schimmer_2013, @Sikora_2015].
+This likely reflects the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed large open access agreements with publishers over the last decade [@Schimmer_2013, @Sikora_2015] leading not only to a large number of supported articles but also to an advantage in monitoring and thus being able to report on this data.
 Many universities and research organisations, on the other hand, reported a remarkably lower number of supported articles.
@@
 This study is limited in some respects.
-One is that we cannot assess whether publishers and journals granted publication fee discounts because the Open APC initiative does not track this kind of information.
+One is that we cannot assess whether publishers and journals granted publication fee discounts because the Open APC initiative, dedicated to make self-reporting easy by using a minimal data schema, does not track this kind of information.
 However, large price ranges of particular journals suggests that varying pricing schemes are in place, an observation also made earlier [@Solomon_2012, @Pinfield_2015].
@@
 Charges reported to the Open APC initiative ranged between 2000 €, the DFG price cap, and 4.403 €.
-Although such payments made from several budgets are a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro rata amounts cannot be recognised as such from the Open APC data, leading to a possibly flawed representations of publication fee spending in Germany.
-In another case, one university included its charges for participating in the German SCOAP consortia and presumably divided the sum by the articles published in SCOAP journals.
-In German SCOAP consortia, managed by the German National Library of Science and Technology, payments were not directly made per article.
-Instead, subscription costs between a participating library and a publisher were reconciled, and the reduction transferred to the SCOAP to finance publications in particle physics journals.
+Although such payments made from several budgets are a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro rata amounts cannot be identified as such within the Open APC data, leading to a possibly flawed representations of publication fee spending in Germany.
+In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the articles published by their authors in SCOAP^3-covered journals.^[26] This is very arbitrary, since the mean APC for an institution can only be determined after the end of a full 3-year funding cycle.
 
@@
 Others pointed out that the workload to produce such a data-set could be too extensive [@deppe_2015].
-As no registry of institutional open access funds or similar support structures exists, we cannot assess the number of non-participants in Germany.
+As no complete registry of institutional open access funds or similar support structures exists, we cannot assess the number of non-participants in Germany.
 
-Our analysis on how institutional spending per articles was distributed over publishers and journal titles indicates that open access publishing is heterogeneous and concentrated at the same time.
+Our analysis of how institutional spending per articles was distributed over publishers and journal titles indicates that open access publishing is heterogeneous and concentrated at the same time.
 While we were able to identify 139 individual publishing houses that were supported by the German universities and research organizations, the distribution is highly skewed.
@@
 However, our study could not confirm that publications in open access journals owned by traditional publishing houses account for most of the spending on publication fees as observed by @Pinfield_2015.
-Rather, open access publishers such as Public Library of Science (PloS), Copernicus GmbH or MPDI AG rank higher in our study than in the analyses of cost data in the UK.
+Rather, open access publishers such as Public Library of Science (PloS), Copernicus or MPDI rank higher in our study than in the analyses of cost data in the UK.
 
@@
 However, reviewing self-reported cost data from funders or countries that also support hybrid open access journals or open access books, we observe smaller proportions of payments in favour of articles in fully open access journals.
-Because open access publication fee spending is fragmented, it remains open to speculation whether authors affiliated with universities and research organisations simply avoided opting for open access when publishing in hybrid journals for political or economical reasons, or used other budgets to pay publication fees required to make their work open access through these kind of journals.
+Because open access publication fee spending is fragmented, it remains open to speculation whether authors affiliated with universities and research organisations simply avoided opting for open access when publishing in hybrid journals for political or economical reasons, or used other budgets for these fees—budgets that did not report data to the Open APC initiative.
+This could support a key operational idea of the initiative: having organisations report on spendings they have made, which might work better when done through organisation-wide funds.
 
@@
 
-[^24]: https://github.com/OpenAPC/openapc-de/tree/v2.4.3
+[^24]: <https://github.com/OpenAPC/openapc-de/tree/v2.4.3>
 
 [^25]: <https://github.com/CrossRef/rest-api-doc/blob/master/rest_api.md>
+
+[^26]: SCOAP^3, the Sponsoring Consortium for Open Access Publishing in Particle Physics, is a unique approach to convert former subscription journals in high-energy physics to open access journals under a CC BY license.
+The consortiums, led by CERN, pays publishers centrally, based on previously agreed APCs and an overall price cap, and retrieves its funds from organisations and countries based on their share in the articles published in the covered journals.
+German universities participate through an initiative led by the German National Library of Science and Technology that received additional funding from DFG.

### 2016-06-05 article.Rmd
conclusion first draft

@@
 
-The analysis of APC spending is difficult.
-The total spending is fragmented across the budgets of grant agencies, research institutions, and libraries, or is covered by personal budgets.
+The study of institutional spending on open access journal articles has been limited for several reasons.
+One is that payment on these charges is fragmented across the budgets of grant agencies, research institutions, and libraries, or is covered by personal budgets.
 Asking 9,645 authors from various disciplines how they financed publication fees, a comprehensive survey in 2010 revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.
@@
 
-Hybrid journals substantially add to this complexity of open access funding [@Kingsley_2014; @Pinfield_2015].
+Hybrid journals substantially add to this complexity of open access funding [@Bj_rk_learned_2014; @Kingsley_2014; @Pinfield_2015].
 These journals, allowing articles to be published immediately as open access after a charge was paid, rely both on subscriptions and publication fees as revenue sources.
-Although the uptake of open access through hybrid journals was described as lower and more expensive compared to that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014], this model has gained attention through recent science policies, notably because of open access policies from the UK [@Pinfield_Review_2015].
+Although the uptake of open access through hybrid journals was described as lower and more expensive compared to that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014], this model has gained increasing attention because of recent open access science policies, notably from the UK [@Pinfield_Review_2015].
 
@@
 The not-for-profit company Jisc followed by collecting data from UK universities [@Lawson_data_2015].
-Disclosed as publicly available spreadsheets, these data-sets self-report what was being spent along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
+Disclosed as publicly available spreadsheets, these data-sets self-report expenditure along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
 Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers to the full text including the Digital Object Identifier (DOI) [@Neylon_2014; @woodward_2014].
 Parts of Jisc's cost data was examined by @Pinfield_2015.
-Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as a varying number of articles UK universities supported between 2007 - 2014, confirming earlier studies that collected price information from journal websites.
-
-<!-- This complex situation of fee-based open access publishing creates difficulties for researchers and practitioners alike.
-Some European initiatives have therefore begun to collect and disclose their expenditures on open access journal articles as open data.
-To our knowledge, the first research funders providing such data were the Wellcome Trust [@fwf_apc_13] and the Austrian Science Fund FWF [@wellcome_apc_12].
-The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015].
-Disclosed as publicly available spreadsheets, these data-sets self-report payments made along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version, and a link to a deposit in a subject repository.
-Curatorial efforts
-focused on the disambiguation of publisher and journal titles as well as on detecting duplicates.
-In the case of the Wellcome Trust, crowd-sourcing data cleaning activities through a Google spreadsheet in combination with checks against bibliographic sources massively improved the spending data (see comments in @wellcome_apc).-->
+Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as a varying number of articles UK universities supported between 2007 - 2014, being consistent with earlier studies that collected price information from journal websites [@Solomon_2012].
 
@@
 This is similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012].
-The DFG has enforced a set of criteria grantees have to comply with  [@Fournier_2013], leading to similar implementations for open access supporting across German universities.
-These criteria exclude sponsoring of articles in hybrid journals and the funding of articles whose publication fee exceeds 2,000 €[^3].
-Grantees agree not only to pay for APCs but also to find ways to improve the handling of those financial transactions.
-This includes central invoicing schemes and memberships, among others, agreed between university libraries and publishers that, in many cases, lead to discounts [@Fournier_2013].
+The DFG has enforced a set of criteria grantees have to comply with  [@Fournier_2013], leading to related implementations for open access supporting across German universities.
+These criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 €[^3].
+Grantees agree not only to pay for APCs, but also to find ways to improve the handling of those financial transactions.
+This includes central invoicing schemes and memberships, among others, agreed between university libraries and publishers that often lead to a discount on publication fees [@Fournier_2013].
 
-Research institutes outside of universities, e.g.
-those institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft, are not eligible for this DFG funding program.
-But in response, some organizations have adopted similar processes to support authors.
-The Max-Planck-Gesellschaft operates their long-lasting open access activities, including handling spending and publisher agreements centrally, through the Max Planck Digital Library [@Schimmer_2013; @Sikora_2015], while the Leibniz-Gemeinschaft set up a dedicated open access fund in 2016.
+Non-university research organisations, mainly those institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft, are not eligible for this DFG funding program.
+But in response, some organisations have adopted similar processes to support authors.
+The Max-Planck-Gesellschaft operates their long-lasting open access activities, including handling spending and publisher agreements centrally, through the Max Planck Digital Library (MPDL) [@Schimmer_2013; @Sikora_2015], while the Leibniz-Gemeinschaft set up a dedicated open access fund in 2016.
 
 The evolving institutional support structures to cover open access publication fees has led to calls for an unified approach towards supporting open access journal publishing in Germany.
+The MPDL called for re-allocating subscriptions in favour of financing open access charges in 2014 [@Schimmer_Geschuhn_Vogler].
 The Allianz der Wissenschaftsorganisationen[^10c], a science policy board representing all major research organisations in Germany, marked price transparency as one way to sustain an "adequate open access publication system" [@allianz].
 Reflecting Austrian and UK initiatives to share institutional spending on open access publication fees as open data, as well as professional discussions on open access publishing, Bielefeld University Library began to openly share its payment of publication fees in May 2014.
-After engaging with the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11], other German institutions joined under the umbrella of the Open APC initiative soon after.
+After engaging with the working group "Electronic Publishing" of the Deutsche Initiative für Netzwerkinformation (DINI)[^11], other German institutions joined under the umbrella of the Open APC initiative soon after [@openapc].
 
@@
 The aim of the study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-Using self-reported cost data from the Open APC initiative, the analysis focused on the amount that was being spent on publication fees, and compared these expenditures with data from related Austrian and UK initiatives, both in terms of size and the proportion of articles being published in fully and hybrid open access journals.
-We also asked how thoroughly self-reported articles were indexed in Crossref, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.
+Using self-reported cost data from the Open APC initiative, the analysis focused on the amount that was being spent on publication fees, and compared these expenditure with data from related Austrian and UK initiatives, both in terms of size and the proportion of articles being published in fully and hybrid open access journals.
+We also asked how thoroughly self-reported articles were indexed in Crossref, a DOI minting agency for scholarly literature, and analysed how the institutional spending was distributed over publishers and journal titles.
 
@@
 
-We analysed self-reported cost data released by the Open APC initiative on May 13, 2016[^24] to assess institutional spending on open access publication fees in Germany.
+We analysed self-reported cost data released by the Open APC initiative on May 13, 2016,[^24] to assess institutional spending on open access publication fees in Germany.
 In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully open access journal or in a hybrid journal as well as the DOI reported in the data-set.
@@
 We fetched bibliographic metadata for each article from Crossref on May 19, 2016, on the basis of the reported DOIs.
-Although the Open APC initiative gathered metadata representing publishers and journals from Crossref as well, this information was retrieved at the time when the participating institutions submitted their expenses.
+Although the Open APC initiative gathered metadata representing publishers and journals from Crossref as well, this information was retrieved at the time when the participating institutions submitted their expenditure.
 The Open APC initiative kept track of the date when these data-sets were submitted with Git, a version control system, increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub to be transparent over time.
@@
 A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that operated as Springer Nature at the time of our study.
-To reflect these dynamics in academic publishing, we decided to retrieve updated metadata from Crossref for the whole Open APC data-set instead of re-using the historical publisher and journal information contained in the Open APC data-set.
+To reflect these dynamics in academic publishing, we decided to retrieve  updated bibliographic metadata from Crossref, and to merge these records with the administrative information instead of re-using the historical publisher and journal information contained in the Open APC data-set.
 
@@
 In the case of FWF, we gathered the cost information from the accompanying spending reports.
-We used the spreadsheet data to calculate Wellcome Trust's and Jisc's spending, and converted the prices from GBP to € in accordance with the average foreign exchange reference rates provided by the European Central Bank.
+We used the spreadsheet data to calculate Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average foreign exchange reference rates provided by the European Central Bank.
 Our comparison of the open data initiatives focussed on the last two years 2014 and 2015.
-Because Wellcome Trust's spending was reported for the financial periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
-We excluded articles from the analysis for which either cost information or journal type was missing.
+Because Wellcome Trust's spending was reported for the fiscal periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
+We excluded articles from the analysis for which  information about the cost or the journal type was missing.
 In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to 4.861.772 € from the analysis because no publication type was given in the data-set.
 
-The data collection methods of the Open APC initiative and those of the other initiatives differed in some aspects.
-For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, in the Wellcome Trust data a publication identifier was added as part of the automated compliance checks.
-Our first screening of the data-sets revealed that some articles lacked a DOI.
-For this reason, and as our main focus was institutional funding for publication fees in Germany, we decided to only compare German APC spending with that reported by other initiatives.
-We did not compare the distribution of spending over publishers and journal titles, or the indexing coverage in Crossref.
+Data curation activities of the Open APC initiative and those of the other initiatives differed in some aspects.
+For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, in case of the Wellcome Trust spending data, DOIs were also being identified by automated compliance checks.
+Our first screening of the data-sets revealed that some articles published in Crossref-indexed journals lacked a DOI.
+Because of these different methods to curate the cost data, and as our main focus was institutional funding for publication fees in Germany, we decided to only compare German APC spending with that reported by other initiatives.
+We, therefore, did not analyse the distribution of spending over publishers and journal titles, or the indexing coverage in Crossref for the Austrian and UK spending data.
 
@@
 As illustrated in Figure 1, payments made for open access journal articles increased over the years.
-While one institution supported `r filter(my_apc, period == min(my_apc$period)) %>% nrow()` articles in `r as.character(min(my_apc$period))`, the majority shared their expenditure from 2013 onwards.
-With `r max(apc_time$articles)` articles, the year `r apc_time[apc_time$articles == max(apc_time$articles),"period"]` was best represented in our data-set.
-However, the data-set contained `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institutions that contributed their cost data for 2015 at the time of this analysis, suggesting a time lag between payments made and reporting these spending to the Open APC initiative.
+While one institution supported `r filter(my_apc, period == min(my_apc$period)) %>% nrow()` articles in `r as.character(min(my_apc$period))`, most institutions included in our study shared their expenditure from 2013 onwards.
+With `r max(apc_time$articles)` articles, the year `r apc_time[apc_time$articles == max(apc_time$articles),"period"]` was best represented.
+However, at the time of analysis, not all, but `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institutions contributed cost information, suggesting a time lag between payments made and reporting these spending to the Open APC initiative.
 
@@
 
-Figure 3 presents the large price variations.
+Figure 3 presents institutional spending per article and year.
+Large price variations could be observed.
 Publication fees that were paid by the German universities and research organisations ranged between `r min(my_apc$euro)` € and `r max(my_apc$euro)` €  (SD `r sd(my_apc$euro)`).
 However, the average price paid varied somewhat during the period 2011 and 2015 (`r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - `r my_apc %>% filter(period == 2015) %>% summarise(mean(euro)) %>% round()` €).
-Average publication fee spending per article in Germany was `r mean(my_apc$euro)` (Median = `r median(my_apc$euro)` €, SD = `r sd(my_apc$euro)` €).
 
@@
 The number of APC payments per institutions varied considerably (see Table 1).
-With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall submissions.
-By contrast, several universities that only recently begun to set up support structures for fee-based open access journal articles reported low numbers of payments.
+With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall article volume.
+By contrast, for several universities that only recently begun to set up centrally managed open access funds to cover publication fees we observed a lower number of support for open access journal articles.
 
@@
 Prices were converted according to the average Euro exchange rate of the examined periods, and gathered for both fully and hybrid open access journals.
-The comparison reveals that the Open APC initiative lacked cost information about hybrid journals, whereas the related Austrian and UK open data initiatives could report a large share of spending on these journals between 2014 and 2015.
-This situation presumably reflected the importance of the DFG as funding source for publication fees in Germany.
-The DFG funding policy excluded support for publications in hybrid journals.
+The comparison revealed that the Open APC initiative lacked cost information about hybrid journals, whereas the related Austrian and UK open data initiatives reported a large share of spending on these journals between 2014 and 2015.
 Over the years 2005 – 2015, `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` German universities and research institutions reported `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` hybrid journal articles to the Open APC initiative, representing `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of the overall supported article volume.
@@
 Comparison of average prices suggests that publishing in hybrid journal was more expensive than in fully open access journals.
-Price differences between these two categories were also reported earlier, indicating that prices for fully open access journals were typically lower [@Solomon_2012; @Pinfield_2015].
+Price differences between these two categories were also reported earlier, indicating that prices for fully open access journals were on average lower [@Solomon_2012; @Pinfield_2015].
 In 2014 and 2015, the mean price for fully open access journals calculated from all data-sets was below the DFG price cap of 2,000 €.
@@
 We used the DOI to automatically fetch publisher and journal names for each article from the Crossref REST API.
-Table 4 shows the top ten publishers in terms of payments made that represented `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall spending on publication fees.
+Table 2 shows the top ten publishers in terms of payments made.
+They represented `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall spending on publication fees.
 In total, payments were made to `r my_apc %>% group_by(publisher) %>% summarize(n()) %>% nrow()` publishing houses.
-In comparison with data from the UK, some open access publishers have a greater share on total spending.
+In comparison with data from the UK, a greater share on total spending of some open access publishers could be observed.
 @Pinfield_2015, for instance, reported remarkably lower numbers for the open access publishers MPDI, Copernicus, and Hindawi.
@@
 
-Most of publication fee spending in Germany was on articles published in Springer Nature journals, most likely profiting from the merge with the open access publisher BioMed Central in 2008, and that between the well-established publishers Springer Science + Business Media and Nature Publishing Group in 2015.
+Most of publication fee spending in Germany was on articles published in Springer Nature journals, likely profiting from the merge with the open access publisher BioMed Central in 2008, and that between the well-established publishers Springer Science + Business Media and Nature Publishing Group in 2015.
 Using the Crossref-Member-ID instead of the publisher name, we were able to differentiate between journals formely published by Springer Science + Business Media and Nature Publishing Group.
 In terms of articles, the majority of payments made were for publications in journals formerly associated with Springer Science + Business Media.
-Springer Science + Business Media journals accounted for `r nrow(springer)` articles, representing `r nrow(springer) / nrow(springer_nat) * 100`% of the overall Springer Nature article volume in the Open APC data, and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % in terms of the amount being spent.
+Springer Science + Business Media journals accounted for `r nrow(springer)` articles, representing `r nrow(springer) / nrow(springer_nat) * 100`% of the overall Springer Nature article volume recorded by the Open APC initiative, and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % in terms of the amount being spent.
 Median publication fee spending slightly differed between Springer Science + Business Media (`r median(springer$euro)` €) and Nature Publishing (`r median(nature$euro)` €).
@@
 
-In contrast to Springer Nature, other well-established publishing houses such as Elsevier and Wiley-Blackwell rank lower in our analysis.
-Table 2, furthermore, illustrates the price variations across and within publishers, which confirms earlier findings [@Solomon_2012; @Pinfield_2015].
+In contrast to Springer Nature, other well-established publishing houses such as Elsevier and Wiley-Blackwell ranked lower in our analysis.
 
@@
 
-Prices also varied within single journals.
+Prices also varied within journals.
 Based on the number of articles paid for, Table 3 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals.
-Payments to these journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall expenditure.
+We normalised PLOS journal titles, because the name chnages from "PLoS" to "PLOS" was only partly represented in the Crossref metadata at the time of our study.
+Payments to the top ten journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall expenditure.
 The multidisciplinary journal PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline, were also well represented in the Open APC data-set.
@@
 These findings are consistent with the general trend of publication fees as revenue source for open access publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-They also reflect the increasing importance of providing central support to cover these charges [@Fournier_2013; @Kingsley_2014].
+They also reflect the evolving institutional support to cover these charges in Germany [@Fournier_2013].
 Similar to investigations of the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending size varies across the German universities and research organisations included in the Open APC initiative.
-With a proportion of 39 % of the overall number of articles, the Max Planck Society, a large non-university research organisation, supported most open access journal articles.
-This likely reflects the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed large open access agreements with publishers over the last decade [@Schimmer_2013, @Sikora_2015] leading not only to a large number of supported articles but also to an advantage in monitoring and thus being able to report on this data.
+With a proportion of 39 % of the overall article volume, the Max Planck Society, a large non-university research organisation, supported most open access journal publications.
+This likely reflects the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed large open access agreements with publishers over the last decade [@Schimmer_2013, @Sikora_2015].
+This presumably not only resulted in a large number of supported articles, but also contributed to an advantage in monitoring and thus being able to report on this data.
 Many universities and research organisations, on the other hand, reported a remarkably lower number of supported articles.
 
-To answer data curation issues while unifying spending data raised by UK initiatives, re-using DOIs to gather bibliographic metadata from Crossref is a promising approach.
+To answer data curation issues while unifying UK spending data [@Neylon_2014; @woodward_2014], re-using DOIs to gather bibliographic metadata from Crossref is a promising approach.
 In our study, Crossref thoroughly indexed open access journal articles disclosed in the Open APC data-set, providing publisher and journal titles for 99 % of the overall article volume.
-Making use of metadata from Crossref, therefore, reduces the extensive validation work of bibliographic information provided that the DOI is made available in the cost data, a requirement needed to be addressed also in other fields of the quantitative study of scholarly communication [@Haustein_2016].
-Drawing on Crossref would also increase the comparability of cost data to prepare future negotiations with publishers on open access agreements, because Crossref's metadata reflects the current landscape of academic publishing in terms of ongoing mergers of publishing houses or name changes.
-Comparative studies of publication fee spending using data on the article-level would also benefit from such an approach.
+Making use of metadata from Crossref, therefore, reduces the extensive validation work of bibliographic information provided that the DOI is made available in the cost data.
+DOI availability is a requirement needed to be addressed in other fields of the quantitative study of scholarly communication as well, such as in altmetrics to deal with heterogeneous data [@Haustein_2016].
+Drawing on Crossref has the potential to increase the comparability of cost data to prepare future negotiations with publishers on open access agreements, because Crossref's metadata reflects the current landscape of academic publishing in terms of ongoing mergers of publishing houses.
+Future comparative studies of publication fee spending using data on the article-level also benefit from such an approach.
 
 This study is limited in some respects.
-One is that we cannot assess whether publishers and journals granted publication fee discounts because the Open APC initiative, dedicated to make self-reporting easy by using a minimal data schema, does not track this kind of information.
+One is that we cannot assess whether publishers and journals granted publication fee discounts because the Open APC initiative uses a minimal data scheme to make self-reporting easy, and, therefore, does not track this kind of information.
 However, large price ranges of particular journals suggests that varying pricing schemes are in place, an observation also made earlier [@Solomon_2012, @Pinfield_2015].
@@
 In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the articles published by their authors in SCOAP^3-covered journals.^[26] This is very arbitrary, since the mean APC for an institution can only be determined after the end of a full 3-year funding cycle.
+Other factors affecting price variations are currency movements over the years, and different tax rates in Germany.
+For instance, the Max Planck Society has a limited input tax reduction.
+The refund of input value added tax for publication fees is 20%.
+Local and payment currency, as well as taxes, as suggested by @Pinfield_2015 to increase the transparency of publication fee spending, were not disclosed in the Open APC initiative's data-set at the time of our study.
 
@@
 Therefore, not all institutions in Germany that provide central funding of publication fees contribute cost data to this initiative.
-In a qualitative survey, asking why German institutions are reluctant to share their cost data through the Open APC initiative, one institution feared that increase in transparency would allow publishers to adjust prices in their favour.
+In a qualitative survey, asking why German institutions are reluctant to share their cost data through the Open APC initiative, one institution feared that an increase in transparency would allow publishers to adjust prices in their favour.
 Others pointed out that the workload to produce such a data-set could be too extensive [@deppe_2015].
-As no complete registry of institutional open access funds or similar support structures exists, we cannot assess the number of non-participants in Germany.
+As no reliable registry of institutional open access funds or similar support structures in Germany exists, we cannot assess the number of non-participants in Germany.
 
@@
 While we were able to identify 139 individual publishing houses that were supported by the German universities and research organizations, the distribution is highly skewed.
-92 % of open access publication fee spending went to ten publishers, being consistent with a observed high concentration of few publishers in current academic publishing [@Larivi_re_2015].
+Ten publishers collected 92 % of open access publication fee spending, being consistent with an observed high concentration of few publishers in current academic publishing [@Larivi_re_2015].
 However, our study could not confirm that publications in open access journals owned by traditional publishing houses account for most of the spending on publication fees as observed by @Pinfield_2015.
@@
 99 % of the overall article volume German universities and research organisations financially supported were published in fully open access journals.
-This presumably reflects the DFG funding policy which excludes the support of articles published in hybrid open access journals through many university open access funds.
-However, reviewing self-reported cost data from funders or countries that also support hybrid open access journals or open access books, we observe smaller proportions of payments in favour of articles in fully open access journals.
-Because open access publication fee spending is fragmented, it remains open to speculation whether authors affiliated with universities and research organisations simply avoided opting for open access when publishing in hybrid journals for political or economical reasons, or used other budgets for these fees—budgets that did not report data to the Open APC initiative.
-This could support a key operational idea of the initiative: having organisations report on spendings they have made, which might work better when done through organisation-wide funds.
+This presumably reflects the DFG funding policy that excludes the support of articles published in hybrid open access journals.
+However, while reviewing self-reported cost data from funders or countries that also support hybrid open access journals, there are smaller proportions of payment on  articles in fully open access journals.
+Because open access publication fee spending is fragmented and often in-transparent, it remains open to speculation whether authors affiliated with German universities and research organisations avoid opting for open access when publishing in hybrid journals, or simply use other budgets that are not covered by the Open APC initiative.
 
@@
 
-<!-- The purpose of the Open APC initiative is to gain better insight on spending for open access.
-The main operational idea—to make data submission easy by building upon existing data and by using popular open tools and data bases—has proven to be a feasible approach and to scale well.
-Starting with a few institutions, the integration of further data contributors has not been a problem.
-While we use article metadata reported by publishers to Crossref, we think it's best to collect information on how much has been spent at the sources: the institutions paying for APCs.
-We look forward to more data contributors to the initiative and to adapt this scheme to new analyses, additional data, or international sources.
-The Open APC initiative can serve as an example of the value that open data, open source software, and the knowledge and skills of tech-savvy librarians add to modern tasks of libraries.-->
+Our study reveals the size and extent of spending on open access journals using publication fees in Germany.
+Drawing on self-reported cost data from the Open APC initiative, payments from German universities and research institutions have grown over the years.
+Comparing these expenditure with those from Austria and the UK, German open access funding is mostly focussed on fully open access journals, raising important questions about the role hybrid open access journals play as publication venue for researchers affiliated with the institutions included in our study.
+Given our findings and the general discussion on funding policies concerning hybrid open access journals, questions about if and to what extent science policy interventions, as well as the availability of institutional support, influence how researchers publish  are of particular concern.
+
+Using self-reported data and gathering publisher and journal information from Crossref, our study extends methods and improves data collection activities for researchers and practitioners alike, as well as contribute to a better understanding of factors affecting the analysis of publication fees in open access publishing.
+In this regard, our research highlights large variation in the distribution of spending that needs to be taken into consideration when studying payment on publication at the institutional level.
+We have also confirmed previous studies showing large price variations across publishers and open access journals that need to be better understood in future.
 

### 2016-06-05 article.Rmd
minor fixes discussioN

@@
 These findings are consistent with the general trend of publication fees as revenue source for open access publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-They also reflect the evolving institutional support to cover these charges in Germany [@Fournier_2013].
-Similar to investigations of the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending size varies across the German universities and research organisations included in the Open APC initiative.
-With a proportion of 39 % of the overall article volume, the Max Planck Society, a large non-university research organisation, supported most open access journal publications.
-This likely reflects the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed large open access agreements with publishers over the last decade [@Schimmer_2013, @Sikora_2015].
-This presumably not only resulted in a large number of supported articles, but also contributed to an advantage in monitoring and thus being able to report on this data.
-Many universities and research organisations, on the other hand, reported a remarkably lower number of supported articles.
+They also demonstrate the growing institutional support for researchers to cover these charges in Germany [@Fournier_2013].
+Similar to the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending size varies across the German universities and research organisations.
+With a proportion of 39 % of the total article volume, the Max Planck Society, a large non-university research organisation, supported most open access journal publications included in our study.
+A possible explanation is the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed open access agreements with publishers over the last decade on behalf of most Max Planck institutes [@Schimmer_2013, @Sikora_2015].
+This centralised approach presumably not only resulted in a large number of supported open access articles, but also advanced capabilities and skills needed to report on these expenditure on a regular basis.
+Many universities and research organisations, on the other hand, disclosed a remarkably lower number of supported articles.
 
 To answer data curation issues while unifying UK spending data [@Neylon_2014; @woodward_2014], re-using DOIs to gather bibliographic metadata from Crossref is a promising approach.
-In our study, Crossref thoroughly indexed open access journal articles disclosed in the Open APC data-set, providing publisher and journal titles for 99 % of the overall article volume.
-Making use of metadata from Crossref, therefore, reduces the extensive validation work of bibliographic information provided that the DOI is made available in the cost data.
-DOI availability is a requirement needed to be addressed in other fields of the quantitative study of scholarly communication as well, such as in altmetrics to deal with heterogeneous data [@Haustein_2016].
-Drawing on Crossref has the potential to increase the comparability of cost data to prepare future negotiations with publishers on open access agreements, because Crossref's metadata reflects the current landscape of academic publishing in terms of ongoing mergers of publishing houses.
-Future comparative studies of publication fee spending using data on the article-level also benefit from such an approach.
+In our study, Crossref thoroughly indexed open access journal articles disclosed in the Open APC data-set, providing information about publisher and journal titles for 99 % of all articles included in the Open APC data-set.
+Making use of metadata from Crossref, therefore, reduces extensive validation work of bibliographic records as long as the DOIs are made available in the cost data.
+Beyond identifying publishers and journals, mandatory reporting of DOIs in the spending data can also increase the re-use of such data to study other aspects of open access journals using publication fees.
+For instance, impact analysis in the field of altmetrics are dependent on the availability of DOIs as well [@Haustein_2016].
+Drawing on Crossref has the potential to increase the comparability of cost data to prepare future negotiations with publishers on open access agreements, because Crossref's metadata represents current developments in academic publishing in terms of ongoing mergers of publishing houses.
+In addition to these practical consequences, future comparative studies of publication fee spending using data on the article-level can also benefit from such an approach.
 
 This study is limited in some respects.
-One is that we cannot assess whether publishers and journals granted publication fee discounts because the Open APC initiative uses a minimal data scheme to make self-reporting easy, and, therefore, does not track this kind of information.
-However, large price ranges of particular journals suggests that varying pricing schemes are in place, an observation also made earlier [@Solomon_2012, @Pinfield_2015].
+One is that we cannot assess whether publishers and journals granted publication fee discounts.
+The   Open APC initiative uses a minimal data scheme to incentives self-reporting, and, therefore, does not track this kind of information.
+However, large price ranges suggests that different pricing levels and varying pricing schemes are in place, an observation also made earlier [@Solomon_2012; @Pinfield_2015].
 Adding to this complexity, it is likely that some institutions only paid parts of the publication fee.
@@
 Charges reported to the Open APC initiative ranged between 2000 €, the DFG price cap, and 4.403 €.
-Although such payments made from several budgets are a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro rata amounts cannot be identified as such within the Open APC data, leading to a possibly flawed representations of publication fee spending in Germany.
-In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the articles published by their authors in SCOAP^3-covered journals.^[26] This is very arbitrary, since the mean APC for an institution can only be determined after the end of a full 3-year funding cycle.
+Although such payments being made from several budgets are a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro-rata payment were not made transparent in the Open APC data, leading to a possibly flawed representations of publication fee spending in Germany.
+In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the articles published by their authors in SCOAP^3-covered journals.^[26] This is very arbitrary, since averages for an institution can only be determined after the end of a full 3-year funding cycle.
 Other factors affecting price variations are currency movements over the years, and different tax rates in Germany.
@@
 Local and payment currency, as well as taxes, as suggested by @Pinfield_2015 to increase the transparency of publication fee spending, were not disclosed in the Open APC initiative's data-set at the time of our study.
+In addition, the Open APC data-set does not track funding sources so that we cannot determine which funders co-financed publication fees.
 
@@
 Others pointed out that the workload to produce such a data-set could be too extensive [@deppe_2015].
-As no reliable registry of institutional open access funds or similar support structures in Germany exists, we cannot assess the number of non-participants in Germany.
+As no reliable registry of institutional open access funds or related support structures in Germany exists, we cannot assess the number of non-participants in Germany.
 
-Our analysis of how institutional spending per articles was distributed over publishers and journal titles indicates that open access publishing is heterogeneous and concentrated at the same time.
+Our analysis of how institutional spending per articles was distributed over publishers and journals indicates that open access publishing is heterogeneous and concentrated at the same time.
 While we were able to identify 139 individual publishing houses that were supported by the German universities and research organizations, the distribution is highly skewed.
@@
 
-One possible explanation why traditional publishers are less well represented in our study, is the lack of cost information about hybrid open access journals.
-99 % of the overall article volume German universities and research organisations financially supported were published in fully open access journals.
+One possible explanation why traditional publishers are less well represented in our study is the lack of cost information about hybrid open access journals.
+99 % of all articles German universities and research organisations financially supported were published in fully open access journals.
 This presumably reflects the DFG funding policy that excludes the support of articles published in hybrid open access journals.
-However, while reviewing self-reported cost data from funders or countries that also support hybrid open access journals, there are smaller proportions of payment on  articles in fully open access journals.
+However, while reviewing self-reported cost data from Austria and the UK where hybrid open access journals are generally supported, we observe a lower proportion of payment on articles in fully open access journals.
 Because open access publication fee spending is fragmented and often in-transparent, it remains open to speculation whether authors affiliated with German universities and research organisations avoid opting for open access when publishing in hybrid journals, or simply use other budgets that are not covered by the Open APC initiative.
@@
 Drawing on self-reported cost data from the Open APC initiative, payments from German universities and research institutions have grown over the years.
-Comparing these expenditure with those from Austria and the UK, German open access funding is mostly focussed on fully open access journals, raising important questions about the role hybrid open access journals play as publication venue for researchers affiliated with the institutions included in our study.
-Given our findings and the general discussion on funding policies concerning hybrid open access journals, questions about if and to what extent science policy interventions, as well as the availability of institutional support, influence how researchers publish  are of particular concern.
+Comparing these expenditure with those from Austria and the UK, German open access funding is mostly focussed on fully open access journals, raising important questions about hybrid open access journals as publication venue.
+Given our findings, and in light of the general discussion about funding policies addressing hybrid open access journals, questions about whether and to what extent science policy interventions, as well as the availability of institutional support, influence how researchers publish  are of particular concern.
 

### 2016-06-05 article.Rmd
abstract + conclusion

@@
 ---
-title: 'Disclosing funding sources for open access publication fees: the Open APC
-  initiative'
+title: 'A Study of Institutional Spending on Open Access Publication Fees in Germany'
 author:
@@
 csl: frontiers.csl
-abstract: Publication fees in open access publishing hold a prominent place on the
-  agenda of researchers, policy-makers, and academic publishers. This paper contributes
-  to the evolving empirical basis on open access funding. It describes the Open APC
-  initiative, in which German universities and research organizations share their
-  expenditures for publication fees. As method, the initiative uses existing open
-  data tools to aggregate and disseminate institutional spending on open access publication
-  fees. In total, 29 German research organizations self-reported funding of 6,279
-  open access journal articles, which amounted to 8,039,339 €. The average payment
-  for each article was 1,280 €, and the median payment 1,209 €. Our data-set comprises
-  only 53 articles in hybrid journals. With an indexing coverage of 99 %, the findings
-  reveal that the DOI agency Crossref provides both comprehensive bibliographic coverage
-  of the funded open access journal literature and disambiguated names of journal
-  titles and publishing houses. We show that authority control of these bibliographic
-  information is particularly relevant for the comparative study of the economical
-  effects of open access publishing.
+
 ---
@@
 ## General Background{.unnumbered}
-
-<!-- Publication fees, often paid by funders or universities, are a widely discussed business model for open access journals.
-Yet, how and to what extent these activities are effective in terms of the number of supported research articles and associated costs remains under debate.
-This study contributes to a growing empirical base about institutional spending on open access publication fees.
-More specifically, it uses self-reported cost data from German universities and research organisations to examine how much they spent, and how the so funded open access journal articles were distributed over publishers and journal titles using bibliographic information from Crossref.
-Such an approach extends methods and improves data collection activities for researchers and practitioners alike, as well as contribute to a better understanding of factors affecting the analysis of publication fees in open access publishing.
--->
 

### 2016-06-06 article.Rmd
reviewed discussion

@@
 
-As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access Crossref's REST API[^25].
+We used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access Crossref's REST API[^25].
 We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the ISSN media types, the International Standard Serial Number used to identify journals, were distinguished.
@@
 Because Wellcome Trust's spending was reported for the fiscal periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
-We excluded articles from the analysis for which  information about the cost or the journal type was missing.
+We excluded articles from the analysis with missing information about the cost or the journal type.
 In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to 4.861.772 € from the analysis because no publication type was given in the data-set.
@@
 Data curation activities of the Open APC initiative and those of the other initiatives differed in some aspects.
-For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, in case of the Wellcome Trust spending data, DOIs were also being identified by automated compliance checks.
+For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, in case of the Wellcome Trust spending data, DOIs were additionally identified by automated compliance checks.
 Our first screening of the data-sets revealed that some articles published in Crossref-indexed journals lacked a DOI.
@@
 Figure 2 presents the distribution of institutional spending on publications.
-We observed that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100`%) of the publication fees were paid in accordance with the DFG price cap of 2,000 €.
+We observed that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100` %) of the publication fees were paid in accordance with the DFG price cap of 2,000 €.
 Most payment on publications ranged between 1,000 - 1,250 €.
@@
 With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall article volume.
-By contrast, for several universities that only recently begun to set up centrally managed open access funds to cover publication fees we observed a lower number of support for open access journal articles.
+By contrast, for several universities that only recently begun to set up centrally managed open access funds to cover publication fees we observed a lower number of supported open access journal articles.
 
@@
 They also demonstrate the growing institutional support for researchers to cover these charges in Germany [@Fournier_2013].
-Similar to the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending size varies across the German universities and research organisations.
+Similar to the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending sizes vary across the German universities and research organisations.
 With a proportion of 39 % of the total article volume, the Max Planck Society, a large non-university research organisation, supported most open access journal publications included in our study.
 A possible explanation is the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed open access agreements with publishers over the last decade on behalf of most Max Planck institutes [@Schimmer_2013, @Sikora_2015].
-This centralised approach presumably not only resulted in a large number of supported open access articles, but also advanced capabilities and skills needed to report on these expenditure on a regular basis.
+This centralised approach presumably not only resulted in a large number of supported open access articles, but also in central access to this data and in advanced capabilities and skills needed to report on these expenditure on a regular basis.
 Many universities and research organisations, on the other hand, disclosed a remarkably lower number of supported articles.
@@
 Beyond identifying publishers and journals, mandatory reporting of DOIs in the spending data can also increase the re-use of such data to study other aspects of open access journals using publication fees.
-For instance, impact analysis in the field of altmetrics are dependent on the availability of DOIs as well [@Haustein_2016].
+For instance, impact analyses in the field of altmetrics depend on the availability of DOIs as well [@Haustein_2016].
 Drawing on Crossref has the potential to increase the comparability of cost data to prepare future negotiations with publishers on open access agreements, because Crossref's metadata represents current developments in academic publishing in terms of ongoing mergers of publishing houses.
@@
 In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the articles published by their authors in SCOAP^3-covered journals.^[26] This is very arbitrary, since averages for an institution can only be determined after the end of a full 3-year funding cycle.
-Other factors affecting price variations are currency movements over the years, and different tax rates in Germany.
+Other factors affecting price variations are exchange rates and differend tax rates for some organisations in Germany.
 For instance, the Max Planck Society has a limited input tax reduction.
-The refund of input value added tax for publication fees is 20%.
-Local and payment currency, as well as taxes, as suggested by @Pinfield_2015 to increase the transparency of publication fee spending, were not disclosed in the Open APC initiative's data-set at the time of our study.
+The refund of input value added tax for publication fees is 20 %.
+Tax rates and the currency the payment was made in, as suggested by @Pinfield_2015 to increase the transparency of publication fee spending, were not disclosed in the Open APC initiative's data-set at the time of our study.
 In addition, the Open APC data-set does not track funding sources so that we cannot determine which funders co-financed publication fees.
@@
 
-Our analysis of how institutional spending per articles was distributed over publishers and journals indicates that open access publishing is heterogeneous and concentrated at the same time.
+Our analysis of how institutional spending for open access articles was distributed over publishers and journals indicates that open access publishing is heterogeneous and concentrated at the same time.
 While we were able to identify 139 individual publishing houses that were supported by the German universities and research organizations, the distribution is highly skewed.
@@
 However, our study could not confirm that publications in open access journals owned by traditional publishing houses account for most of the spending on publication fees as observed by @Pinfield_2015.
-Rather, open access publishers such as Public Library of Science (PloS), Copernicus or MPDI rank higher in our study than in the analyses of cost data in the UK.
+Rather, open access publishers such as Public Library of Science (PloS), Copernicus, or MPDI rank higher in our study than in the analyses of cost data in the UK.
 
@@
 99 % of all articles German universities and research organisations financially supported were published in fully open access journals.
-This presumably reflects the DFG funding policy that excludes the support of articles published in hybrid open access journals.
-However, while reviewing self-reported cost data from Austria and the UK where hybrid open access journals are generally supported, we observe a lower proportion of payment on articles in fully open access journals.
-Because open access publication fee spending is fragmented and often in-transparent, it remains open to speculation whether authors affiliated with German universities and research organisations avoid opting for open access when publishing in hybrid journals, or simply use other budgets that are not covered by the Open APC initiative.
+This presumably reflects the DFG funding policy that declares paying for open access articles published in hybrid journals undesirable.
+However, while reviewing self-reported cost data from Austria and the UK where hybrid open access journals are generally supported, we observe a much lower share of payments for articles in fully open access journals.
+Because APC spending is fragmented and often intransparent, it remains open to speculation whether authors affiliated with German universities and research organisations avoid opting for open access when publishing in hybrid journals, or simply use other budgets that are not covered by the Open APC initiative.
 

### 2016-06-06 article.Rmd
standardized white space: 5 %, 10 €...

@@
 In terms of articles, the majority of payments made were for publications in journals formerly associated with Springer Science + Business Media.
-Springer Science + Business Media journals accounted for `r nrow(springer)` articles, representing `r nrow(springer) / nrow(springer_nat) * 100`% of the overall Springer Nature article volume recorded by the Open APC initiative, and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % in terms of the amount being spent.
+Springer Science + Business Media journals accounted for `r nrow(springer)` articles, representing `r nrow(springer) / nrow(springer_nat) * 100` % of the overall Springer Nature article volume recorded by the Open APC initiative, and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % in terms of the amount being spent.
 Median publication fee spending slightly differed between Springer Science + Business Media (`r median(springer$euro)` €) and Nature Publishing (`r median(nature$euro)` €).

### 2016-06-06 article.Rmd
final minor fixes

@@
 
-Non-university research organisations, mainly those institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft, are not eligible for this DFG funding program.
-But in response, some organisations have adopted similar processes to support authors.
+Non-university research organisations, i.e.
+institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft, are not eligible for this DFG funding program.
+But in response, some have adopted similar processes to support authors.
 The Max-Planck-Gesellschaft operates their long-lasting open access activities, including handling spending and publisher agreements centrally, through the Max Planck Digital Library (MPDL) [@Schimmer_2013; @Sikora_2015], while the Leibniz-Gemeinschaft set up a dedicated open access fund in 2016.
 
-The evolving institutional support structures to cover open access publication fees has led to calls for an unified approach towards supporting open access journal publishing in Germany.
-The MPDL called for re-allocating subscriptions in favour of financing open access charges in 2014 [@Schimmer_Geschuhn_Vogler].
-The Allianz der Wissenschaftsorganisationen[^10c], a science policy board representing all major research organisations in Germany, marked price transparency as one way to sustain an "adequate open access publication system" [@allianz].
+The evolving institutional support to cover open access publication fees has led to calls for an unified approach towards open access funding in Germany.
+The MPDL called for re-allocating subscriptions in favour of open access  journals in 2015 [@Schimmer_Geschuhn_Vogler].
+At the same time, the Allianz der Wissenschaftsorganisationen[^10c], a science policy board representing all major research organisations in Germany, marked price transparency as one way to sustain an "adequate open access publication system" [@allianz].
 Reflecting Austrian and UK initiatives to share institutional spending on open access publication fees as open data, as well as professional discussions on open access publishing, Bielefeld University Library began to openly share its payment of publication fees in May 2014.
@@
 We analysed self-reported cost data released by the Open APC initiative on May 13, 2016,[^24] to assess institutional spending on open access publication fees in Germany.
-In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully open access journal or in a hybrid journal as well as the DOI reported in the data-set.
+In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully open access journal or in a hybrid journal, as well as the DOI reported in the data-set.
 
@@
 A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that operated as Springer Nature at the time of our study.
-To reflect these dynamics in academic publishing, we decided to retrieve  updated bibliographic metadata from Crossref, and to merge these records with the administrative information instead of re-using the historical publisher and journal information contained in the Open APC data-set.
+To reflect these dynamics in academic publishing, we decided to retrieve  updated bibliographic metadata from Crossref, and to merge these records with the administrative information rather than re-using the historical publisher and journal information contained in the Open APC data-set.
 
@@
 
-Data curation activities of the Open APC initiative and those of the other initiatives differed in some aspects.
+Data curation activities of the Open APC initiative and those of the other initiatives differed in some respects.
 For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, in case of the Wellcome Trust spending data, DOIs were additionally identified by automated compliance checks.
 Our first screening of the data-sets revealed that some articles published in Crossref-indexed journals lacked a DOI.
-Because of these different methods to curate the cost data, and as our main focus was institutional funding for publication fees in Germany, we decided to only compare German APC spending with that reported by other initiatives.
+Because of these different methods to curate the cost data, and as our main focus was institutional funding for publication fees in Germany, we decided to only compare German spending with that reported by other initiatives.
 We, therefore, did not analyse the distribution of spending over publishers and journal titles, or the indexing coverage in Crossref for the Austrian and UK spending data.
@@
 
-Among all articles, fees amounted to `r sum(my_apc$euro)` € including VAT, the average payment was `r mean(my_apc$euro)` € and the median value `r median(my_apc$euro)` € .
+Among all articles, fees amounted to `r sum(my_apc$euro)` € including value added tax, the average payment was `r mean(my_apc$euro)` € (Median =  `r median(my_apc$euro)` €) .
 Figure 2 presents the distribution of institutional spending on publications.
 We observed that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100` %) of the publication fees were paid in accordance with the DFG price cap of 2,000 €.
-Most payment on publications ranged between 1,000 - 1,250 €.
+Most payment on publication ranged between 1,000 - 1,250 €.
 
@@
 Large price variations could be observed.
-Publication fees that were paid by the German universities and research organisations ranged between `r min(my_apc$euro)` € and `r max(my_apc$euro)` €  (SD `r sd(my_apc$euro)`).
+Publication fees that were paid by the German universities and research organisations ranged between `r min(my_apc$euro)` € and `r max(my_apc$euro)` €  (SD = `r sd(my_apc$euro)`).
 However, the average price paid varied somewhat during the period 2011 and 2015 (`r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - `r my_apc %>% filter(period == 2015) %>% summarise(mean(euro)) %>% round()` €).
@@
 With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall article volume.
-By contrast, for several universities that only recently begun to set up centrally managed open access funds to cover publication fees we observed a lower number of supported open access journal articles.
+By contrast,  we observed a lower number of supported open access journal articles for several universities that only recently begun to set up centrally managed open access funds to cover publication fees.
 
@@
 The comparison revealed that the Open APC initiative lacked cost information about hybrid journals, whereas the related Austrian and UK open data initiatives reported a large share of spending on these journals between 2014 and 2015.
-Over the years 2005 – 2015, `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` German universities and research institutions reported `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` hybrid journal articles to the Open APC initiative, representing `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of the overall supported article volume.
+Over the years 2005 – 2015, `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` German universities and research institutions reported `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` hybrid journal articles to the Open APC initiative, representing `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of all articles included in the data-set.
 In terms of the number of supported articles and the amount being spent on publication fees, by contrast, the Open APC data-set provided the most comprehensive price information for fully open access journals compared to what the Austrian and UK initiatives had reported.
@@
 
-Comparison of average prices suggests that publishing in hybrid journal was more expensive than in fully open access journals.
+Comparison of average prices revealed that publishing in hybrid journals was, on average, more expensive than in fully open access journals.
 Price differences between these two categories were also reported earlier, indicating that prices for fully open access journals were on average lower [@Solomon_2012; @Pinfield_2015].
@@
 We used the DOI to automatically fetch publisher and journal names for each article from the Crossref REST API.
-Table 2 shows the top ten publishers in terms of payments made.
-They represented `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall spending on publication fees.
+Table 2 shows the top ten publishers in terms of the number of financially supported articles.
+These publishers represented `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of all articles include din our data-set.
 In total, payments were made to `r my_apc %>% group_by(publisher) %>% summarize(n()) %>% nrow()` publishing houses.
-In comparison with data from the UK, a greater share on total spending of some open access publishers could be observed.
+In comparison with data from the UK, we observed a greater share on total spending of some open access publishers.
 @Pinfield_2015, for instance, reported remarkably lower numbers for the open access publishers MPDI, Copernicus, and Hindawi.
@@
 Based on the number of articles paid for, Table 3 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals.
-We normalised PLOS journal titles, because the name chnages from "PLoS" to "PLOS" was only partly represented in the Crossref metadata at the time of our study.
-Payments to the top ten journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall expenditure.
-The multidisciplinary journal PLOS ONE and the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline, were also well represented in the Open APC data-set.
-In general, an estimated 14 out of more than 10,000 journals registered in DOAJ in 2015 accounted for up to 15–20 % of all articles published in fully open access journals [@Bj_rk_2015] In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
+We normalised PLOS journal titles, because the name changes from "PLoS" to "PLOS" was only partly represented in the Crossref metadata at the time of our study.
+Articles published in the top ten journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall article volume.
+The multidisciplinary journal PLOS ONE ranked highest.
+In addition, the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline, were also well represented in our study.
+In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
 
@@
 These findings are consistent with the general trend of publication fees as revenue source for open access publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-They also demonstrate the growing institutional support for researchers to cover these charges in Germany [@Fournier_2013].
-Similar to the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending sizes vary across the German universities and research organisations.
+They also demonstrate the growing institutional support for researchers in Germany to cover these charges [@Fournier_2013].
+Similar to the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending volume varies across the German universities and research organisations.
 With a proportion of 39 % of the total article volume, the Max Planck Society, a large non-university research organisation, supported most open access journal publications included in our study.
-A possible explanation is the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed open access agreements with publishers over the last decade on behalf of most Max Planck institutes [@Schimmer_2013, @Sikora_2015].
-This centralised approach presumably not only resulted in a large number of supported open access articles, but also in central access to this data and in advanced capabilities and skills needed to report on these expenditure on a regular basis.
+A possible explanation is the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed open access agreements with publishers over the last decade on behalf of most Max Planck institutes [@Schimmer_2013; @Sikora_2015].
+This centralised approach presumably not only resulted in a large number of supported open access articles, but also in central access to cost data provided by publishers, as well as in advanced capabilities and skills needed to report on these expenditure on a regular basis.
 Many universities and research organisations, on the other hand, disclosed a remarkably lower number of supported articles.
@@
 Beyond identifying publishers and journals, mandatory reporting of DOIs in the spending data can also increase the re-use of such data to study other aspects of open access journals using publication fees.
-For instance, impact analyses in the field of altmetrics depend on the availability of DOIs as well [@Haustein_2016].
+For instance, impact analyses in the field of altmetrics make heavy use of DOIs as well [@Haustein_2016].
 Drawing on Crossref has the potential to increase the comparability of cost data to prepare future negotiations with publishers on open access agreements, because Crossref's metadata represents current developments in academic publishing in terms of ongoing mergers of publishing houses.
@@
 One is that we cannot assess whether publishers and journals granted publication fee discounts.
-The Open APC initiative uses a minimal data scheme to incentives self-reporting, and, therefore, does not track this kind of information.
+The Open APC initiative uses a minimal data scheme to encourage self-reporting, and, therefore, does not track this kind of information.
 However, large price ranges suggests that different pricing levels and varying pricing schemes are in place, an observation also made earlier [@Solomon_2012; @Pinfield_2015].
@@
 Take for instance the journal Nature Communication.
-Charges reported to the Open APC initiative ranged between 2000 €, the DFG price cap, and 4.403 €.
+Charges reported to the Open APC initiative ranged between 2000 €, the DFG price cap, and 4.403 €, a journal that can be grouped as pricy high-impact journal according to @Solomon_2012.
 Although such payments being made from several budgets are a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro-rata payment were not made transparent in the Open APC data, leading to a possibly flawed representations of publication fee spending in Germany.
@@
 The refund of input value added tax for publication fees is 20 %.
-Tax rates and the currency the payment was made in, as suggested by @Pinfield_2015 to increase the transparency of publication fee spending, were not disclosed in the Open APC initiative's data-set at the time of our study.
+To increase the transparency of publication fee spending, @Pinfield_2015 generally suggested to disclose tax rates and the currency the payment was made in future cost data-sets.
 In addition, the Open APC data-set does not track funding sources so that we cannot determine which funders co-financed publication fees.
@@
 It must also be noted that reporting to the Open APC initiative is voluntary.
-Therefore, not all institutions in Germany that provide central funding of publication fees contribute cost data to this initiative.
+Therefore, not all institutions in Germany that provide central funding for publication fees contribute cost data to this initiative.
 In a qualitative survey, asking why German institutions are reluctant to share their cost data through the Open APC initiative, one institution feared that an increase in transparency would allow publishers to adjust prices in their favour.
@@
 Our analysis of how institutional spending for open access articles was distributed over publishers and journals indicates that open access publishing is heterogeneous and concentrated at the same time.
-While we were able to identify 139 individual publishing houses that were supported by the German universities and research organizations, the distribution is highly skewed.
+While we were able to identify 139 individual publishing houses that were supported by the German universities and research organisations, the distribution is highly skewed.
 Ten publishers collected 92 % of open access publication fee spending, being consistent with an observed high concentration of few publishers in current academic publishing [@Larivi_re_2015].
@@
 99 % of all articles German universities and research organisations financially supported were published in fully open access journals.
-This presumably reflects the DFG funding policy that declares paying for open access articles published in hybrid journals undesirable.
-However, while reviewing self-reported cost data from Austria and the UK where hybrid open access journals are generally supported, we observe a much lower share of payments for articles in fully open access journals.
-Because APC spending is fragmented and often intransparent, it remains open to speculation whether authors affiliated with German universities and research organisations avoid opting for open access when publishing in hybrid journals, or simply use other budgets that are not covered by the Open APC initiative.
+This presumably reflects the DFG funding policy that opposes paying for open access articles published in hybrid journals.
+However, while reviewing self-reported cost data from Austria and the UK where hybrid open access journals are generally supported, we observe a much higher share of payments for articles in hybrid open access journals.
+Because publication fee spending is fragmented and often in-transparent, it remains open to speculation whether authors affiliated with German universities and research organisations avoid opting for open access when publishing in hybrid journals, or simply use other budgets that are not covered by the Open APC initiative.
 
@@
 Our study reveals the size and extent of spending on open access journals using publication fees in Germany.
-Drawing on self-reported cost data from the Open APC initiative, payments from German universities and research institutions have grown over the years.
+Drawing on self-reported cost data from the Open APC initiative, support from German universities and research institutions have grown over the years.
 Comparing these expenditure with those from Austria and the UK, German open access funding is mostly focussed on fully open access journals, raising important questions about hybrid open access journals as publication venue.

### 2016-06-10 article.Rmd
proofreading introduction finished

@@
 author:
-- Najko Jahn, Bielefeld University Library, Bielefeld University (TIB) - German National
-  Library of Science and Technology, Hannover, Germany
-- <a href="http://orcid.org/0000-0002-5111-2788">Marco Tullney</a>, Technische Informationsbibliothek
+- Najko Jahn, Bielefeld University Library, Bielefeld University 
+- <a href="http://orcid.org/0000-0002-5111-2788">Marco Tullney</a>, TIB - German National Library of Science and Technology, Hannover, Germany
 date: "7 April 2016"
@@
 
-The rise of open access journals matches the increasing relevance of publication fees in academic publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-To cover these fees, also referred to as article-processing charges (APC), authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
-Yet, how and to what extent these research support activities are effective in terms of the number of supported articles and associated costs remains under debate.
+The rise of open access journals has been matched by the increasing relevance of publication fees in academic publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
+To cover these fees, also referred to as article-processing charges (APCs), authors tend to make use of funding that grant agencies or academic institutions provide [@Suber_2012].
+However, the question of how and to what extent these research support activities are effective in terms of the number of supported articles and their associated costs remains under debate.
 
 The study of institutional spending on open access journal articles has been limited for several reasons.
-One is that payment on these charges is fragmented across the budgets of grant agencies, research institutions, and libraries, or is covered by personal budgets.
-Asking 9,645 authors from various disciplines how they financed publication fees, a comprehensive survey in 2010 revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.
+The first is that payment of these charges is fragmented across the budgets of grant agencies, research institutions, and libraries, or is covered by personal budgets.
+A comprehensive 2015 survey asked 9,645 authors from various disciplines how they financed publication fees, and it revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.
 By contrast, 12 % paid publication fees individually [@soap].
-These findings are consistent with that of other studies, adding that sources of funding mostly exists in higher income countries, mainly to support research articles in the bio- and physical sciences [@Solomon_2011].
-Personal budgets, on the other hand, are likely used to cover low price publication fees [@Solomon_2011; @Bj_rk_2015].
+These results are consistent with similar findings from other studies.
+Previous studies also found that funding sources exist primarily in higher-income countries, mainly to support research articles in the biological and physical sciences [@Solomon_2011].
+Personal budgets, however, are likely used to cover lower publication fees [@Solomon_2011; @Bj_rk_2015].
 
-Another key problem in this regard is that funding for open access journals using publication fees lacks transparency because the parties involved - authors, universities, funders, publishers - neither release information on who pays for what nor the costs of publishing [@Bj_rk_learned_2014], a situation similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
-Empirical studies examining publication fees so far obtained price estimates by surveying authors [@soap], or from journal websites.
-Using the latter method, two studies investigating journals across a broad range of disciplines calculated similar price averages that ranged between 904 $ [@Solomon_2012] and 923 $ [@Walters_2011], as well as considerable price variances across journals and publishers.
-Accordingly, @Solomon_2012 suggested to cluster fully open access journal using publication fees into several groups.
+Another key problem in this regard is that funding for open access journals using publication fees lacks transparency because the parties involved -- authors, universities, funders, and publishers -- do not  release information about who pays for what or the costs of publishing [@Bj_rk_learned_2014]; this situation is similar to the lack of transparency regarding journal subscriptions [@Lawson_Meghreblian_2015].
+To date, empirical studies examining publication fees have obtained price estimates by surveying authors [@soap] or obtained them from journal websites.
+Using the latter method, two studies investigating journals across a broad range of disciplines calculated similar price averages that ranged between $ 904 [@Solomon_2012] and $ 923 [@Walters_2011], as well as considerable price variation across journals and publishers.
+Accordingly, @Solomon_2012 suggested using publication fees to cluster fully open access journal into several groups.
 In descending order, these are high-impact journals, followed by biomedicine journals from commercial publishers, large multi-disciplinary journals, and mid-price journals from commercial publishers covering a large spectrum of disciplines.
-Lower priced journals were published by academic societies and by publishers from low-income countries.
+Lower-priced journals are those published by academic societies and by publishers from low-income countries.
 
@@
 
-Hybrid journals substantially add to this complexity of open access funding [@Bj_rk_learned_2014; @Kingsley_2014; @Pinfield_2015].
-These journals, allowing articles to be published immediately as open access after a charge was paid, rely both on subscriptions and publication fees as revenue sources.
-Although the uptake of open access through hybrid journals was described as lower and more expensive compared to that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014], this model has gained increasing attention because of recent open access science policies, notably from the UK [@Pinfield_Review_2015].
+Hybrid journals substantially add to the complexity of open access funding [@Bj_rk_learned_2014; @Kingsley_2014; @Pinfield_2015].
+These journals, which allow articles to be published immediately as open access after a charge is paid, rely on both subscriptions and publication fees as revenue sources.
+Although the uptake of open access through hybrid journals was described as lower and more expensive than that of fully open access journals [@Solomon_2012; @Bj_rk_learned_2014], this model has gained increasing attention because of recent open access science policies, notably from the UK [@Pinfield_Review_2015].
 
-To address these problems of fragmented spending on publication fees and of in-transparency about what was being paid, some European research funders and research performing institutions have recently begun to disclose their payment for publication as open data.
-To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc_12] and the Austrian Science Fund FWF [@fwf_apc_13].
-The not-for-profit company Jisc followed by collecting data from UK universities [@Lawson_data_2015].
-Disclosed as publicly available spreadsheets, these data-sets self-report expenditure along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version.
-Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers to the full text including the Digital Object Identifier (DOI) [@Neylon_2014; @woodward_2014].
-Parts of Jisc's cost data was examined by @Pinfield_2015.
-Although the average spending on publication fees remained stable across the universities, they found large price variances, as well as a varying number of articles UK universities supported between 2007 - 2014, being consistent with earlier studies that collected price information from journal websites [@Solomon_2012].
+To address the problems of fragmented spending on publication fees and the lack of transparency about what is being paid, some European research funders and research-performing institutions have recently begun to disclose their expenditures for publication as open data.
+To the best of our knowledge, the first research funders to provide such data were the Wellcome Trust [@wellcome_apc_12] and the Austrian Science Fund FWF [@fwf_apc_13].
+The not-for-profit company Jisc followed this example by collecting data from UK universities [@Lawson_data_2015].
+Disclosed as publicly available spreadsheets, these data-sets self-report expenditures along with bibliographic information, including title, journal and publisher, and a persistent identifier for the publisher's version.
+Curatorial efforts focussed on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers to the full text including the Digital Object Identifier (DOI) [@Neylon_2014; @woodward_2014].
+A preliminary version of Jisc's cost data was examined by @Pinfield_2015.
+Although the average spending on publication fees remained stable across universities, the authors found large price variations, as well as a varying number of articles supported by UK universities between 2007 and 2014, findings which confirm earlier studies that collected price information from journal websites [@Solomon_2012].
 
@@
 
-This paper focuses on how much German universities and research organisations spent on open access publication fees.
+This paper focuses on how much German universities and research organisations spend on open access publication fees.
 In Germany, the Deutsche Forschungsgemeinschaft (DFG), the largest German research funder, has strongly influenced how universities manage institutional support for these charges.
-Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" program in 2011, only a few central funds existed [@eppelin_2012].
-This is similar to the situation described in Canada [@Hampson_2014] or the UK [@PINFIELD_2012].
-The DFG has enforced a set of criteria grantees have to comply with  [@Fournier_2013], leading to related implementations for open access supporting across German universities.
-These criteria exclude sponsoring of articles in hybrid journals, and the funding of articles whose publication fee exceeds 2,000 €[^3].
+Before the DFG started to pay for centrally funded publication fees on a pro rata basis through its "Open-Access Publishing" programme in 2011, only a few central funds existed [@eppelin_2012].
+This is similar to the situation described in Canada [@Hampson_2014] and the UK [@PINFIELD_2012].
+The DFG has enforced a set of criteria with which grantees have to comply and which has resulted in similar policies regarding support for publication fees across German universities [@Fournier_2013].
+These criteria exclude the sponsorship of articles in hybrid journals and the funding of articles for which the publication fee exceeds € 2,000[^3].
 Grantees agree not only to pay for APCs, but also to find ways to improve the handling of those financial transactions.
-This includes central invoicing schemes and memberships, among others, agreed between university libraries and publishers that often lead to a discount on publication fees [@Fournier_2013].
+These ways include central invoicing schemes and memberships that are agreed upon by university libraries and publishers and that often lead to a discount on publication fees [@Fournier_2013].
 
-Non-university research organisations, i.e.
-institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft, are not eligible for this DFG funding program.
-But in response, some have adopted similar processes to support authors.
-The Max-Planck-Gesellschaft operates their long-lasting open access activities, including handling spending and publisher agreements centrally, through the Max Planck Digital Library (MPDL) [@Schimmer_2013; @Sikora_2015], while the Leibniz-Gemeinschaft set up a dedicated open access fund in 2016.
+Non-university research organisations, i.e., institutes organised in the Fraunhofer-Gesellschaft, Helmholtz-Gemeinschaft, Leibniz-Gemeinschaft, and Max-Planck-Gesellschaft, are not eligible for this DFG funding programme.
+However, in response, some organisations have adopted similar processes to support authors.
+The Max-Planck-Gesellschaft operates its long-standing open access activities, including handling spending and publisher agreements centrally, through the Max Planck Digital Library (MPDL) [@Schimmer_2013; @Sikora_2015], while the Leibniz-Gemeinschaft set up a dedicated open access fund in 2016.
 
-The evolving institutional support to cover open access publication fees has led to calls for an unified approach towards open access funding in Germany.
+The evolving institutional support for covering open access publication fees has led to calls for a unified approach towards open access funding in Germany.
 The MPDL called for re-allocating subscriptions in favour of open access  journals in 2015 [@Schimmer_Geschuhn_Vogler].
@@
 
-The aim of the study was to examine how much German universities and research organisations spent on open access publication fees until 2015.
-Using self-reported cost data from the Open APC initiative, the analysis focused on the amount that was being spent on publication fees, and compared these expenditure with data from related Austrian and UK initiatives, both in terms of size and the proportion of articles being published in fully and hybrid open access journals.
-We also asked how thoroughly self-reported articles were indexed in Crossref, a DOI minting agency for scholarly literature, and analysed how the institutional spending was distributed over publishers and journal titles.
+The aim of this study was to examine how much German universities and research organisations spent on open access publication fees.
+Using self-reported cost data from the Open APC initiative, the analysis focussed on the amount that was being spent on publication fees, and compared these expenditure with data from related Austrian and UK initiatives, in terms  of both size and the proportion of articles being published in fully and hybrid open access journals.
+We also investigated how thoroughly self-reported articles were indexed in Crossref, a DOI minting agency for scholarly literature, and analysed how the institutional spending was distributed across publishers and journal titles.
 

### 2016-06-10 article.Rmd
proofreading abstract

@@
 - Najko Jahn, Bielefeld University Library, Bielefeld University 
-- <a href="http://orcid.org/0000-0002-5111-2788">Marco Tullney</a>, TIB - German National Library of Science and Technology, Hannover, Germany
-date: "7 April 2016"
+- Marco Tullney, German National Library of Science and Technology (TIB), Hannover, Germany
+date: "10 June 2016"
 output:

### 2016-06-10 article.Rmd
proofreading methods section

@@
 We analysed self-reported cost data released by the Open APC initiative on May 13, 2016,[^24] to assess institutional spending on open access publication fees in Germany.
-In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully open access journal or in a hybrid journal, as well as the DOI reported in the data-set.
+In addition to administrative data about the amount paid per article, including value-added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully open access journal or in a hybrid journal, as well as the DOI reported in the data-set.
 
-We fetched bibliographic metadata for each article from Crossref on May 19, 2016, on the basis of the reported DOIs.
+We obtained bibliographic metadata for each article from Crossref on May 19, 2016, on the basis of the reported DOIs.
 Although the Open APC initiative gathered metadata representing publishers and journals from Crossref as well, this information was retrieved at the time when the participating institutions submitted their expenditure.
-The Open APC initiative kept track of the date when these data-sets were submitted with Git, a version control system, increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub to be transparent over time.
-However, during these data collection activities Crossref had regularly updated metadata to represent ongoing mergers of publishing houses.
-A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that operated as Springer Nature at the time of our study.
-To reflect these dynamics in academic publishing, we decided to retrieve  updated bibliographic metadata from Crossref, and to merge these records with the administrative information rather than re-using the historical publisher and journal information contained in the Open APC data-set.
+To be transparent over time, the Open APC initiative kept track of the date when these data-sets were submitted with Git, a version control system that is increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub  However, during these data collection activities, Crossref regularly updated the metadata to represent ongoing mergers of publishing houses.
+A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, which operated as Springer Nature at the time of our study.
+To reflect these dynamics in academic publishing, we decided to retrieve  updated bibliographic metadata from Crossref and to merge these records with the administrative information rather than re-using the historical publisher and journal information contained in the Open APC data-set.
 
-We used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access Crossref's REST API[^25].
-We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the ISSN media types, the International Standard Serial Number used to identify journals, were distinguished.
-It also contained normalised publisher information, thus avoiding confusion about naming of publishing houses other studies were faced with when working with self-reported data [@woodward_2014].
+We used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access Crossref's REST API.[^25] We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the ISSN media types (the International Standard Serial Number used to identify journals) were distinguished.
+This source also contained normalised publisher information, thus avoiding confusion regarding the naming of publishing houses that other studies faced when working with self-reported data [@woodward_2014].
 In cases where no bibliographic information could be obtained, we used the Open APC values.
-Because Crossref was not the only registration agency for DOIs, but also the agencies DataCite and Medra minted DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
+Because Crossref is not the only registration agency for DOIs – the agencies DataCite and Medra also mint DOIs for scholarly works – we also identified the DOI agency for each article with the help of the rcrossref client.
 
 Data collection also involved obtaining cost data from related open data initiatives.
-To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we consulted the openly available data-sets from the the Austrian Science Fund FWF [@fwf_apc_14; @fwf_apc_15], Jisc [@jisc_14; @jisc_15] and the Wellcome Trust [@wellcome_apc_13_14; @wellcome_apc_14_15].
+To compare self-reported spending on open access journal articles by German universities and research organisations with that of other initiatives, we consulted the openly available data-sets from the Austrian Science Fund (FWF) [@fwf_apc_14; @fwf_apc_15], Jisc [@jisc_14; @jisc_15] and the Wellcome Trust [@wellcome_apc_13_14; @wellcome_apc_14_15].
 For analysis, we obtained the overall publication fee spending on both fully and hybrid open access journal articles.
@@
 We used the spreadsheet data to calculate Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average foreign exchange reference rates provided by the European Central Bank.
-Our comparison of the open data initiatives focussed on the last two years 2014 and 2015.
-Because Wellcome Trust's spending was reported for the fiscal periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
+Our comparison of the open data initiatives focussed on the last two years: 2014 and 2015.
+Because Wellcome Trust's spending was reported for the fiscal periods 2013-2014 and 2014-2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data.
 We excluded articles from the analysis with missing information about the cost or the journal type.
-In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to 4.861.772 € from the analysis because no publication type was given in the data-set.
+In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to € 4.861.772 from the analysis because no publication type was given in the data-set.
 
 Data curation activities of the Open APC initiative and those of the other initiatives differed in some respects.
-For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, in case of the Wellcome Trust spending data, DOIs were additionally identified by automated compliance checks.
+For instance, the DOI was a mandatory element in the Open APC data template that the participating institutions were required to report, whereas in the case of the Wellcome Trust spending data, DOIs were additionally identified by automated compliance checks.
 Our first screening of the data-sets revealed that some articles published in Crossref-indexed journals lacked a DOI.
-Because of these different methods to curate the cost data, and as our main focus was institutional funding for publication fees in Germany, we decided to only compare German spending with that reported by other initiatives.
-We, therefore, did not analyse the distribution of spending over publishers and journal titles, or the indexing coverage in Crossref for the Austrian and UK spending data.
+Because of these different methods to curate the cost data and because our main focus was institutional funding for publication fees in Germany, we decided to compare only German spending with that reported by other initiatives.
+We did not, therefore, analyse the distribution of spending over publishers and journal titles or the indexing coverage in Crossref for the Austrian and UK spending data.
 

### 2016-06-10 article.Rmd
proofreading results section

@@
 
-After excluding payments being made for non-journal articles as well as articles invoiced in 2016, we retrieved `r nrow(my_apc)` open access journal articles `r length(unique(my_apc$institution))` German universities and research institutions financially supported between 2005 and 2015.
+After excluding payments for non-journal articles as well as articles invoiced in 2016, we retrieved information on `r nrow(my_apc)` open access journal articles that `r length(unique(my_apc$institution))` German universities and research institutions supported financially between 2005 and 2015.
 As illustrated in Figure 1, payments made for open access journal articles increased over the years.
 While one institution supported `r filter(my_apc, period == min(my_apc$period)) %>% nrow()` articles in `r as.character(min(my_apc$period))`, most institutions included in our study shared their expenditure from 2013 onwards.
-With `r max(apc_time$articles)` articles, the year `r apc_time[apc_time$articles == max(apc_time$articles),"period"]` was best represented.
-However, at the time of analysis, not all, but `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institutions contributed cost information, suggesting a time lag between payments made and reporting these spending to the Open APC initiative.
+The best represented year was `r apc_time[apc_time$articles == max(apc_time$articles),"period"]`, with `r max(apc_time$articles)` articles.
+However, at the time of analysis only `r my_apc %>% filter(period == 2015) %>% group_by(institution) %>% summarise(n()) %>% nrow()` institutions had contributed cost information for 2015, suggesting a lag between the time that payments are made and expenditures are reported to the Open APC initiative.
 
-**Figure 1: Growth of Open APC Initiative**
+**Figure 1: Growth of the Open APC Initiative**
 
@@
 
-Among all articles, fees amounted to `r sum(my_apc$euro)` € including value added tax, the average payment was `r mean(my_apc$euro)` € (Median =  `r median(my_apc$euro)` €) .
+The fees for all of the articles amounted to € `r sum(my_apc$euro)`, including value-added tax; the average payment was € `r mean(my_apc$euro)` (median =  € `r median(my_apc$euro)`) .
 Figure 2 presents the distribution of institutional spending on publications.
-We observed that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100` %) of the publication fees were paid in accordance with the DFG price cap of 2,000 €.
-Most payment on publication ranged between 1,000 - 1,250 €.
+We observed that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100` %) of the publication fees were paid in accordance with the DFG price cap of € 2,000.
+Most payments for publications ranged from € 1,000 to € 1,250.
 
@@
 Figure 3 presents institutional spending per article and year.
-Large price variations could be observed.
-Publication fees that were paid by the German universities and research organisations ranged between `r min(my_apc$euro)` € and `r max(my_apc$euro)` €  (SD = `r sd(my_apc$euro)`).
-However, the average price paid varied somewhat during the period 2011 and 2015 (`r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - `r my_apc %>% filter(period == 2015) %>% summarise(mean(euro)) %>% round()` €).
+Large price variations can be observed.
+Publication fees that were paid by German universities and research organisations ranged from € `r min(my_apc$euro)` to € `r max(my_apc$euro)` (SD = `r sd(my_apc$euro)`).
+However, the average price paid varied somewhat during the period from 2011 to 2015 (€ `r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - € `r my_apc %>% filter(period == 2015) %>% summarise(mean(euro)) %>% round()`).
 
@@
 
-The number of APC payments per institutions varied considerably (see Table 1).
+The number of APC payments per institution varied considerably (see Table 1).
 With `r filter(my_apc, institution == "MPG") %>% nrow()` reported articles, the Max Planck Society contributed `r (filter(my_apc, institution == "MPG") %>% nrow() / nrow(my_apc)) * 100` % of the overall article volume.
-By contrast,  we observed a lower number of supported open access journal articles for several universities that only recently begun to set up centrally managed open access funds to cover publication fees.
+By contrast,  we observed a lower number of supported open access journal articles for several universities that had only recently begun to set up centrally managed open access funds to cover publication fees.
 
@@
 
-## Comparision of related cost data-sets
+## Comparison of related cost data-sets
 
-Table 2 compares Open APC spending data with that of the Austrian FWF, as well as with Jisc's and Wellcome Trust's expenditure.
-Prices were converted according to the average Euro exchange rate of the examined periods, and gathered for both fully and hybrid open access journals.
+Table 2 compares the Open APC spending data with that of the Austrian FWF, as well as with Jisc's and Wellcome Trust's expenditures.
+Prices were converted according to the average Euro exchange rate during the examined periods and were gathered for both fully open access journals and hybrid journals.
 The comparison revealed that the Open APC initiative lacked cost information about hybrid journals, whereas the related Austrian and UK open data initiatives reported a large share of spending on these journals between 2014 and 2015.
-Over the years 2005 – 2015, `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` German universities and research institutions reported `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` hybrid journal articles to the Open APC initiative, representing `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of all articles included in the data-set.
-In terms of the number of supported articles and the amount being spent on publication fees, by contrast, the Open APC data-set provided the most comprehensive price information for fully open access journals compared to what the Austrian and UK initiatives had reported.
+Over the years 2005-2015, `r my_apc %>% group_by(institution) %>% filter(is_hybrid == TRUE) %>% summarise() %>% nrow()` out of `r length(unique(my_apc$institution))` German universities and research institutions reported `r my_apc %>% filter(is_hybrid == TRUE) %>% nrow()` hybrid journal articles to the Open APC initiative, representing `r (my_apc %>% filter(is_hybrid == TRUE) %>% nrow() / my_apc %>% nrow()) * 100` % of all articles included in the data-set.
+In contrast, in terms of the number of supported articles and the amount spent on publication fees, the Open APC data-set provided more comprehensive price information for fully open access journals than did the Austrian and UK initiatives.
 
@@
 
-Comparison of average prices revealed that publishing in hybrid journals was, on average, more expensive than in fully open access journals.
-Price differences between these two categories were also reported earlier, indicating that prices for fully open access journals were on average lower [@Solomon_2012; @Pinfield_2015].
-In 2014 and 2015, the mean price for fully open access journals calculated from all data-sets was below the DFG price cap of 2,000 €.
+A comparison of average prices revealed that publishing in hybrid journals was, on average, more expensive than publishing in fully open access journals.
+Price differences between these two categories were also reported earlier, indicating that prices for fully open access journals were lower on average [@Solomon_2012; @Pinfield_2015].
+In 2014 and 2015, the mean price for fully open access journals calculated from all data-sets was below the DFG price cap of € 2,000.
 
@@
 agencies <- data.frame(table(unlist(doi_miss$doi_agency)))
-
 ```
 
-To identify publication fee spending on the article-level, as well as to gather bibliographic metadata, DOIs were a mandatory part of the Open APC initiative's data collection activities.
+To identify publication fee spending at article level and to gather bibliographic metadata, DOIs were a mandatory part of the Open APC initiative's data collection activities.
 The participating institutions reported DOIs for `r my_apc %>% filter(!is.na(doi)) %>% nrow()` out of `r nrow(my_apc)` articles.
@@
 For eight articles, our parser could not gather the XML resource, although these publications were registered with Crossref at the time of our study.
-Seven DOIs reported to the Open APC initiative did not resolve.
+Seven DOIs reported to the Open APC initiative could not be resolved.
 
@@
 Table 2 shows the top ten publishers in terms of the number of financially supported articles.
-These publishers represented `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of all articles include din our data-set.
+These publishers represented `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of all articles included in our data-set.
 In total, payments were made to `r my_apc %>% group_by(publisher) %>% summarize(n()) %>% nrow()` publishing houses.
-In comparison with data from the UK, we observed a greater share on total spending of some open access publishers.
+Comparing these data with data from the UK, we observed that a greater share of total spending was directed to some open access publishers.
 @Pinfield_2015, for instance, reported remarkably lower numbers for the open access publishers MPDI, Copernicus, and Hindawi.
@@
 
-Most of publication fee spending in Germany was on articles published in Springer Nature journals, likely profiting from the merge with the open access publisher BioMed Central in 2008, and that between the well-established publishers Springer Science + Business Media and Nature Publishing Group in 2015.
-Using the Crossref-Member-ID instead of the publisher name, we were able to differentiate between journals formely published by Springer Science + Business Media and Nature Publishing Group.
+Most of the publication fee spending in Germany was on articles published in Springer Nature journals, which likely reflects the results of mergers with the open access publisher BioMed Central in 2008 and between the well-established publishers Springer Science + Business Media and Nature Publishing Group in 2015.
+Using the Crossref-Member-ID instead of the publisher name, we were able to differentiate between journals formerly published by Springer Science + Business Media and Nature Publishing Group.
 In terms of articles, the majority of payments made were for publications in journals formerly associated with Springer Science + Business Media.
-Springer Science + Business Media journals accounted for `r nrow(springer)` articles, representing `r nrow(springer) / nrow(springer_nat) * 100` % of the overall Springer Nature article volume recorded by the Open APC initiative, and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % in terms of the amount being spent.
-Median publication fee spending slightly differed between Springer Science + Business Media (`r median(springer$euro)` €) and Nature Publishing (`r median(nature$euro)` €).
-However, price variation was higher for Nature Publishing journals (SD = `r sd(nature$euro)`€) compared with that of the former Springer Science + Business Media titles (SD = `r sd(springer$euro)` €).
+Springer Science + Business Media journals accounted for `r nrow(springer)` articles, representing `r nrow(springer) / nrow(springer_nat) * 100` % of the overall Springer Nature article volume recorded by the Open APC initiative and `r sum(springer$euro) / sum(springer_nat$euro) * 100` % of the amount that was spent.
+Median publication fee spending differed slightly between Springer Science + Business Media (€ `r median(springer$euro)` €) and Nature Publishing (€ `r median(nature$euro)`).
+However, the price variation was higher for Nature Publishing journals (SD = € `r sd(nature$euro)`) than for the former Springer Science + Business Media titles (SD = € `r sd(springer$euro)`).
 
@@
 Prices also varied within journals.
-Based on the number of articles paid for, Table 3 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals.
-We normalised PLOS journal titles, because the name changes from "PLoS" to "PLOS" was only partly represented in the Crossref metadata at the time of our study.
-Articles published in the top ten journals represent `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall article volume.
+Table 3 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals, based on the number of supported articles.
+We normalised PLOS journal titles because the name change from "PLoS" to "PLOS" was only partially represented in the Crossref metadata at the time of our study.
+Articles published in the top ten journals represented `r apc_journal %>% filter(!journal_full_title == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of the overall article volume.
 The multidisciplinary journal PLOS ONE ranked highest.
 In addition, the journals New Journal of Physics, Atmospheric Chemistry and Physics Discussions and Frontiers in Psychology, all of which publish contributions from all branches of their respective discipline, were also well represented in our study.
-In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and also takes the submission's file format into consideration.
+In the case of Atmospheric Chemistry and Physics Discussions, the large price range can be explained by the fact that this journal charges per page and takes the submission's file format into consideration.
 

### 2016-06-10 article.Rmd
proofreading discussions section

@@
 In Germany, institutional spending on publication fees charged by open access journals has increased over the years.
-These findings are consistent with the general trend of publication fees as revenue source for open access publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-They also demonstrate the growing institutional support for researchers in Germany to cover these charges [@Fournier_2013].
-Similar to the expenditure on publication fees at an institutional level in the UK [@Pinfield_2015], spending volume varies across the German universities and research organisations.
-With a proportion of 39 % of the total article volume, the Max Planck Society, a large non-university research organisation, supported most open access journal publications included in our study.
-A possible explanation is the centralised library support at the Max Planck Society where the Max Planck Digital Library has managed open access agreements with publishers over the last decade on behalf of most Max Planck institutes [@Schimmer_2013; @Sikora_2015].
-This centralised approach presumably not only resulted in a large number of supported open access articles, but also in central access to cost data provided by publishers, as well as in advanced capabilities and skills needed to report on these expenditure on a regular basis.
-Many universities and research organisations, on the other hand, disclosed a remarkably lower number of supported articles.
+These findings are consistent with the general trend towards using publication fees as a revenue source for open access publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
+They also demonstrate the growing trend among institutions to both encourage their researchers to publish in open access journals and to offer financial support [@Fournier_2013].
+Similar to the expenditures on publication fees at an institutional level in the UK [@Pinfield_2015], spending volume varies across German universities and research organisations.
+With a proportion of 39 % of the total article volume, the Max Planck Society, a large non-university research organisation, supported the most open access journal publications included in our study.
+A possible explanation could be the centralised library support at the Max Planck Society, where the Max Planck Digital Library has managed open access agreements with publishers over the last decade on behalf of most Max Planck institutes [@Schimmer_2013; @Sikora_2015].
+This centralised approach presumably resulted not only in a large number of supported open access articles but also in central access to cost data provided by publishers, as well as in posessing the advanced capabilities and skills needed to report these expenditures on a regular basis.
+Many universities and research organisations, by contrast, disclosed a remarkably lower number of supported articles.
 
-To answer data curation issues while unifying UK spending data [@Neylon_2014; @woodward_2014], re-using DOIs to gather bibliographic metadata from Crossref is a promising approach.
+Re-using DOIs to gather bibliographic metadata from Crossref is a promising approach to addressing data curation issues while unifying UK spending data [@Neylon_2014; @woodward_2014].
 In our study, Crossref thoroughly indexed open access journal articles disclosed in the Open APC data-set, providing information about publisher and journal titles for 99 % of all articles included in the Open APC data-set.
-Making use of metadata from Crossref, therefore, reduces extensive validation work of bibliographic records as long as the DOIs are made available in the cost data.
-Beyond identifying publishers and journals, mandatory reporting of DOIs in the spending data can also increase the re-use of such data to study other aspects of open access journals using publication fees.
+Making use of metadata from Crossref, therefore, reduces the need for extensive validation of bibliographic records as long as the DOIs are made available in the cost data.
+Beyond identifying publishers and journals, mandatory reporting of DOIs in the spending data can also increase the use of such data to study other aspects of APC-based open access journals.
 For instance, impact analyses in the field of altmetrics make heavy use of DOIs as well [@Haustein_2016].
-Drawing on Crossref has the potential to increase the comparability of cost data to prepare future negotiations with publishers on open access agreements, because Crossref's metadata represents current developments in academic publishing in terms of ongoing mergers of publishing houses.
-In addition to these practical consequences, future comparative studies of publication fee spending using data on the article-level can also benefit from such an approach.
+Drawing on Crossref has the potential to increase the comparability of cost data to prepare for future negotiations with publishers regarding open access agreements because Crossref's metadata represent current developments in academic publishing in terms of ongoing mergers of publishing houses.
+In addition to these practical benefits, future comparative studies of publication fee spending using data at the article level can also benefit from such an approach.
 
 This study is limited in some respects.
-One is that we cannot assess whether publishers and journals granted publication fee discounts.
-The Open APC initiative uses a minimal data scheme to encourage self-reporting, and, therefore, does not track this kind of information.
-However, large price ranges suggests that different pricing levels and varying pricing schemes are in place, an observation also made earlier [@Solomon_2012; @Pinfield_2015].
-Adding to this complexity, it is likely that some institutions only paid parts of the publication fee.
-Take for instance the journal Nature Communication.
-Charges reported to the Open APC initiative ranged between 2000 €, the DFG price cap, and 4.403 €, a journal that can be grouped as pricy high-impact journal according to @Solomon_2012.
-Although such payments being made from several budgets are a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro-rata payment were not made transparent in the Open APC data, leading to a possibly flawed representations of publication fee spending in Germany.
-In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the articles published by their authors in SCOAP^3-covered journals.^[26] This is very arbitrary, since averages for an institution can only be determined after the end of a full 3-year funding cycle.
-Other factors affecting price variations are exchange rates and differend tax rates for some organisations in Germany.
+First, we cannot assess whether publishers and journals granted publication fee discounts.
+The Open APC initiative uses a minimal data scheme to encourage self-reporting, and it therefore does not track this type of information.
+However, large price variations suggest that different pricing levels and pricing schemes are in place, as previously observed [@Solomon_2012; @Pinfield_2015].
+Adding to this complexity, it is likely that some institutions paid only part of the publication fee.
+Consider, for instance, the journal Nature Communication: Charges reported to the Open APC initiative ranged between € 2000, the DFG price cap, and € 4,403 for a journal that can be categorised as a pricy, high-impact journal according to @Solomon_2012.
+Although making such payments from several budgets is a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro-rata payments were not made transparent in the Open APC data, leading to a possibly flawed representation of publication fee spending in Germany.
+In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the number of articles published by their authors in SCOAP^3-covered journals.[^26] This approach is arbitrary, because averages for an institution can be determined only after the end of a full 3-year funding cycle.
+Other factors affecting price variations are exchange rates and different tax rates for some organisations in Germany.
 For instance, the Max Planck Society has a limited input tax reduction.
-The refund of input value added tax for publication fees is 20 %.
-To increase the transparency of publication fee spending, @Pinfield_2015 generally suggested to disclose tax rates and the currency the payment was made in future cost data-sets.
-In addition, the Open APC data-set does not track funding sources so that we cannot determine which funders co-financed publication fees.
+The refund of input value-added tax for publication fees is 20 %.
+To increase the transparency of publication fee spending, @Pinfield_2015 suggested disclosing tax rates and payment currencies in future cost data-sets.
+In addition, the Open APC data-set does not track funding sources; thus, we could not determine which funders co-financed publication fees.
 
@@
 Therefore, not all institutions in Germany that provide central funding for publication fees contribute cost data to this initiative.
-In a qualitative survey, asking why German institutions are reluctant to share their cost data through the Open APC initiative, one institution feared that an increase in transparency would allow publishers to adjust prices in their favour.
-Others pointed out that the workload to produce such a data-set could be too extensive [@deppe_2015].
+According to a qualitative survey that asked why German institutions are reluctant to share their cost data through the Open APC initiative, one institution feared that an increase in transparency would allow publishers to adjust prices in their favour.
+Others noted that the workload to produce such a data-set could be too extensive [@deppe_2015].
 As no reliable registry of institutional open access funds or related support structures in Germany exists, we cannot assess the number of non-participants in Germany.
 
-Our analysis of how institutional spending for open access articles was distributed over publishers and journals indicates that open access publishing is heterogeneous and concentrated at the same time.
+Our analysis of how institutional spending for open access articles was distributed over publishers and journals indicated that open access publishing is heterogeneous and concentrated at the same time.
 While we were able to identify 139 individual publishing houses that were supported by the German universities and research organisations, the distribution is highly skewed.
-Ten publishers collected 92 % of open access publication fee spending, being consistent with an observed high concentration of few publishers in current academic publishing [@Larivi_re_2015].
-However, our study could not confirm that publications in open access journals owned by traditional publishing houses account for most of the spending on publication fees as observed by @Pinfield_2015.
-Rather, open access publishers such as Public Library of Science (PloS), Copernicus, or MPDI rank higher in our study than in the analyses of cost data in the UK.
+Ten publishers collected 92 % of open access publication fee spending, which is consistent with an observed high concentration of a few publishers in current academic publishing [@Larivi_re_2015].
+However, our study could not confirm that publications in open access journals owned by traditional publishing houses accounted for most of the spending on publication fees as observed by @Pinfield_2015.
+Rather, open access publishers such as Public Library of Science (PloS), Copernicus, or MPDI ranked higher in our study than they did in the analyses of cost data in the UK.
 
-One possible explanation why traditional publishers are less well represented in our study is the lack of cost information about hybrid open access journals.
-99 % of all articles German universities and research organisations financially supported were published in fully open access journals.
-This presumably reflects the DFG funding policy that opposes paying for open access articles published in hybrid journals.
-However, while reviewing self-reported cost data from Austria and the UK where hybrid open access journals are generally supported, we observe a much higher share of payments for articles in hybrid open access journals.
-Because publication fee spending is fragmented and often in-transparent, it remains open to speculation whether authors affiliated with German universities and research organisations avoid opting for open access when publishing in hybrid journals, or simply use other budgets that are not covered by the Open APC initiative.
+One possible explanation for why traditional publishers are less well represented in our study is the lack of cost information about hybrid open access journals.
+In fact, 99 % of all articles that German universities and research organisations supported financially were published in fully open access journals.
+This result presumably reflects the DFG funding policy that opposes paying for open access articles published in hybrid journals.
+However, while reviewing self-reported cost data from Austria and the UK, where hybrid open access journals are generally supported, we observed a much higher share of payments for articles in hybrid open access journals.
+Because publication fee spending is fragmented and often lacks transparency, it remains open to speculation whether authors affiliated with German universities and research organisations avoid opting for open access when publishing in hybrid journals or whether they simply use other budgets that are not covered by the Open APC initiative.
 

### 2016-06-10 article.Rmd
proofreading conclusion

@@
 
-Our study reveals the size and extent of spending on open access journals using publication fees in Germany.
-Drawing on self-reported cost data from the Open APC initiative, support from German universities and research institutions have grown over the years.
-Comparing these expenditure with those from Austria and the UK, German open access funding is mostly focussed on fully open access journals, raising important questions about hybrid open access journals as publication venue.
-Given our findings, and in light of the general discussion about funding policies addressing hybrid open access journals, questions about whether and to what extent science policy interventions, as well as the availability of institutional support, influence how researchers publish  are of particular concern.
+Our study revealed the size and extent of spending on open access journals using publication fees in Germany.
+According to self-reported cost data from the Open APC initiative, this type of support from German universities and research institutions has grown over the years.
+Comparing these expenditure with those from Austria and the UK, German open access funding is focussed primarily on fully open access journals, raising important questions about hybrid open access journals as a publication venue.
+Given our findings and the general discussion about funding policies addressing hybrid open access journals, questions about whether and to what extent science policy interventions and the availability of institutional support influence how researchers publish are of particular concern.
 
-Using self-reported data and gathering publisher and journal information from Crossref, our study extends methods and improves data collection activities for researchers and practitioners alike, as well as contribute to a better understanding of factors affecting the analysis of publication fees in open access publishing.
-In this regard, our research highlights large variation in the distribution of spending that needs to be taken into consideration when studying payment on publication at the institutional level.
-We have also confirmed previous studies showing large price variations across publishers and open access journals that need to be better understood in future.
+Using self-reported data and gathering publisher and journal information from Crossref, our study extends methods and improves data collection activities for researchers and practitioners alike, as well as contributing to a better understanding of the factors affecting the analysis of publication fees in open access publishing.
+In this regard, our research highlights large variations in the distribution of spending that need to be taken into consideration when studying payments on publications at the institutional level.
+We have also confirmed previous studies that showed large price variations across publishers and open access journals; these variations need to be better understood in future.
 
-# Acknowledgment{.unnumbered}
+# Acknowledgement{.unnumbered}
 
-We thank Andrea Hacker and Ada-Charlotte Regelmann for valuable comments on the first draft of this paper.
-We also thank Christoph Broschinski, Vitali Peil, and Dirk Pieper, the members of the DINI working group "Electronic publishing", and all data contributors[^20] of the Open APC initiative.
+We thank Andrea Hacker and Ada-Charlotte Regelmann for their valuable comments on the first draft of this paper.
+We also thank Christoph Broschinski, Vitali Peil, Dirk Pieper, the members of the DINI working group "Electronic publishing", and all data contributors[^20] of the Open APC initiative.
 

### 2016-06-10 article.Rmd
fix superscript in scoap3

@@
 Although making such payments from several budgets is a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro-rata payments were not made transparent in the Open APC data, leading to a possibly flawed representation of publication fee spending in Germany.
-In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the number of articles published by their authors in SCOAP^3-covered journals.[^26] This approach is arbitrary, because averages for an institution can be determined only after the end of a full 3-year funding cycle.
+In another case, one university included its contributions to the SCOAP^3^ consortia and presumably divided the sum by the number of articles published by their authors in SCOAP^3^-covered journals.[^26] This approach is arbitrary, because averages for an institution can be determined only after the end of a full 3-year funding cycle.
 Other factors affecting price variations are exchange rates and different tax rates for some organisations in Germany.
@@
 
-[^26]: SCOAP^3, the Sponsoring Consortium for Open Access Publishing in Particle Physics, is a unique approach to convert former subscription journals in high-energy physics to open access journals under a CC BY license.
+[^26]: SCOAP^3^, the Sponsoring Consortium for Open Access Publishing in Particle Physics, is a unique approach to convert former subscription journals in high-energy physics to open access journals under a CC BY license.
 The consortiums, led by CERN, pays publishers centrally, based on previously agreed APCs and an overall price cap, and retrieves its funds from organisations and countries based on their share in the articles published in the covered journals.

### 2016-06-10 article.Rmd
fixed typos/numbering

@@
 
-## Comparison of related cost data-sets
+## Comparison of related cost data-sets{.unnumbered}
 
@@
 [^26]: SCOAP^3^, the Sponsoring Consortium for Open Access Publishing in Particle Physics, is a unique approach to convert former subscription journals in high-energy physics to open access journals under a CC BY license.
-The consortiums, led by CERN, pays publishers centrally, based on previously agreed APCs and an overall price cap, and retrieves its funds from organisations and countries based on their share in the articles published in the covered journals.
+The consortium, led by CERN, pays publishers centrally, based on previously agreed APCs and an overall price cap, and retrieves its funds from organisations and countries based on their share in the articles published in the covered journals.
 German universities participate through an initiative led by the German National Library of Science and Technology that received additional funding from DFG.

### 2016-06-10 article.Rmd
added scoap3 website

@@
 
-[^26]: SCOAP^3^, the Sponsoring Consortium for Open Access Publishing in Particle Physics, is a unique approach to convert former subscription journals in high-energy physics to open access journals under a CC BY license.
+[^26]: SCOAP^3^, the Sponsoring Consortium for Open Access Publishing in Particle Physics, is a unique approach to convert former subscription journals in high-energy physics to open access journals under a CC BY license, see <https://scoap3.org/what-is-scoap3/>.
 The consortium, led by CERN, pays publishers centrally, based on previously agreed APCs and an overall price cap, and retrieves its funds from organisations and countries based on their share in the articles published in the covered journals.

### 2016-06-10 article.Rmd
some minor fixes

--- 2016-06-10 article.Rmd
@@
 ```
+**Table 2: Comparison of data by journal type and financial period**
 
@@
 We used the DOI to automatically fetch publisher and journal names for each article from the Crossref REST API.
-Table 2 shows the top ten publishers in terms of the number of financially supported articles.
+Table 3 shows the top ten publishers in terms of the number of financially supported articles.
 These publishers represented `r apc_publisher %>% filter(!publisher == "Other") %>% nrow() / my_apc %>% nrow() * 100` % of all articles included in our data-set.
@@
 Comparing these data with data from the UK, we observed that a greater share of total spending was directed to some open access publishers.
-@Pinfield_2015, for instance, reported remarkably lower numbers for the open access publishers MPDI, Copernicus, and Hindawi.
+@Pinfield_2015, for instance, reported remarkably lower proportions for the open access publishers MPDI, Copernicus, and Hindawi.
 
-**Table 2: Publication fees paid per publisher (in €)**
+**Table 3: Publication fees paid per publisher (in €)**
 
@@
 Prices also varied within journals.
-Table 3 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals, based on the number of supported articles.
+Table 4 illustrates the top ten out of `r my_apc %>% group_by(journal_full_title) %>% summarize(n()) %>% nrow()` journals, based on the number of supported articles.
 We normalised PLOS journal titles because the name change from "PLoS" to "PLOS" was only partially represented in the Crossref metadata at the time of our study.
@@
 
-Re-using DOIs to gather bibliographic metadata from Crossref is a promising approach to addressing data curation issues while unifying UK spending data [@Neylon_2014; @woodward_2014].
+Re-using DOIs to gather bibliographic metadata from Crossref is a promising approach to addressing data curation issues raised by UK initiatives [@Neylon_2014; @woodward_2014].
 In our study, Crossref thoroughly indexed open access journal articles disclosed in the Open APC data-set, providing information about publisher and journal titles for 99 % of all articles included in the Open APC data-set.
@@
 Adding to this complexity, it is likely that some institutions paid only part of the publication fee.
-Consider, for instance, the journal Nature Communication: Charges reported to the Open APC initiative ranged between € 2000, the DFG price cap, and € 4,403 for a journal that can be categorised as a pricy, high-impact journal according to @Solomon_2012.
+Consider, for instance, the journal Nature Communication, a journal that can be categorised as a pricy, high-impact journal according to @Solomon_2012: Charges reported to the Open APC initiative ranged between € 2000, the DFG price cap, and € 4,403.
 Although making such payments from several budgets is a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro-rata payments were not made transparent in the Open APC data, leading to a possibly flawed representation of publication fee spending in Germany.
@@
 In fact, 99 % of all articles that German universities and research organisations supported financially were published in fully open access journals.
-This result presumably reflects the DFG funding policy that opposes paying for open access articles published in hybrid journals.
+This result presumably reflects the DFG funding programme that excludes paying for open access articles published in hybrid journals.
 However, while reviewing self-reported cost data from Austria and the UK, where hybrid open access journals are generally supported, we observed a much higher share of payments for articles in hybrid open access journals.
@@
 Comparing these expenditure with those from Austria and the UK, German open access funding is focussed primarily on fully open access journals, raising important questions about hybrid open access journals as a publication venue.
-Given our findings and the general discussion about funding policies addressing hybrid open access journals, questions about whether and to what extent science policy interventions and the availability of institutional support influence how researchers publish are of particular concern.
+Given our findings and the general discussion about funding policies addressing hybrid open access journals, questions about whether and to what extent science policies and the availability of institutional support influence how researchers publish are of particular concern.
 
@@
 In this regard, our research highlights large variations in the distribution of spending that need to be taken into consideration when studying payments on publications at the institutional level.
-We have also confirmed previous studies that showed large price variations across publishers and open access journals; these variations need to be better understood in future.
+We have also confirmed the findings of other studies that showed large price variations across publishers and open access journals.
+This complex situation of fee-based open access publishing need to be better understood by both researchers and practitioners.
 

### 2016-06-10 data/cost_data_uk/README.Rmd
minor

@@
 
-sum(tt$`Fees paid in GBP`)
+sum(tt$`Fees paid in Euro`)
 
@@
 
-### wellcome 2014 - 2014
+### wellcome 2014 - 2015
 
@@
 ```
+
+### open apc
+
+```{r}
+my_apc <- jsonlite::stream_in(file("../cr-apc-all.json"), verbose = FALSE)
+
+# data cleaning
+library(dplyr)
+## remove Bielefelds Zero value
+my_apc <- dplyr::filter(my_apc, euro > 0)
+# remove intech book chapters
+my_apc <-  dplyr::filter(my_apc, !is.na(journal_full_title))
+
+# remove Austrian FWF data
+my_apc <- dplyr::filter(my_apc, !institution == "FWF - Austrian Science Fund")
+# data from 2014 and 2015
+my_apc <- dplyr::filter(my_apc, period == "2015" | period == "2014")
+
+aggregate(my_apc$euro, by = list(my_apc$is_hybrid, my_apc$period), sum)
+aggregate(my_apc$euro, by = list(my_apc$is_hybrid, my_apc$period), mean)
+aggregate(my_apc$euro, by = list(my_apc$is_hybrid, my_apc$period), length)
+```

### 2016-06-10 article.Rmd
minor corrections

@@
 The first is that payment of these charges is fragmented across the budgets of grant agencies, research institutions, and libraries, or is covered by personal budgets.
-A comprehensive 2015 survey asked 9,645 authors from various disciplines how they financed publication fees, and it revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.
+A comprehensive 2010 survey asked 9,645 authors from various disciplines how they financed publication fees, and it revealed that the majority of the respondents had access to research funding or institutional support to cover these charges.
 By contrast, 12 % paid publication fees individually [@soap].
-These results are consistent with similar findings from other studies.
-Previous studies also found that funding sources exist primarily in higher-income countries, mainly to support research articles in the biological and physical sciences [@Solomon_2011].
+These results are consistent with similar findings from other studies: Previous studies also found that funding sources exist primarily in higher-income countries, mainly to support research articles in the biological and physical sciences [@Solomon_2011].
 Personal budgets, however, are likely used to cover lower publication fees [@Solomon_2011; @Bj_rk_2015].
@@
 Disclosed as publicly available spreadsheets, these data-sets self-report expenditures along with bibliographic information, including title, journal and publisher, and a persistent identifier for the publisher's version.
-Curatorial efforts focussed on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers to the full text including the Digital Object Identifier (DOI) [@Neylon_2014; @woodward_2014].
+Curatorial efforts focussed on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers for the full text including the Digital Object Identifier (DOI) [@Neylon_2014; @woodward_2014].
 A preliminary version of Jisc's cost data was examined by @Pinfield_2015.
@@
 We excluded articles from the analysis with missing information about the cost or the journal type.
-In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to € 4.861.772 from the analysis because no publication type was given in the data-set.
+In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to € 4,861,772 from the analysis because no publication type was given in the data-set.
 
@@
 
-The fees for all of the articles amounted to € `r sum(my_apc$euro)`, including value-added tax; the average payment was € `r mean(my_apc$euro)` (median =  € `r median(my_apc$euro)`) .
+The fees for all of the articles amounted to € `r sum(my_apc$euro)`, including value-added tax; the average payment was € `r mean(my_apc$euro)` (median =  € `r median(my_apc$euro)`, SD = € `r sd(my_apc$euro)`).
 Figure 2 presents the distribution of institutional spending on publications.
@@
 Large price variations can be observed.
-Publication fees that were paid by German universities and research organisations ranged from € `r min(my_apc$euro)` to € `r max(my_apc$euro)` (SD = `r sd(my_apc$euro)`).
+Publication fees that were paid by German universities and research organisations ranged from € `r min(my_apc$euro)` to € `r max(my_apc$euro)`.
 However, the average price paid varied somewhat during the period from 2011 to 2015 (€ `r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - € `r my_apc %>% filter(period == 2015) %>% summarise(mean(euro)) %>% round()`).
@@
 
-[^24]: <https://github.com/OpenAPC/openapc-de/tree/v2.4.3>
+[^24]: <hhttps://github.com/OpenAPC/openapc-de/releases/tag/v2.4.3>
 

### 2016-06-11 article.Rmd
fixed typo
modified csl to list up to 100 authors

@@
 bibliography: literatur.bib
-csl: frontiers.csl
+csl: frontiers100.csl
 
@@
 
-[^24]: <hhttps://github.com/OpenAPC/openapc-de/releases/tag/v2.4.3>
+[^24]: <https://github.com/OpenAPC/openapc-de/releases/tag/v2.4.3>
 

### 2016-06-11 article.Rmd
revised version

@@
 Although the Open APC initiative gathered metadata representing publishers and journals from Crossref as well, this information was retrieved at the time when the participating institutions submitted their expenditure.
-To be transparent over time, the Open APC initiative kept track of the date when these data-sets were submitted with Git, a version control system that is increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub  However, during these data collection activities, Crossref regularly updated the metadata to represent ongoing mergers of publishing houses.
+To be transparent over time, the Open APC initiative kept track of the date when these data-sets were submitted with Git, a version control system that is increasingly used for enabling reproducible research [@Ram_2013], and made this information available via GitHub.
+However, during these data collection activities, Crossref regularly updated the metadata to represent ongoing mergers of publishing houses.
 A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, which operated as Springer Nature at the time of our study.
@@
   scale_x_continuous(limits = c(0,8000)) +
-  xlab("Publication fee in €") +
+  xlab("Publication fee in Euro") +
   ylab("Articles") +
@@
 
-**Figure 2: Distribution of institutional spending on publication fees by German research organisations**
+**Figure 2: Institutional spending on publication fees by German research organisations per article (in €)**
 
-**Figure 3: Institutional spending on publication fees by German research organisations per year**
+**Figure 3: Institutional spending on publication fees by German research organisations per year (in €)**
 
@@
   xlab("Year") +
-  ylab("Publication fee paid in €") +
+  ylab("Publication fee paid in Euro") +
   scale_y_continuous(limits = c(0,8000)) + 
@@
 These findings are consistent with the general trend towards using publication fees as a revenue source for open access publishing [@Davis_2011; @Laakso_2012; @Pinfield_Review_2015].
-They also demonstrate the growing trend among institutions to both encourage their researchers to publish in open access journals and to offer financial support [@Fournier_2013].
+They also demonstrate the growing trend among institutions in Germany to both encourage their researchers to publish in open access journals and to offer financial support [@Fournier_2013].
 Similar to the expenditures on publication fees at an institutional level in the UK [@Pinfield_2015], spending volume varies across German universities and research organisations.
@@
 To increase the transparency of publication fee spending, @Pinfield_2015 suggested disclosing tax rates and payment currencies in future cost data-sets.
-In addition, the Open APC data-set does not track funding sources; thus, we could not determine which funders co-financed publication fees.
+Likewise, the Open APC data-set does not track funding sources; thus, we could not determine which funders co-financed publication fees.
 
